### 2020-03-06 sources/tools/make-tableau-data.Rmd
start on tableau test

---
title: "Make Database for Tableau Prototype"
output: html_document
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse",
                     "jsonlite",
                     "visioneval")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

# import::here(ve.export,.from="./exporter.R")
```

# Remote I/O
```{r remote-io}
viewer_dir <- "~/Documents/GitHub/VisionEval-Dev/sources/VEScenarioViewer/data/VERSPM/"
output_dir <- "./example_dir/tableau_test/"

viewer_file_name <- paste0(viewer_dir,  "verspm.js")

scenario_config_file_name <- paste0(output_dir, "example_scenario_manage.csv")

output_file_name <- paste0(output_dir, "database-for-tableau.csv")
```

# Convert Javascript data to CSV
```{r convert-java}
temp_out_name <- paste0(output_dir, "verspm.JSON")
input  <- readLines(viewer_file_name)
  step_1 <- gsub(pattern = "var data =", replace = "", x = input)
  step_2 <- str_replace(step_1, ";", "")
  writeLines(step_2, con = temp_out_name)

temp_df <- fromJSON(temp_out_name, flatten = TRUE)

write_csv(temp_df, path = paste0(output_dir, "verspm.csv"))
```

# Prepare Tableau
```{r prep-tableau}
measure_names_vector <- c("GHGReduction", 
                          "DVMTPerCapita",
                          "WalkTravelPerCapita",
                          "TruckDelay",
                          "AirPollutionEm",
                          "FuelUse",
                          "VehicleCost",
                          "VehicleCostLow")

config_df <- read_csv(scenario_config_file_name, 
                      col_types = cols(.default = col_character(), execute = col_logical()))

category_name_dict_df <- distinct(config_df, category_name, category_label) %>%
  mutate(category_label = str_replace(category_label, " ", "_")) %>%
  select(short = category_name, long = category_label)

level_name_dict_df <- distinct(config_df, category_name, level_name, level_label)

outcomes_df <- read_csv(paste0(output_dir, "verspm.csv"),
                        col_types = cols(.default = col_character())) %>%
  mutate_at(vars(measure_names_vector), as.numeric)

working_df <- outcomes_df

for (category in unique(level_name_dict_df$category_name)) {

  level_dict_df <- level_name_dict_df %>%
    filter(category_name == category) %>%
    distinct(level_name, level_label)

  new_category_name <- paste0(category, "_NAME")

  working_df <- working_df %>%
    rename(subject := !!category) %>%
    left_join(., level_dict_df, by = c("subject" = "level_name")) %>%
    mutate(subject_copy = level_label) %>%
    rename(!!new_category_name := subject_copy) %>%
    rename(!!category := subject) %>%
    select(-level_label)

}

for (category in category_name_dict_df$short) {

  long_name <- filter(category_name_dict_df, short == category)$long

  new_category_name <- paste0(category, "_NAME")

  working_df <- working_df %>%
    rename(!!long_name := !!new_category_name) 
}

output_df <- working_df %>%
  mutate(scenario_index = row_number())

```

# Write
```{r}
write_csv(output_df, path = output_file_name)
```

### 2020-03-10 sources/tools/make-tableau-data.Rmd
beta version functioning

@@
 
+# TODO: 
+# make a vignette, 
+# make a proper tool (in R)
+
 # Overhead
@@
 packages_vector <- c("tidyverse",
-                     "jsonlite",
-                     "visioneval")
+                     "jsonlite")
 
@@
 
-# import::here(ve.export,.from="./exporter.R")
 ```
@@
 
-viewer_file_name <- paste0(viewer_dir,  "verspm.js")
-
-scenario_config_file_name <- paste0(output_dir, "example_scenario_manage.csv")
-
 output_file_name <- paste0(output_dir, "database-for-tableau.csv")
@@
 
-# Convert Javascript data to CSV
-```{r convert-java}
-temp_out_name <- paste0(output_dir, "verspm.JSON")
-input  <- readLines(viewer_file_name)
-  step_1 <- gsub(pattern = "var data =", replace = "", x = input)
-  step_2 <- str_replace(step_1, ";", "")
+# Private Methods
+```{r methods}
+.MakeDataframeFromJavaScript <- function(input_js_var_file_name, var_name_string) {
+
+  temp_out_name <-  paste0(output_dir, "temp.JSON")
+  input_line  <- readLines(input_js_var_file_name)
+  step_1 <- stringr::str_replace(input_line, paste0("var ", var_name_string, " ="), "")
+  step_2 <- stringr::str_replace(step_1, ";", "")
   writeLines(step_2, con = temp_out_name)
 
-temp_df <- fromJSON(temp_out_name, flatten = TRUE)
+  return_df <- jsonlite::fromJSON(temp_out_name, simplifyVector = TRUE, flatten = TRUE)
 
-write_csv(temp_df, path = paste0(output_dir, "verspm.csv"))
+  return(return_df)
+
+}
+
+.MakeScenarioDataframes <- function(input_scenario_dir, model_name = "verspm") {
+
+  file_name <-  paste0(input_scenario_dir, model_name, ".js")
+  results_df <- .MakeDataframeFromJavaScript(file_name, "data")
+
+  file_name <-  paste0(input_scenario_dir, "output-cfg.js")
+  output_config_df <- .MakeDataframeFromJavaScript(file_name, "outputcfg")
+
+  file_name <-  paste0(input_scenario_dir, "scenario-cfg.js")
+  scenario_config_df <- .MakeDataframeFromJavaScript(file_name, "scenconfig") %>%
+    dplyr::rename(category_name = NAME, 
+                  category_label = LABEL, 
+                  category_description = DESCRIPTION,
+                  category_instructions = INSTRUCTIONS) %>%
+    tidyr::unnest(.,cols = c(category_name, category_label, category_description, 
+                             category_instructions, LEVELS)) %>%
+    dplyr::bind_cols(., tibble::tibble(level_name = unlist(.$NAME))) %>%
+    dplyr::bind_cols(., tibble::tibble(level_label = unlist(.$LABEL))) %>%
+    dplyr::bind_cols(., tibble::tibble(level_description = unlist(.$DESCRIPTION))) %>%
+    dplyr::select(-NAME, -LABEL, -DESCRIPTION)
+
+  file_name <-  paste0(input_scenario_dir, "category-cfg.js")
+  category_config_df <- .MakeDataframeFromJavaScript(file_name, "catconfig") %>%
+    dplyr::rename(strategy_label = NAME,
+                  strategy_description = DESCRIPTION) %>%
+    tidyr::unnest(., cols = c(strategy_label, strategy_description, LEVELS)) %>%
+    dplyr::rename(strategy_bundle = NAME) %>%
+    tidyr::unnest(., cols = c(strategy_bundle, INPUTS)) %>%
+    dplyr::bind_cols(., tibble::tibble(category_name = unlist(.$NAME))) %>%
+    dplyr::bind_cols(., tibble::tibble(level_name = unlist(.$LEVEL))) %>%
+    dplyr::select(-LEVEL, -NAME)
+
+  return_list <- list("results" = results_df,
+                      "output_config" = output_config_df,
+                      "scenario_config" = scenario_config_df,
+                      "category_config" = category_config_df)
+
+  return(return_list)
+}
+
+.notAllNA <- function(x) {
+
+  return(any(!is.na(x)))
+
+}
+
 ```
 
-# Prepare Tableau
-```{r prep-tableau}
-measure_names_vector <- c("GHGReduction", 
-                          "DVMTPerCapita",
-                          "WalkTravelPerCapita",
-                          "TruckDelay",
-                          "AirPollutionEm",
-                          "FuelUse",
-                          "VehicleCost",
-                          "VehicleCostLow")
+# Public Methods
+```{r public-methods}
+MakeTableauScenarioViewer <- function(input_scenario_dir,  model_name = "verspm") {
 
-config_df <- read_csv(scenario_config_file_name, 
-                      col_types = cols(.default = col_character(), execute = col_logical()))
+  df_list <- .MakeScenarioDataframes(input_scenario_dir, model_name)
 
-category_name_dict_df <- distinct(config_df, category_name, category_label) %>%
-  mutate(category_label = str_replace(category_label, " ", "_")) %>%
-  select(short = category_name, long = category_label)
+  measure_names_vector <- df_list$output_config$COLUMN
 
-level_name_dict_df <- distinct(config_df, category_name, level_name, level_label)
+  outcomes_df <- df_list$results %>%
+    dplyr::mutate_at(dplyr::vars(all_of(measure_names_vector)), as.numeric)
 
-outcomes_df <- read_csv(paste0(output_dir, "verspm.csv"),
-                        col_types = cols(.default = col_character())) %>%
-  mutate_at(vars(measure_names_vector), as.numeric)
+  category_name_dict_df <- df_list$scenario_config %>%
+    dplyr::distinct(category_name, category_label) %>%
+    dplyr::mutate(category_label = stringr::str_replace_all(category_label, " ", "_")) %>%
+    dplyr::select(short = category_name, long = category_label)
+
+  level_name_dict_df <- df_list$scenario_config %>%
+    dplyr::distinct(category_name, level_name, level_label)
 
@@
     level_dict_df <- level_name_dict_df %>%
-    filter(category_name == category) %>%
-    distinct(level_name, level_label)
+      dplyr::filter(category_name == category) %>%
+      dplyr::distinct(level_name, level_label)
 
@@
     working_df <- working_df %>%
-    rename(subject := !!category) %>%
-    left_join(., level_dict_df, by = c("subject" = "level_name")) %>%
-    mutate(subject_copy = level_label) %>%
-    rename(!!new_category_name := subject_copy) %>%
-    rename(!!category := subject) %>%
-    select(-level_label)
+      dplyr::rename(subject := !!category) %>%
+      dplyr::left_join(., level_dict_df, by = c("subject" = "level_name")) %>%
+      dplyr::mutate(subject_copy = level_label) %>%
+      dplyr::rename(!!new_category_name := subject_copy) %>%
+      dplyr::rename(!!category := subject) %>%
+      dplyr::select(-level_label)
 
@@
 
-  long_name <- filter(category_name_dict_df, short == category)$long
+    long_name <- dplyr::filter(category_name_dict_df, short == category)$long
 
@@
     working_df <- working_df %>%
-    rename(!!long_name := !!new_category_name) 
+      dplyr::rename(!!long_name := !!new_category_name) 
   }
 
-output_df <- working_df %>%
-  mutate(scenario_index = row_number())
+  strategy_name_df <- df_list$category_config %>%
+    tidyr::pivot_wider(., names_from = category_name, values_from = level_name)
 
+  for (strategy in unique(strategy_name_df$strategy_label)) {
+
+    join_df <- dplyr::filter(strategy_name_df, strategy_label == strategy) %>%
+      dplyr::select(-strategy_description) %>%
+      dplyr::select_if(.notAllNA)
+
+    join_vector <- colnames(dplyr::select(join_df, -strategy_label, -strategy_bundle))
+
+    strategy <- stringr::str_replace_all(strategy, " ", "_")
+    strategy <- stringr::str_replace_all(strategy, "\\/", "_")
+    strategy <- paste0("Strategy_", strategy)
+
+    working_df <- dplyr::left_join(working_df, join_df, by = join_vector) %>%
+      dplyr::rename(!!strategy := strategy_bundle) %>%
+      dplyr::select(-strategy_label)
+
+  }
+
+  return_df <- working_df %>%
+    dplyr::mutate(scenario_index = dplyr::row_number())
+
+  return(return_df)
+
+}
+```
+
+# Apply Method
+```{r apply-method}
+output_df <- MakeTableauScenarioViewer(viewer_dir, model_name = "verspm")
 ```

### 2020-03-11 sources/tools/Tableau Scenario Viewer Vignette.Rmd
ready for PR

---
title: "Tableau Scenario Viewer Vignette"
output:
  html_document: 
    theme: cosmo
---

*tableau_scenario_viewer.R (defines ve.scenario_management tools)*

This file documents the `ve.tableau_scenario_viewer` tools for VisionEval, which allows the user to view scenarios in [Tableau](https://www.tableau.com/) rather than the custom JavaScript HTML viewer distributed as part of VisionEval.

## Background

A powerful feature of VisionEval is the ability to quickly run 100s or even 1000s of scenarios. Background on running scenarios in VisionEval is available in the [project documentation](https://github.com/VisionEval/VisionEval/wiki/Multiple-Scenarios). Scenarios outcomes can be viewed in a web browser using a HTML viewer distributed as part of VisionEval. The advantages of the HTML viewer are that it is free and can be viewed in a web browser. The cons are that it is difficult to modify/customize.

Tableau is a popular data visualization software product. The tool is expensive, but allows for a rich set of visualizations that can be quickly developed and modified. We have created a set of tools that allow users to explore VisionEval scenario outputs or in Tableau.

## Tableau Scenario Viewer  Demonstration
As an alternative to the HTML viewer, we have created a method that creates a Tableau-ready database using the files created for the HTML viewer. Users can then import this database into Tableau to create their own visualizations, or start with the `Scenario Viewer Prototype.twb` Tableau workbook we have provided.

Here's how the tools work.

To access the tools you must, let R know where they are. The code below sets the working directory to `VisionEval/sources/`. Please modify the code to point to where you have saved the `VisionEval/sources` folder.
```{r setup}
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/VisionEval-Dev/sources/")
```

We can now load the VisionEval scenario management tools.
```{r load-tools, message = FALSE}
source("tools/tableau_scenario_viewer.R")
```

To convert the HTML viewer inputs into a Tableau-ready CSV database, the user must specify the following three inputs:
1. The location of the HTML viewer inputs;
2. Whether the outputs are specific to VERSPM or VERPAT; and,
3. The location of the output Tableau-ready CSV database file.

For example, these values can be:
```{r set-inputs}
viewer_dir <- "./VEScenarioViewer/data/VERSPM/"
model_name <- "verspm"
output_file_name <- paste0("./tools/example_dir/database-for-tableau.csv")
```

We can now call the `ve.tableau_scenario_viewer.MakeCSV` method, which returns a Tableau-ready dataframe and then write this dataframe to disk as follows.

```{r method-write}
output_df <- ve.tableau_scenario_viewer.MakeCSV(viewer_dir, model_name)
readr::write_csv(output_df, path = output_file_name)
```

You can now open up the `Scenario Viewer Prototype.twb` Tableau workbook and point to the output data file.

### 2020-03-11 sources/tools/tableau_scenario_viewer.R

@@ -1,40 +1,19 @@
----
-title: "Make Database for Tableau Prototype"
-output: html_document
----
+requireNamespace("magrittr")
+requireNamespace("dplyr")
+requireNamespace("tidyr")
+requireNamespace("readr")
+requireNamespace("jsonlite")
 
-# TODO: 
-# make a vignette, 
-# make a proper tool (in R)
+library(magrittr)
 
-# Overhead
-```{r overhead, include = FALSE}
-packages_vector <- c("tidyverse",
-                     "jsonlite")
+tool.contents <- c(
+  "ve.tableau_scenario_viewer.MakeCSV"
+)
 
-need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]
-
-if (length(need_to_install)) install.packages(need_to_install)
-
-for (package in packages_vector) {
-  library(package, character.only = TRUE)
-}
-
-```
-
-# Remote I/O
-```{r remote-io}
-viewer_dir <- "~/Documents/GitHub/VisionEval-Dev/sources/VEScenarioViewer/data/VERSPM/"
-output_dir <- "./example_dir/tableau_test/"
-
-output_file_name <- paste0(output_dir, "database-for-tableau.csv")
-```
-
-# Private Methods
-```{r methods}
+# private methods --------------------------------------------------------------
 .MakeDataframeFromJavaScript <- function(input_js_var_file_name, var_name_string) {
   
-  temp_out_name <-  paste0(output_dir, "temp.JSON")
+  temp_out_name <- "temp.JSON"
   input_line  <- readLines(input_js_var_file_name)
   step_1 <- stringr::str_replace(input_line, paste0("var ", var_name_string, " ="), "")
   step_2 <- stringr::str_replace(step_1, ";", "")
@@ -42,11 +21,18 @@ output_file_name <- paste0(output_dir, "database-for-tableau.csv")
   
   return_df <- jsonlite::fromJSON(temp_out_name, simplifyVector = TRUE, flatten = TRUE)
   
+  file.remove(temp_out_name)
+  
   return(return_df)
   
 }
 
 .MakeScenarioDataframes <- function(input_scenario_dir, model_name = "verspm") {
+  
+  if (!(model_name %in% c("verspm", "verpat"))) {
+    stop("model_name must be 'verspm' or 'verpat'")
+    return("Error")
+  }
   
   file_name <-  paste0(input_scenario_dir, model_name, ".js")
[…]

### 2020-03-18 sources/tools/Tableau Scenario Viewer Vignette.Rmd
magrittr assign, tableau path

@@
 model_name <- "verspm"
-output_file_name <- paste0("./tools/example_dir/database-for-tableau.csv")
+output_file_name <- "./tools/example_dir/database-for-tableau.csv"
 ```

### 2020-03-18 sources/tools/tableau_scenario_viewer.R

@@ -4,7 +4,7 @@ requireNamespace("tidyr")
 requireNamespace("readr")
 requireNamespace("jsonlite")
 
-library(magrittr)
+assign("%>%",getFromNamespace("%>%","magrittr"))
 
 tool.contents <- c(
   "ve.tableau_scenario_viewer.MakeCSV"

### 2020-06-30 sources/modules/VEScenario/R/NewScenarioSet.R
Add NewScenarioSet module to VEScenario

@@ -1,0 +1,313 @@
+#=================
+#NewScenarioSet.R
+#=================
+# This module builds scenarios from the combinations scenario input levels.
+
+
+
+
+
+
+library(visioneval)
+library(jsonlite)
+
+#=================================================
+#SECTION 1: ESTIMATE AND SAVE Scenario PARAMETERS
+#=================================================
+
+
+#================================================
+#SECTION 2: DEFINE THE MODULE DATA SPECIFICATIONS
+#================================================
+
+NewScenarioSetSpecifications <- list(
+  # Level of geography module is applied at
+  RunBy = "Region",
+  # Specify new tables to be created by Inp if any
+  # Specify new tables to be created by Set if any
+  # NewSetTable
+  # Specify Input data
+  # Specify data to be loaded from the datastore
+  Get = items(
+
+  item(
+    NAME = "ModelFolder",
+    TABLE = "Model",
+    GROUP = "Global",
+    TYPE = "character",
+    UNITS = "NA",
+    SIZE = 20,
+    PROHIBIT = "NA",
+    ISELEMENTOF = ""
+  )
+  ,
+  
+    item(
+      NAME = "ScenarioInputFolder",
+      TABLE = "Model",
+      GROUP = "Global",
+      TYPE = "character",
+      UNITS = "NA",
+      PROHIBIT = "NA",
+      ISELEMENTOF = "",
+      DESCRIPTION = "Scenario Input folder"
+    
+  )
+  ,
+  item(
+    NAME = "ScenarioManagerFile",
+    TABLE = "Model",
+    GROUP = "Global",
+    TYPE = "character",
+    UNITS = "NA",
+    PROHIBIT = "NA",
+    ISELEMENTOF = "",
+    DESCRIPTION = "name of scenario management csv database"
+    
+  )
+),
+Set = items(
+  item(
+    NAME = "Scenarioset",
+    TABLE = "Model",
+    GROUP = "Global",
+    TYPE = "integer",
+    UNITS = "NA",
+    PROHIBIT = "NA",
+    ISELEMENTOF = "",
+    DESCRIPTION = "Returns 1 if scenario set build is complete"
+  )
+)
+)
+
+#Save the data specifications list
+#---------------------------------
+#' Specifications list for NewScenarioSet module
+#'
+#' A list containing specifications for the NewScenarioSet module.
+#'
+#' @format A list containing 4 components:
+#' \describe{
[…]

### 2020-06-30 sources/models/VERPAT_Scenarios/tutorial.Rmd
update to model folders for VERSPM_Scenarios and VERPAT Scenarios

---
title: "Introduction to VERPAT_Scenario"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to VERPAT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Strategic planning often requires the assessment of large numbers of future scenarios, each assessing a different combination of prioritizations, policy decisions, and constraints. The VEScenarios module provides the capability to quickly set up, run, and visualize large numbers of VERSPM scenarios using a baseline scenario combined with multiple changes to model inputs.

Both VERSPM and VERPAT models are capable of running multiple scenarios efficiently

### Scenario manager database
The first step to run the VERPAT_scnerios is to build the scenario manager database. A sample file is 
located in `scenario_root/verpat_scenario_manager.csv`. Scenarios should be defined in this table with the same format. 
At first level, you should define the strategies you want to run.

These are the current list of VERPAT_Scenarios strategies:

* **Bicycles** :  Network improvements, incentives, and technologies that encourage bicycling and other light-weight vehicle travel. ( group of B ) 
* **Demand Management** : Programs to encourage less private vehicle travel.( group of D ) 
* **Transit** Extent of and charges for priced parking. ( T )
* **Land Use** : Factors representing changes to future vehicles and fuels ( group of L ) 
* **Vehicle Travel Cost** : Context factor on the assumed growth of statewide average per capita income(C)
* **Parking** : Extent of and charges for priced parking. ( P )

Each strategy covers couple of categories in a lower level.

These are the current list of VERSPM_Scenarios categories:

  - C - Vehicle Travel Cost (`model_parameters.json`) : Combination of fuel prices and charges to pay for roadway costs and possibly externalities.
  - B - Bicycles (`region_light_vehicles.csv`) : Network improvements, incentives, and technologies that encourage bicycling and other light-weight vehicle travel

  - D - DemandManagement (`region_commute_options.csv`) : Programs to encourage less private vehicle travel

  - L - LandUse (`bzone_pop_emp_prop.csv`) : Distribution of population and employment by place type

  - P - Parking (`marea_parking_growth.csv`) : The extent of paid parking and its price

  - T - Transit (`model_parameters.json`) : The extent and frequency of transit service

  - V - Vehicle Characteristics (`bzone_pop_emp_prop.csv`): The combination of fuel prices and vehicle travel charges to pay for roadways and to pay for externalities such as carbon pricing

 For each category, you should define the scenario levels you would like to run. For each scenario level you need to change an input based on the scenario definition.

 You need to have each scenario level as a row in this table. The input column should be populated with the modified input file names for each scenario level.

### defs

`VERPAT_Scenarios/defs` directory contains the same files as the `/defs` directory in VERSPM, but the `model_parameters.json` is different.:

```
[

  {
    "NAME": "ModelFolder",
    "VALUE": "VERPAT_base_model",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME": "ScenarioInputFolder",
    "VALUE": "scenario_root",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioManagerFile",
    "VALUE": "verspm_scenario_manager.csv",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 30,
   "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioOutputFolder",
    "VALUE": "scenarios",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 20,
   "ISELEMENTOF": ""
  },  
  {
    "NAME" : "NWorkers",
    "VALUE": "6",
    "TYPE" : "integer",
    "UNITS" : "NA",
    "PROHIBIT" : "c('NA', '< 0')",
    "ISELEMENTOF" : ""
  }
]

```
** Model folder : The VERSPM base model folder name 
** ScenarioInputFolder: Folder name for scenario root where scenario manager file is located. All the scenario level folders will be created here
** ScenarioManagerFile : Scenario manager database file name
** ScenarioOutputFolder: Name of the folder where all the scenarios are created

### Running VERPAT_Scnearios

Running VERSPM_Scenario has 5 major steps. After initalizing the module there are 5 modules to call

** NewScenarioSet : This module creates the required categories and scenarios json files from csv database. It also creates the folder structure and required inputs for all scenario levels  
** BuildScenarios: This module creates all possible scnearios which are defined in previous step. It copies the base VERSPM files for each scenario folder 
** RunScenarios : This module runs all scenarios 
** VERSPMResults: This model summarizes all different scenario outputs from datastor to standard format
** ViewResults:  This module create the visualization and summaries for different performance measures

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(visioneval)

ptm <- proc.time()

#Initialize model
#----------------
initializeModel(
  ModelScriptFile = "run_model.R",
  ParamDir = "defs",
  RunParamFile = "run_parameters.json",
  GeoFile = "geo.csv",
  ModelParamFile = "model_parameters.json",
  LoadDatastore = FALSE,
  DatastoreName = NULL,
  SaveDatastore = TRUE
)

#Run all demo module for all years
#---------------------------------
for(Year in getYears()) {

  runModule(
    ModuleName = "NewScenarioSet",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )

  runModule(
    ModuleName = "BuildScenarios",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "RunScenarios",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "VERPATResults",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "ViewResults",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
}

```

### 2020-06-30 sources/models/VERSPM_Scenarios/tutorial.Rmd
update to model folders for VERSPM_Scenarios and VERPAT Scenarios

---
title: "Introduction to VERRSPM_Scenario"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to VETRSPM_Scenario}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Strategic planning often requires the assessment of large numbers of future scenarios, each assessing a different combination of prioritizations, policy decisions, and constraints. The VEScenarios module provides the capability to quickly set up, run, and visualize large numbers of VERSPM scenarios using a baseline scenario combined with multiple changes to model inputs.

Both VERSPM and VERPAT models are capable of running multiple scenarios efficiently

### Scenario manager database
The first step to run the VERSPM_scnerios is to build the scenario manager database. A sample file is 
located in `scenario_root/verspm_scenario_manager.csv`. Scenarios should be defined in this table with the same format. 
At first level, you should define the strategies you want to run.

These are the current list of VERSPM_Scenarios strategies:

* **Community Design** :  Policies that seek to enable shorter trips and alternate modes such as promotion of mixed use land use, transit service, bicycling, and parking management. ( group of L, B, P , T ) 
* **Marketing/Incentive** : Policies that improve driving efficiency such as eco driving,and Intelligent Transportation System efforts, as well as programs that reduce auto demand such as carsharing, and home or work-based transportation demand management.( group of D, E ) 
* **Pricing : Policies** that move towards true cost pricing such as road user fees to pay for the cost of operating, maintaining and improving roads, pay-as-you-drive (PAYD) insurance, and environmental impact fees such as a carbon tax. ( C )
* **Vehicles/Fuels** : Factors representing changes to future vehicles and fuels ( group of V, F ) 
* **Income** : Context factor on the assumed growth of statewide average per capita income ( I )

Each strategy covers couple of categories in a lower level.

These are the current list of VERSPM_Scenarios categories:

  - B - Bicycles (`azone_prop_sov_dvmt_diverted.csv`) : Network improvements, incentives, and technologies that encourage bicycling and other light-weight vehicle travel

  - C - Vehicle Travel Cost (`azone_hh_veh_own_taxes.csv`) and (`region_prop_externalities_paid.csv`) : Combination of fuel prices and charges to pay for roadway costs and possibly externalities

  - D - DemandManagement (`bzone_travel_demand_mgt.csv`) : Programs to encourage less private vehicle travel

  - E - Driving Efficiency (`marea_operations_deployment.csv`) and (`marea_speed_smooth_ecodrive.csv`) and (`other_ops_effectiveness.csv`) : Driving efficiency by increasing implementation of ITS

  - F - Technology Mix and CI (`marea_transit_powertrain_prop.csv`) and (`region_carsvc_powertrain_prop.csv`) and (region_comsvc_powertrain_prop.csv`) : Vehicle technology mix and carbon intensity of fuels.

  - G - Fuel Price (`azone_fuel_power_cost.csv`) : Real fuel price in 2010 USD

  - I - Income (`azone_fuel_power_cost.csv`) : Real average household income in 2010 USD

  - L - LandUse (`azone_hhsize_targets.csv`) and (`bzone_urban_du_proportions.csv`) : Distribution of population and employment by place type

  - P - Parking (`bzone_parking.csv`) : The extent of paid parking and its price

  - T - Transit (`marea_transit_service.csv`) : The extent and frequency of transit service

  - V - Vehicle Characteristics (`azone_hh_veh_mean_age.csv`) and (`azone_lttrk_prop`): The combination of fuel prices and vehicle travel charges to pay for roadways and to pay for externalities such as carbon pricing

 For each category, you should define the scenario levels you would like to run. For each scenario level you need to change an input based on the scenario definition.

 You need to have each scenario level as a row in this table. The input column should be populated with the modified input file names for each scenario level.

### defs

`VERSPM_Scenarios/defs` directory contains the same files as the `/defs` directory in VERSPM, but the `model_parameters.json` is different.:

```
[

  {
    "NAME": "ModelFolder",
    "VALUE": "VERSPM_base_model",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME": "ScenarioInputFolder",
    "VALUE": "scenario_root",
    "TYPE": "character",
    "UNITS": "NA",
    "PROHIBIT": "NA",
    "SIZE": 20,
    "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioManagerFile",
    "VALUE": "verspm_scenario_manager.csv",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 30,
   "ISELEMENTOF": ""
  },
  {
    "NAME" : "ScenarioOutputFolder",
    "VALUE": "scenarios",
   "TYPE": "character",
   "UNITS": "NA",
   "PROHIBIT": "NA",
   "SIZE": 20,
   "ISELEMENTOF": ""
  },  
  {
    "NAME" : "NWorkers",
    "VALUE": "6",
    "TYPE" : "integer",
    "UNITS" : "NA",
    "PROHIBIT" : "c('NA', '< 0')",
    "ISELEMENTOF" : ""
  }
]

```
** Model folder : The VERSPM base model folder name 
** ScenarioInputFolder: Folder name for scenario root where scenario manager file is located. All the scenario level folders will be created here
** ScenarioManagerFile : Scenario manager database file name
** ScenarioOutputFolder: Name of the folder where all the scenarios are created

### Running VERSPM_Scnearios

Running VERSPM_Scenario has 5 major steps. After initalizing the module there are 5 modules to call

** NewScenarioSet : This module creates the required categories and scenarios json files from csv database. It also creates the folder structure and required inputs for all scenario levels  
** BuildScenarios: This module creates all possible scnearios which are defined in previous step. It copies the base VERSPM files for each scenario folder 
** RunScenarios : This module runs all scenarios 
** VERSPMResults: This model summarizes all different scenario outputs from datastor to standard format
** ViewResults:  This module create the visualization and summaries for different performance measures

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(visioneval)

ptm <- proc.time()

#Initialize model
#----------------
initializeModel(
  ModelScriptFile = "run_model.R",
  ParamDir = "defs",
  RunParamFile = "run_parameters.json",
  GeoFile = "geo.csv",
  ModelParamFile = "model_parameters.json",
  LoadDatastore = FALSE,
  DatastoreName = NULL,
  SaveDatastore = TRUE
)

#Run all demo module for all years
#---------------------------------
for(Year in getYears()) {

  runModule(
    ModuleName = "NewScenarioSet",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )

  runModule(
    ModuleName = "BuildScenarios",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "RunScenarios",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "VERSPMResults",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
  runModule(
    ModuleName = "ViewResults",
    PackageName = "VEScenario",
    RunFor = "AllYears",
    RunYear = Year
  )
}

```

### 2020-06-30 sources/modules/VETravelPerformance/R/CalculateSafetyMeasures.R
update for dplyr import warning message in CalculateSaferyMeasure

@@ -5,7 +5,7 @@
 #<doc>
 #
 ## CalculateSafetyMeasures Module
-#### June 02, 1992
+#### June 02, 2020
 #
 # This module aims to calculate safety measures for the model area. It uses the fatality and injury crash rates for different modes
 # All these rates are per miles traveled. Therefore this module simply uses calculated VMT from model to estimate the 
@@ -290,10 +290,11 @@ usethis::use_data(CalculateSafetyMeasuresSpecifications, overwrite = TRUE)
 #' @return A list containing data produced by the function consistent with the
 #' module specifications.
 #' @name CalculateSafetyMeasures
-#' @import visioneval dplyr
+#' @import visioneval 
 #' @export
 CalculateSafetyMeasures <- function(L) {
   
+  assign("%>%",getFromNamespace("%>%","magrittr"))
   
   HH_df= data.frame(L$Year$Household)
   Marea_df= data.frame(L$Year$Marea)
@@ -301,12 +302,12 @@ CalculateSafetyMeasures <- function(L) {
 # calcualte Bike and Walk PMT from Houshold table and aggregate all housholds to geth the total Marea PMTs
 # apply the injury and fatal rates to estiamte the crashes by type.  
     Crashes_HH <- HH_df %>%
-      mutate(BikeMT = BikeTrips * BikeAvgTripDist,
+      dplyr::mutate(BikeMT = BikeTrips * BikeAvgTripDist,
              WalkMT = WalkTrips * WalkAvgTripDist) %>%
-      group_by(Marea) %>%
-      summarise(BikePMT = sum(BikeMT),
+      dplyr::group_by(Marea) %>%
+      dplyr::summarise(BikePMT = sum(BikeMT),
               WalkPMT = sum(WalkMT)) %>%
-      mutate(walkfatal = Marea_df$WalkFatal[[1]] *WalkPMT / (10^8),
+      dplyr::mutate(walkfatal = Marea_df$WalkFatal[[1]] *WalkPMT / (10^8),
              walkinjury = Marea_df$WalkInjur[[1]] *WalkPMT / (10^8),
              bikefatal = Marea_df$BikeFatal[[1]] *BikePMT / (10^8),
              bikeinjury = Marea_df$BikeInjur[[1]] *BikePMT / (10^8)
[…]

### 2020-06-30 build/scripts/build-docs-framework.R
fixed missing development dependency

@@ -28,8 +28,8 @@
 if ( ! exists("ve.installer" ) ) ve.installer <- getwd()
 source(file.path(ve.installer,"scripts","get-runtime-config.R"))
 
-if ( ! suppressWarnings(requireNamespace("jsonlite",quietly=TRUE)) ) {
-  install.packages("jsonlite", lib=dev.lib, dependencies=NA, type=.Platform$pkgType )
+if ( ! suppressWarnings(requireNamespace("Rd2md",quietly=TRUE)) ) {
+  install.packages("Rd2md", lib=dev.lib, dependencies=NA, type=.Platform$pkgType )
 }
 
 message("========== BUILD FRAMEWORK DOCS ==========")

### 2020-07-01 build/boilerplate/VisionEval.R
Fixed up exporter/list functions; transparent tweak to framework

@@ -181,7 +181,11 @@ if ( install.success ) {
     tools <- character(0)
     for ( tf in tool.files ) {
       # Add error checking for tool.contents not present
-      eval(parse(text=paste0("import::here(tool.contents,.from='",tf,"')")))
+      try(
+        silent=TRUE,
+        eval(parse(text=paste0("import::here(tool.contents,.from='",tf,"')")))
+      )
+      if ( ! exists("tool.contents") ) next
       tools <- c(tools,tool.contents)
       eval(parse(text=paste0("import::into(.into='",env.loc,"',",paste(tool.contents,collapse=","),",.from='",tf,"')")))
       rm(tool.contents)

### 2020-07-01 sources/framework/visioneval/R/datastore.R

@@ -314,8 +314,7 @@ readFromTableRD <- function(Name, Table, Group, DstoreLoc = NULL, Index = NULL, 
   if (is.null(DstoreLoc)) {
     DstoreDir <- ""
   } else {
-    ParseDstoreLoc_ <- unlist(strsplit(DstoreLoc, "/"))
-    DstoreDir <- paste(ParseDstoreLoc_[-length(ParseDstoreLoc_)], collapse = "/")
+    DstoreDir <- dirname(DstoreLoc)
   }
   #Load the model state file
   if (DstoreDir == "") {

### 2020-07-01 sources/tools/exporter.R

@@ -33,15 +33,18 @@ tool.contents <- c("ve.export","ve.list")
 #     'Vehicle' from VERPAT includes incompatible numbers of rows
 #   overwrite=TRUE
 #     Blow away the outputFolder if it already exists
+#   all.tables=FALSE
+#     If TRUE, include EVERYTHING (not just output files)
 #   quiet=FALSE
 #     If FALSE, dish up some progress messages and extra debugging data
 #     If TRUE, only report errors
 
 ve.export <- function( modelStateFile = "ModelState.Rda",
                        outputFolder   = "output",
-                       includeTables  = character(0),  # default: all of them
-                       excludeTables  = c("Vehicle"),  # Vehicle doesn't work for VERPAT
+                       includeTables  = character(0),  # default: include all of them
+                       excludeTables  = character(0),  # default: don't exclude any
                        overwrite      = FALSE,         # TRUE to destroy outputFolder first if it exists
+                       all.tables     = FALSE,         # TRUE will ignore include/exclude and export everything
                        quiet          = FALSE          # TRUE to suppress all 
                        ) {
   owd <- getwd()
@@ -49,49 +52,48 @@ ve.export <- function( modelStateFile = "ModelState.Rda",
   msf <- getModelStateFile(modelStateFile,quiet=quiet)
   idx <- indexModelState(msf,quiet=quiet)
   odt <- outputData(idx,quiet=quiet)
+  outputFolder <- file.path(dirname(msf),outputFolder)
   exportData(  outputFolder  = outputFolder,
                includeTables = includeTables,
                excludeTables = excludeTables,
+               all.tables    = all.tables,
                overwrite     = overwrite,
                quiet         = quiet,
                odt
              )
-  if ( exists("ModelState_ls",envir=.GlobalEnv) ) rm(ModelState_ls,pos=.GlobalEnv) # Hack to support framework
+  if ( exists("ModelState_ls",where=.GlobalEnv,inherits=FALSE) ) rm(ModelState_ls,pos=.GlobalEnv) # Hack to support framework
   setwd(owd)
 }
 
 # Locate Model State File
[…]

### 2020-07-01 sources/tools/helpers.R

@@ -5,7 +5,7 @@
 # See VisionEval.R in the installation runtime (VE-Installer/boilerplate)
 # In particular, functions call models at locations relative to ve.root
 
-tool.contents <- c("vegui","verpat","verspm","vestate")
+tool.contents <- c("vegui","verpat","verspm","verspm_mm","vestate")
 
 # Function starts the VEGUI
 vegui <- function(ve.root=ve.runtime) {

### 2020-07-01 sources/tools/exporter.R
fixed namespace resolution problem in exporter.R

@@ -200,7 +200,7 @@ outputData <- function(Output,quiet=FALSE) {
 makeDataFrame <- function(Table, Output){
   OutputAllYr <- data.frame()
 
-  for ( year in getYears()){
+  for ( year in visioneval::getYears()){
     # Looping over each year's worth of data in the given table
     # OutputIndex is a vector of booleans true for each row in GroupTableName that
     # contains data for the sought table in the sought year

### 2020-08-25 build/scripts/build-runtime.R
Updated runtime build to fix annoyances.

@@ -59,7 +59,7 @@ message("========== BUILD RUNTIME ENVIRONMENT (scripts, models) ==========")
 cat("Runtime sources...\n")
 copy.runtime <- pkgs.db[pkgs.runtime,c("Root","Path","Package")]
 paths <- file.path(copy.runtime$Root, copy.runtime$Path, copy.runtime$Package)
-copy.paths <- dir(paths,all.files=TRUE,full.names=TRUE)
+copy.paths <- dir(paths,all.files=TRUE,full.names=TRUE,no..=TRUE)
 if ( length(copy.paths) > 0 ) {
   any.newer <- FALSE
   for ( f in seq_along(copy.paths) ) {

### 2020-08-26 sources/modules/VEPowertrainsAndFuels/inst/extdata/annualise-csv-files.Rmd

## Interpolate annual values for all PowertrainsAndFuels-new extdata
[Rick Donnelly](mailto:rick.donnelly@wsp.com) | _WSP_ | `r date()`
```{r init, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse); library(knitr)
options(width = 98)
```

<br>
Many of the PowertrainsAndFuels extdata obtained from Oregon DOT includes values in five-year intervals. The CI tests assume the data are in annual series, so we will replace the data to match this requirement. Start by finding the CSV files in question.
```{r set_path}
PATH <- "~/Models/VisionEval-Dev/sources/modules/VEPowertrainsAndFuels-new/inst/extdata"
all_csv <- Sys.glob(file.path(PATH, "*.csv"))
```

The year is coded in all of the files, but is not consistently named. Thus, we'll specify the column names that qualify as such.
```{r finger_colnames}
possible_colnames <- c("Year", "ModelYear")
```

With those preliminaries out the way let's update the data. We'll process each of the CSV files in turn, checking first whether it needs to be processed. Some files are already in annual time series, so don't touch them. For those that are not start by enumerating the full range of annual steps, which we'll merge with the existing data. Those intermediate years will have missing values in all of the remaining fields. We can use the `na_interpolate` function in the `imputeTS` package to make short order of the process. We finish by overwriting the original file with the updated contents.
```{r process_files}
for (this_csv in all_csv) {
  contents <- readr::read_csv(this_csv, col_types = cols())

  # Make sure at least one of our target field names is present, and that there is only one such field
  target_field <- intersect(possible_colnames, colnames(contents))
  if (length(target_field) != 1) {
    stop(paste("In", this_csv, "target fields found=", target_field))
  }

  # Find out what intervals are coded in the file
  coded_intervals <- contents[[target_field]]
  range <- max(coded_intervals) - min(coded_intervals)
  if (range > length(coded_intervals)) {
    print(paste("Interpolating missing years in", basename(this_csv)), quote = FALSE)
    # If we do detect one or more gaps in the annual series then create a second
    # tibble with only the years
    dummy <- tibble(!!target_field := min(coded_intervals):max(coded_intervals))

    # Now merge these years with the original data, which will create rows with missing values for all of
    # the columns. Then we'll replace all missing values in one pass.
    add_missing_years <- left_join(dummy, contents, by = target_field)
    interpolated <- imputeTS::na_interpolation(add_missing_years)

    # Replace the original data with the new annual time series
    readr::write_csv(interpolated, this_csv)
  } else {
    print(paste(basename(this_csv), "is already coded in annual steps"), quote = FALSE)
  }
}
```

We see that in one case that there were no non-missing values in one of the fields, so `na_interpolate_ had nothing to work with. Otherwise the results appear to be robust.

### 2020-08-26 sources/modules/VESimLandUse/inst/module_docs/analyze_3Ds.Rmd

---
title: "VE-State Bzone Synthesis: Synthesizing Zones with Activity Density, Diversity, and Destination Accessibility Characteristics"
author: "Brian Gregor"
output: html_document
---

# Introduction
This paper documents the process and results for creating a methodology for synthesizing Bzones in urbanized areas. To date, a method has been developed that synthesizes Bzones which reflect the activity density, diversity, and destination accessibility characteristics of urbanized area block groups. Activity density is a measure of the amount activity present in each block group. This is the *D1d* measure in the Smart Location Database (SLD), the primary data source used for developing the zone synthesis methods. The synthesized measure is the number of households and jobs per acre of land area in the block group. Diversity is a measure of the mixing of different types of land uses in each block group. The synthesized measure is the ratio of jobs to housing in the block group. This is the *D2a_JpHH* measure in the SLD. Destination accessibility is a measure of the amount of activity present in the vicinity of each block group. The measure used is the harmonic mean of the population of block groups within 5 miles and employment within 2 miles of the block group.  (Distance measurements are from block group centroid to block group centroid.) This measure is used instead of destination accessibility measures in the SLD are auto oriented (for example, jobs within 45 minute travel time) and not very useful for smaller urbanized areas. The harmonic mean of population and employment was found to be useful for distinguishing *area types*, one of the dimensions in the *place type* system that will be implemented in the Bzone synthesis process. A national mapping of place types can be seen at this [link](https://github.com/gregorbj/Placetypes_USA).

Creating SimBzones and attributing them with activity density, diversity, and destination accessibility characteristics is the first step in the VE-State Bzone synthesis process. In addition, as part of this process, numbers of households and jobs are assigned to each SimBzone. Later steps will splits jobs by employment type (retail, service, other), determine numbers of single family and multi-family housing units, identify place types, and model design and distance to transit attributes.

## Data  
EPA's [Smart Location Database](https://www.epa.gov/smartgrowth/smart-location-mapping) is the principal source of data used for developing the method for the synthesizing Bzones. The SLD includes a large number of land use and transportation measures at the Census block group level for the year 2010. The destination accessibility measure described above was added to the SLD dataset using the locations of the block group centroids.

The SLD includes data on urbanized and non-urbanized areas. Only the data for urbanized areas having populations of 50,000 or more were used to develop the zone synthesis methodology. A separate method will be developed to synthesize Bzones representing smaller urban areas (i.e. towns) and rural areas. In addition, some data cleaning was done by removing block group records that had no activity or land area. This dataset includes data on 139,564 block groups in 431 urbanized areas. The separate state components of urbanized areas that cross state boundaries were identified because the characteristics of block groups in each state may differ substantially and it is common practice for state models to only include the portion of the urbanized area located in the state.

## Summary of Methodology
The following list summarizes the Bzone synthesis methods developed to date:

1) The main inputs that drive the process are numbers of households by Azone, employment by Azone, the geographic extent of the urbanized area, and the proportional distribution of Azone households and employment by location type (rural, town, metropolitan). (Note that metropolitan in this context refers to the urbanized portion of the metropolitan area.) From these data, the amounts of jobs and households to be allocated to urbanized area SimBzones and the average activity density of the urbanized area can be computed.

2) SimBzones are created to have approximately equal amounts of activity: approximately the median value for urbanized area block groups in the SLD (750/ac).

3) The distributions of activity density, diversity, and destination accessibility are simplified by splitting them into levels: 20 levels for density, 5 levels for diversity, and 20 levels for destination accessibility.

4) A profile is developed for each urbanized area in the SLD dataset. In addition, profiles are developed for 6 urbanized area size categories to use when data for a specific metropolitan area is not available. The profiles include:
  - The proportion of total urbanized area activity at each density level;
  - The average block group activity density at each density level;
  - The proportion of activity by diversity level for each density level;
  - Distributions of the employment proportions of activity at each diversity level;
  - The proportion of activity by destination accessibility level for each density level;
  - The average block group destination accessibility at each destination accessibility level.

5) A profile for the entire set of urbanized areas is also developed to fill to fill missing values in individual urbanized area profiles when necessary to project future (or past) conditions.

6) The first step after the total amount of activity and land area has been assigned to an urbanized area is to create SimBzones to contain the projected urbanized area activity. The number of SimBzones for an urbanized area is the whole number result (by rounding) of dividing the total activity by 750. The remainder (positive or negative) is randomly allocated to the SimBzones. This allocation is done for each Azone portion of the urbanized area to assure that the values for SimBzones in each Azone are consistent with Azone household and jobs inputs.

7) Next, each SimBzone is assigned an activity density level. First a proportional allocation of activity to activity density levels is calculated such that the distribution averages to the average density for the urbanized area. This is accomplished by starting with the urbanized area profile of the proportional distribution by activity density level and then adjusting those proportions in small increments until the average density for the urbanized area is achieved. The mean density profile for the urbanized area is used in this calculation. Missing values in the mean density profile are filled in with values from the profile for urbanized areas as a whole. The proportional allocation of activity by density level is used as a sampling distribution to assign each SimBzone a density level. This process is carried out separately for each Azone portion of the urbanized area. A final density adjustment is made so that the average activity density in each Azone urbanized area matches input assumptions.

8) After each SimBzone has an assigned activity level, it is assigned a diversity level. This is accomplished by using the urbanized area profile of proportional allocations of activity by diversity level for each activity density level for sampling distributions. Missing values are filled in from the overall urbanized area profile. Once the diversity level has been chosen a value for the employment proportion of activity is picked for each SimBzone from the employment proportions distributions. The employment proportions are used to make an initial calculation of employment in each SimBzone. That employment is adjusted as necessary to match the urbanized area employment total in each Azone. The number of SimBzone households is then calculated.

9) Bzones are assigned destination accessibility levels in a similar manner. The urbanized area profile of proportional allocations of activity by destination accessibility level are used as sampling distributions to determine destination accessibility levels. Missing values are filled in from the overall urbanized area profile. The specific destination accessibility values are the mean values computed for the levels for the urbanized area. If destination accessibility proportions are missing for one or more density levels, they are filled in from the profile for urbanized areas as a whole.

## Issues
The current GreenSTEP model allows users to specify a proportion of households located in urban mixed-use neighborhoods. Should users of VE-State be able to similarly specify the proportion of activity to be allocated to SimBzones that are in the mixed-use category? If so, the way that users would specify this and the method for adjusting SimBzone densities to match user inputs need to be developed. In addition, some guards should be established to prevent users from specifying unreasonable mix proportions.

The current GreenSTEP model does not allow users to adjust the density distribution. I don't think VE-State should allow this either. Nor should it allow users to adjust the destination accessibility distribution.

## Next Steps
A number of things need to be done to complete the synthetic zone generation process:
1) Develop method for splitting employment into categories used by multi-modal model: retail, service, other.
2) Develop method for determining place types.
3) Tabulate design and distance to transit attributes used by the multi-modal model by place type for each urbanized area. Develop a method for users to input targets for these attributes by place type.
4) Create the method for determining SimBzone housing type (single-family, multifamily) proportions.
5) Modify the method in the VEPredictHousing module to work with SimBzones. The VEPredictHousing method requires inputs on the mix of households by income quartile for each Bzone. This enables the distributions of households by income to reflect observed distributions and enables users to model the effects of gentrification. Should something similar be incorporated in VE-State?
6) Develop approaches for user inputs for parking policies, travel demand management policies, and car service availability.

```{r echo=FALSE}
#======
#SET UP
#======
#Load libraries
library(visioneval)
library(plot3D)

#Load 5Ds data frame
#D_df <- readRDS("5Ds_df.rds")
D_df <- VESimLandUseData::SimLandUseData_df

#==================================================
#DEFINE FUNCTIONS USED IN MULTIPLE PLACES IN SCRIPT
#==================================================

#Define function to add a smoothed line to binned values
#-------------------------------------------------------
addSmoothLine <- function(X_, Y_, ...) {
  X_ <- X_[!is.na(Y_)]
  Y_ <- Y_[!is.na(Y_)]
  XY_SS <-smooth.spline(Y_ ~ X_)
  XPred_ <- seq(X_[1], X_[length(X_)], length = length(X_) * 10)
  lines(XPred_, predict(XY_SS, XPred_)$y, ...)
}

#Define a function to plot distributions for an urbanized area and compare with
#the distribution for all urbanized areas
#------------------------------------------------------------------------------
plotDist <- function(UZA_NAME, VarName, LogTransform = TRUE, ...) {
  UzaDen <- round(ActDen_Ua[UZA_NAME], 1)
  Title <- paste0(UZA_NAME, "\n", "(Ave ", UzaDen, " HH & Jobs per acre)")
  if (LogTransform) {
    Dat_ <- log(D_df[,VarName])
    Xlim_ <- range(Dat_[!is.infinite(Dat_)])
    Var_ <- Dat_[D_df$UZA_NAME == UZA_NAME]
    Var_ <- Var_[!is.infinite(Var_)]
  } else {
    Dat_ <- D_df[,VarName]
    Xlim_ <- range(Dat_)
    Var_ <- Dat_[D_df$UZA_NAME == UZA_NAME]
  }
  plot(density(Var_), xlim = Xlim_, main = Title, ...)
  lines(density(Dat_), lty = 2)
}

#Function to calculate average density using group data
#------------------------------------------------------
calcAveDensity <- function(AveDensity_, PropActivity_) {
  sum(1 / sum(PropActivity_ / AveDensity_, na.rm = TRUE), na.rm = TRUE)
}

#===============================
#CALCULATE URBANIZED AREA VALUES
#===============================

#Limit dataset to areas with populations >= 50,000
#-------------------------------------------------
#Calculate area population and activity
TotPop_Ua <- tapply(D_df$TOTPOP10, D_df$UA_NAME, sum)
#Limit dataset to urbanized areas having populations >= 50,000
AllD_df <- D_df
D_df <- D_df[D_df$UA_NAME %in% names(TotPop_Ua)[TotPop_Ua >= 50000],]

#Calculate activity size category
#--------------------------------
TotAct_Ua <- tapply(D_df$TOTACT, D_df$UA_NAME, sum)
SzBrk_ <- c(0, 5e4, 1e5, 5e5, 1e6, 5e6, max(TotAct_Ua))
Sz <- c("small", "medium-small", "medium", "medium-large", "large", "very-large")
SzGrp_Ua <- cut(TotAct_Ua, SzBrk_, labels = Sz, include.lowest = TRUE)
names(SzGrp_Ua) <- names(TotAct_Ua)
D_df$UZA_SIZE <- SzGrp_Ua[match(D_df$UA_NAME, names(SzGrp_Ua))]
rm(TotAct_Ua, SzBrk_, SzGrp_Ua)

#Calculate overall urbanized area activity density
#-------------------------------------------------
Tmp_ls <- split(D_df[,c("TOTACT", "AC_LAND", "TOTPOP10")], D_df$UZA_NAME)
ActDen_Ua <- unlist(lapply(Tmp_ls, function(x){
  sum(x$TOTACT) / sum(x$AC_LAND)
}))
D_df$UZA_ACTDEN <- ActDen_Ua[D_df$UZA_NAME]
rm(Tmp_ls)

#Identify selected urbanized areas to use in comparison plots
#------------------------------------------------------------
UzaToPlot_ <- c(
  "Atlanta, GA",
  "Jacksonville, FL",
  "Cincinnati, OH",
  "Dallas-Fort Worth-Arlington, TX",
  "Baltimore, MD",
  "Denver-Aurora, CO",
  "Portland, OR",
  "San Francisco-Oakland, CA",
  "New York-Newark, NY"
)
```

# Evaluate the Distributions of 3D Variables for Urbanized Areas
The block group values for activity density, diversity, and destination accessibility are approximately log normally distributed as shown in the following graphs.

```{r echo=FALSE}
#==============================================================
#EVALUATE THE DISTRIBUTIONS OF 3D VARIABLES FOR URBANIZED AREAS
#==============================================================

#Plot overall urbanized area log-transformed distributions
#---------------------------------------------------------
#Plot log distributions - these look fairly normal
#D2A_JPHH has some 0 and infinite values so just plot non-zero values for that
InitPar_ls <- par(mfrow = c(2,2), oma = c(0,0,3,0))
with(D_df, {
  plot(density(log(D1D)),
       xlab = "Log of Jobs & HHs Per Acre",
       main = "Density of Jobs & HHs\n(D1D)")
  plot(density(log(D2A_JPHH[D2A_JPHH != 0 & !is.infinite(D2A_JPHH)])), 
       xlab = "Log of Jobs / HHs",
       main = "Diversity: Jobs-Housing Ratio\n(D2A_JPHH)")
  plot(density(log(D5)), 
       xlab = "Log of Accessibility\n(see text)",
       main = "Destination Accessibility\n(D5)")
  mtext("Distributions of Census Block Group 3D Measures for All Urbanized Areas",
        side = 3, outer = TRUE)
})
par(InitPar_ls)
rm(InitPar_ls)
```

The following figure shows the density distributions for 9 example urbanized areas arranged from least to most densely developed. The solid lines show the density distributions for the urbanized areas. The dashed lines show the average density distribution for all urbanized areas. Note that as urbanized area density increases the density distribution shifts to the right.

```{r echo=FALSE}
#Plot activity density (D1D) distribution for selected urbanized areas
#---------------------------------------------------------------------
InitPar_ls <- par(mfrow = c(3,3), oma = c(0, 0, 2.2, 0))
for (Ua in UzaToPlot_) {
  plotDist(Ua, "D1D", ylim = c(0, 0.8), xlab = "Natural log of D1D")
}
mtext("Block Group Activity Density Distribution (D1D) for Selected Urbanized Areas\nCompared to Distribution for All Urbanized Areas (dashed line)", outer = TRUE, line = -0.5)
par(InitPar_ls)
```

The following figure shows the diversity distributions for the example urbanized areas. Notice that there is little variation in the distributions among the urbanized areas. All are very similar to the overall urbanized area distribution.

```{r echo=FALSE}
#Plot diversity (D2A_JPHH) distribution for selected urbanized areas
#-------------------------------------------------------------------
InitPar_ls <- par(mfrow = c(3,3), oma = c(0, 0, 2.2, 0))
for (Ua in UzaToPlot_[order(ActDen_Ua[UzaToPlot_])]) {
  plotDist(Ua, "D2A_JPHH", ylim = c(0, 0.35), xlab = "Natural log of D2A_JPHH")
}
mtext("Block Group Activity Diversity Distribution (D2A_JPHH) for Selected Urbanized Areas\nCompared to Distribution for All Urbanized Areas (dashed line)", outer = TRUE, line = -0.5)
par(InitPar_ls)
```

The following figure shows the destination accessibility distributions for the example urbanized areas. As with the density distributions, these shift to the right as the urbanized area density increases.

```{r echo=FALSE}
#Plot destination accessibility (D5) distribution for selected urbanized areas
#-----------------------------------------------------------------------------
InitPar_ls <- par(mfrow = c(3,3), oma = c(0, 0, 2.2, 0))
for (Ua in UzaToPlot_[order(ActDen_Ua[UzaToPlot_])]) {
  plotDist(Ua, "D5", ylim = c(0, 0.5), xlab = "Natural log of D5")
}
mtext("Block Group Destination Accessibility Distribution (D5) for Selected Urbanized Areas\nCompared to Distribution for All Urbanized Areas (dashed line)", outer = TRUE, line = -0.5)
par(InitPar_ls)
#Clean up
rm(Ua, InitPar_ls)
```

The following 3 figures present relationships between these 3Ds measures. Diversity is weakly correlated with density and destination accessibility. Density and destination accessibility are more strongly correlated.

```{r echo=FALSE}
#Prepare data to compare correlations between 3Ds
#------------------------------------------------
T_df <- D_df[,c("D1D", "D2A_JPHH", "D5")]
T_df$LogD1D <- log(T_df$D1D)
T_df$LogD2A_JPHH <- log(T_df$D2A_JPHH)
T_df$LogD5 <- log(T_df$D5)
T_df <- T_df[!is.infinite(T_df$LogD2A_JPHH),]
D1D_ <- T_df$LogD1D
D2A_JPHH_ <- T_df$LogD2A_JPHH
D5_ <- T_df$LogD5

#Present the relationship of diversity with density
#--------------------------------------------------
plot(D1D_, D2A_JPHH_, pch = ".", xlab = "Log D1D (density)", 
     ylab = "Log D2A_JPHH (diversity)",
     main = "Relationship of Diversity with Density")
Cor <- round(cor(D2A_JPHH_, D1D_), 2)
Text <- paste("Correlation Coefficient =", Cor)
text(-6, 9, labels = Text, pos = 4)
rm(Cor, Text)

#Present the relationship of destination accessibility with density
#------------------------------------------------------------------
plot(D1D_, D5_, pch = ".", xlab = "Log D1D (density)", 
     ylab = "Log D5 (destination accessibility)",
     main = "Relationship of Destination Accessibility with Density")
Cor <- round(cor(D5_, D1D_), 2)
Text <- paste("Correlation Coefficient =", Cor)
text(-6, 13, labels = Text, pos = 4)
rm(Cor, Text)

#Present the relationship of diversity with destination accessibility
#--------------------------------------------------------------------
plot(D5_, D2A_JPHH_, pch = ".", xlab = "Log D5 (destination accessibility)", 
     ylab = "Log D2A_JPHH (diversity)",
     main = "Relationship of Diversity with Destination Accessibility")
Cor <- round(cor(D2A_JPHH_, D5_), 2)
Text <- paste("Correlation Coefficient =", Cor)
text(1, 9, labels = Text, pos = 4)
rm(Cor, Text)
#Clean up
rm(T_df, D1D_, D2A_JPHH_, D5_)
```

# Define Density Levels
Activity density levels are defined by dividing the log transformation of the distribution into 20 equal intervals and then untransforming the result. The proportion of the urbanized area activity in each bin is tabulated for all urbanized areas and for each individual urbanized area. It is also tabulated by urbanized area size group where size groups are defined by the total of households and jobs in the urbanized area as follows:

| Category     | Value Range |
|:-------------|:------------|
|small         |0 - 5e4      |
|medium-small  |5e4 - 1e5    |
|medium        |1e5 - 5e5    | 
|medium-large  |5e5 - 1e6    |
|large         |1e6 - 5e6    |
|very-large    |5e6+         |

```{r echo=FALSE}
#================================================
#DEFINE AND ANALYZE ACTIVITY DENSITY (D1D) LEVELS
#================================================
#
#Define activity density levels
#------------------------------
#Determine breaks by dividing log of D1D into 20 equal intervals
D1DGrpBrk_ <- local({
  LogD1D_ <- log(D_df$D1D)
  Interval <- diff(range(LogD1D_)) / 20
  LogBreaks_ <- min(LogD1D_) + 0:20 * Interval
  exp(LogBreaks_)
})
#Make sure the minimum and maximum values bound the breaks
D1DGrpBrk_[1] <- min(D_df$D1D)
D1DGrpBrk_[length(D1DGrpBrk_)] <- max(D_df$D1D)
#Identify the density group for each block group
D1DGrp_ <- cut(D_df$D1D, breaks = D1DGrpBrk_, include.lowest = TRUE)
#Calculate the average density for each quantile
D1DGrpAve_ <- unlist(lapply(split(D_df, D1DGrp_), function(x) {
  sum(x$TOTACT) / sum(x$AC_LAND)
}))
#Calculate total activity in each quantile
D1DGrpTotAct_ <- tapply(D_df$TOTACT, D1DGrp_, sum)
#Check average urbanized area density calculated from groups
AveD1DCompare_ <- c(
  Ave = with(D_df, sum(TOTACT) / sum(AC_LAND)),
  BinAve = sum(D1DGrpTotAct_) / sum(D1DGrpTotAct_ / D1DGrpAve_)
)
#Add D1DGrp to D_df
D_df$D1DGrp <- D1DGrp_
#Calculate the proportion of total activity by group
D1DGrpPropAct_ <- D1DGrpTotAct_ / sum(D1DGrpTotAct_)
#Create a list to save the urbanized area D1D group data
D1DGrp_ls <- list(
  AveDensity = D1DGrpAve_,
  PropActivity = D1DGrpPropAct_
)
```

The following figure shows the overall proportional distribution of urbanized area activity by density group.

```{r echo=FALSE}
#Plot activity proportions and average density by density group
#--------------------------------------------------------------
#Plot the proportion of total activity by group
plot(1:20, D1DGrpPropAct_,
     xlab = "Activity Density (D1D) Group",
     ylab = "Proportion of Activity (HHs & Jobs)",
     main = "Proportions of Activity by Activity Density Group\nBlock Groups in All Urbanized Areas")
addSmoothLine(1:20, D1DGrpPropAct_, lty = 2)
```

In addition, for each of these groupings the average density by activity level is computed. This information is used to calculate overall density and to make density adjustments so that the density of urbanized area SimBzones 'sums up' to the overall density of the urbanized area. The following figure shows the average density by density level for all urbanized areas. Average density by density level is also computed for each urbanized area identified in the SLD and for each of the urbanized area size categories.

```{r echo=FALSE}
#Plot the group D1D averages
plot(1:20, D1DGrpAve_,
     xlab = "Activity Density (D1D) Group",
     ylab = "Average Density (HHs + Jobs per Acre)",
     main = "Average Activity Density by Activity Density Group\nBlocks Groups in All Urbanized Areas")
addSmoothLine(1:20, D1DGrpAve_, lty = 2)
#Clean up
rm(D1DGrpBrk_, D1DGrp_, D1DGrpTotAct_, D1DGrpAve_, D1DGrpPropAct_)

#Calculate activity density bin information for urbanized areas
#--------------------------------------------------------------
#Split data by urbanized area
D_Ua_df <- split(D_df, D_df$UZA_NAME)
#Calculate average density and proportion of activity by density group
D1DGrp_Ua_ls <- lapply(D_Ua_df, function(x) {
  GrpSplit_ls <- split(x, x$D1DGrp)
  list(
    AveDensity =
      unlist(lapply(GrpSplit_ls, function(x) sum(x$TOTACT) / sum(x$AC_LAND))),
    PropActivity =
      tapply(x$TOTACT, x$D1DGrp, sum) / sum(x$TOTACT)
  )
})
rm(D_Ua_df)
#Split data by urbanized area size category
D_Sz_df <- split(D_df, D_df$UZA_SIZE)
#Calculate average density and proportion of activity by density group
D1DGrp_Sz_ls <- lapply(D_Sz_df, function(x) {
  GrpSplit_ls <- split(x, x$D1DGrp)
  list(
    AveDensity =
      unlist(lapply(GrpSplit_ls, function(x) sum(x$TOTACT) / sum(x$AC_LAND))),
    PropActivity =
      tapply(x$TOTACT, x$D1DGrp, sum) / sum(x$TOTACT)
  )
})
#Add generic urbanized areas by size to urbanized area list
for (sz in Sz) {
  D1DGrp_Ua_ls[[sz]] <- D1DGrp_Sz_ls[[sz]]
}
rm(D_Sz_df, D1DGrp_Sz_ls)

#Check that density average calculated from bins are same as overall average
#for every urbanized area
#---------------------------------------------------------------------------
#Calculate average density for all urbanized areas
AveDen_Ua <- unlist(lapply(D1DGrp_Ua_ls, function(x) {
  calcAveDensity(x$AveDensity, x$PropActivity)
}))
#Values are the same
#summary(AveDen_Ua[!(names(AveDen_Ua) %in% Sz)] / ActDen_Ua)
rm(AveDen_Ua)
```

# Predicting Changes in Zone Density
While the synthetic zone generation process uses observed urban area distributions, it is necessary to incorporate a process for adjusting the distributions so that they are consistent with model input assumptions for overall urbanized area activity density. If urbanized area density decreases, the distribution needs to shift to the left, and if overall density increases, the distribution needs to shift to the right. The method shifts the distribution to the left or right in small increments by weighted averaging of the activity proportions value at each level with the value of the bin to the right (if the distribution shifts to the left) and vise verse if the distribution shifts to the right. With each increment, the average urbanized area density is calculated using the revised activity proportions and the average density per bin. The iteration stops when the average density is within 1% of the target. At that point, final adjustments are made to the bin average densities so that the target density is matched exactly.

Regarding the average density per bin, the values start with those calculated for the urbanized area. Where values are missing because activity is not present in some density bins, values are substituted from the overall urbanized area averages.

The proportional distribution of activity by density group is used in the Bzone synthesis process as a sampling distribution to determine the density group of each SimBzone in the urbanized area of an Azone. The average density per group is used to determine the average activity density of each SimBzone.

```{r echo=FALSE}
#=========================================================================
#DEVELOP A MODEL FOR PREDICTING DENSITY BY BIN FROM URBANIZED AREA DENSITY
#=========================================================================
#The model adjusts the distribution of activity by density group in response to
#a change in the overall density of the urbanized area. This is done by
#iteratively taking a weighted moving average of the activity proportion in each
#bin with the activity proportion in the bin to the left (if overall density
#increases) or to the right (if overall density decreases).

#Define function to adjust density distribution
#----------------------------------------------
#' Adjust urbanized area density distribution to match average density target
#' 
#' @param DenDist_ A numeric vector of the proportions of urbanized activity by 
#' activity density bin for the urbanized area.
#' @param AveDensity_ A numeric vector of the average activity density by
#' activity density bin for the urbanized area.
#' @param Target A number specifying the average activity density for the
#' urbanized area.
#' @param UzaAveDensity_ A numeric vector of the average activity density by
#' activity density bin for all urbanized areas.
#' @return A data frame having 20 rows and 2 columns: ActProp, the proportion of
#' urbanized activity by activity density bin; and AveDensity, the average
#' density by activity density bin.
adjDenDist <- function(DenDist_, AveDensity_, Target, UzaAveDensity_) {
  #Fill in NA values in DenDist_ and AveDensity_
  DenDist_[is.na(DenDist_)] <- 0
  AveDensity_[is.na(AveDensity_)] <- UzaAveDensity_[is.na(AveDensity_)]
  #Define function to calculate overall average density
  calcAveDensity <- function(AveDensity_, PropActivity_) {
    sum(1 / sum(PropActivity_ / AveDensity_, na.rm = TRUE), na.rm = TRUE)
  }
  #Initial calculation of average density
  AveDensity <- calcAveDensity(AveDensity_, DenDist_)
  #Define a function to make an incremental adjustment in density distribution
  makeAdj <- function(DenDist_, AdjProp){
    if (Target > AveDensity) {
      ShiftDist_ <- c(0, DenDist_[-length(DenDist_)])
    } else {
      ShiftDist_ <- c(DenDist_[-1], 0)
    }
    AdjDenDist_ <- (1 - AdjProp) * DenDist_ + AdjProp * ShiftDist_
    AdjDenDist_ / sum(AdjDenDist_)    
  }
  #Make incremental adjustments until target is approximately achieved
  while (abs(1 - (Target / AveDensity)) > 0.01) {
    TargetDiff <- abs(Target - AveDensity)
    if (TargetDiff > 0.1) {
      DenDist_ <- makeAdj(DenDist_, 0.01)
    } else {
      DenDist_ <- makeAdj(DenDist_, 0.001)
    }
    AveDensity <- calcAveDensity(AveDensity_, DenDist_)
  }
  #Make final adjustments to density bin averages
  AveDensity_ <- AveDensity_ * Target / AveDensity
  #Return data frame of density distribution and average density
  data.frame(
    ActProp = DenDist_,
    AveDensity = AveDensity_
  )
}
```

The following figure shows how the density distribution adjustment works when the density of the urbanized area decreases. In this test, the overall activity density of the Portland (OR) urbanized area is decreased to be equal to that of the Atlanta urbanized area (`r round(ActDen_Ua["Atlanta, GA"], 2)` households and jobs per acre).

```{r echo=FALSE}
#Test function with with Portland decreasing to density of Atlanta
#-----------------------------------------------------------------
AdjDenDist_df <- adjDenDist(
  D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity, 
  D1DGrp_Ua_ls[["Portland, OR"]]$AveDensity, 
  ActDen_Ua["Atlanta, GA"], 
  D1DGrp_ls$AveDensity)
```

The adjustment method produces a density distribution for the hypothetical Portland that has an an equivalent overall activity density (`r round(calcAveDensity(AdjDenDist_df$AveDensity, AdjDenDist_df$ActProp), 2)`). The distribution of activity by density group is similar to that of Atlanta. It is not the same, however, because it reflects the Portland area starting conditions as well.

```{r echo=FALSE}
#Plot comparison of density distributions
plot(D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity,
     xlab = "Activity Density (D1D) Group",
     ylab = "Proportion of Activity (HHs & Jobs)",
     main = "Test of Adjusting Urbanized Area Average Activity Density (D1D) Downward\nPortland (OR) Activity Density -> Atlanta Activity Density")
addSmoothLine(1:20, D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity)
points(1:20, AdjDenDist_df$ActProp, col="red")
addSmoothLine(1:20, AdjDenDist_df$ActProp, col="red")
points(1:20, D1DGrp_Ua_ls[["Atlanta, GA"]]$PropActivity, col = "green")
addSmoothLine(1:20, D1DGrp_Ua_ls[["Atlanta, GA"]]$PropActivity, col = "green")
legend("topleft", lty = 1, col = c("black", "red", "green"), bty = "n",
       legend = c("Portland Density", "Portland at Atlanta Density", "Atlanta Density"))
#Clean up
rm(AdjDenDist_df, calcAveDensity)
```

The following figure shows how the density distribution adjustment works when the density of the urbanized area increases. In this test, the overall activity density of the Portland (OR) urbanized area is increased to be to be equal to that of the New York (NY) urbanized area (`r round(ActDen_Ua["New York--Newark, NY"], 2)` households and jobs per acre).

```{r echo=FALSE}
#Test function with Portland increasing to density of New York
#-------------------------------------------------------------
calcAveDensity <- function(AveDensity_, PropActivity_) {
  sum(1 / sum(PropActivity_ / AveDensity_, na.rm = TRUE), na.rm = TRUE)
}
AdjDenDist_df <- adjDenDist(
  D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity, 
  D1DGrp_Ua_ls[["Portland, OR"]]$AveDensity, 
  ActDen_Ua["New York-Newark, NY"], 
  D1DGrp_ls$AveDensity)
```

The adjustment method produces a density distribution for the hypothetical Portland that has an an equivalent overall activity density (`r round(calcAveDensity(AdjDenDist_df$AveDensity, AdjDenDist_df$ActProp), 2)`). The distribution of activity by density group differs from that of the New York, especially at the top end of the distribution. This is reasonable because the density distribution of the New York area is a result of the history of that area. Even if the Portland area reached the New York area density at some point in the future, its history would be different and so its density distribution is unlikely to be the same.

```{r echo=FALSE}
#Plot comparison of density distributions
plot(D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity,
     xlab = "Activity Density (D1D) Group",
     ylab = "Proportion of Activity (HHs & Jobs)",
     main = "Test of Adjusting Urbanized Area Average Activity Density (D1D) Upward\nPortland (OR) Activity Density -> New York (NY) Activity Density")
addSmoothLine(1:20, D1DGrp_Ua_ls[["Portland, OR"]]$PropActivity)
points(1:20, AdjDenDist_df$ActProp, col="red")
addSmoothLine(1:20, AdjDenDist_df$ActProp, col="red")
points(1:20, D1DGrp_Ua_ls[["New York-Newark, NY"]]$PropActivity, col = "green")
addSmoothLine(1:20, D1DGrp_Ua_ls[["New York-Newark, NY"]]$PropActivity, col = "green")
legend("topleft", lty = 1, col = c("black", "red", "green"), bty = "n",
       legend = c("Portland Density", "Portland at New York Density", "New York Density"))
rm(AdjDenDist_df)
```

# Modeling Activity Diversity as a Function of Activity Density
Although the correlation between block group activity diversity and activity density is weak, there are some relationships at an aggregate level that can be caputured and used in the model. Census block groups that have a high level of activity density tend to also have an activity mix that is dominated by employment. Census block groups that have a low level of activity density tend to have an activity mix that is dominated by households. A more balanced mix of jobs and households is found more often at mid-range densities.

Because of the weak relationship between density and diversity and reflecting earlier work on [place types](https://github.com/gregorbj/Placetypes_USA), 5 diversity categories are defined as follows based on the ratio of jobs to households:

| Category     | Value Range |
|:-------------|:------------|
|primarily-hh  |0 - 0.25     |
|largely-hh    |0.25 - 0.5   |
|mixed         |0.5 - 2      | 
|largely-job   |2 - 4        |
|primarily-job |4 - infinity |

```{r echo=FALSE}
#==============================================================
#ANALYSE RELATIONSHIP OF DIVERSITY (D2A_JPHH) AND DENSITY (D1D)
#==============================================================
#Although the correlation between block group activity diversity (D2A_JPHH) and
#density (D1D) is weak, there are relationships at an aggregate level that are
#useful to capture. For example, urbanized areas characterized by very high
#density levels tend to have high ratios to jobs to households.

#Identify diversity group levels
#-------------------------------
#5 diversity group levels are defined from a lowest level where households
#dominate to the highest level where jobs dominate. The middle level has the
#most balance where the ratio of jobs to households and the inverse ratio does
#not exceed 2:1.
D2GrpBrk_ <- c(
  min(D_df$D2A_JPHH), 0.25, 0.5, 2, 4, max(D_df$D2A_JPHH)
)
D2 <- c("primarily-hh", "largely-hh", "mixed", "largely-job", "primarily-job")
D2Grp_ <- cut(D_df$D2A_JPHH, D2GrpBrk_, labels = D2, include.lowest = TRUE)
D_df$D2Grp <- D2Grp_
```

The following figure illustrates how the activity is distributed among diversity groups at each activity density level. The 5 diversity groups are shown along the y-axis while the 20 density groups are shown along the x-axis from lowest to highest density. The colors show the proportion of activity in each diversity group for each density group (the values in each column add to 1). It can be seen that except for the very lowest density levels, the employment proportion of activity increases as density increases.

```{r echo=FALSE}
#Calculate proportions of activity by diversity group for each density group
#All urbanized areas
#---------------------------------------------------------------------------
TotAct_D1D2 <- with(D_df, tapply(TOTACT, list(D1DGrp, D2Grp), sum))
TotAct_D1D2[is.na(TotAct_D1D2)] <- 0
D2ActProp_D1D2 <- sweep(TotAct_D1D2, 1, rowSums(TotAct_D1D2), "/")
DispPal_ <- colorRampPalette(c("black", "yellow"))(10)
image2D(D2ActProp_D1D2, 
        x = 1:20, y = 1:5,
        xlab = "Density Group",
        ylab = "Diversity Group",
        zlim = c(0,1),
        col = DispPal_, NAcol = "black",
        axes = FALSE,
        main = "Diversity Group Activity Proportions by Density Group\nAll Urbanized Areas")
axis(1)
axis(2, at = 1:5, labels = D2)
```

Distributions of activity by diversity group for each density group are also tabulated for each named urbanized area in the SLD. These distributions are used as sampling distributions to determine the diversity group of each SimBzone once the density group of the SimBzone has been determined. Where an urbanized area does not have diversity proportions for one or more density groups, the values for the overall tabulation for urbanized areas are used.

The following figure displays the tabulations for the same 9 urbanized areas compared previously. The images illustrate that there is a lot of variation in the diversity-density patterns of urbanized area. It also illustrates some general trends. Lower density urbanized areas tend to have greater segmentation of diversity by density group, and a narrower range of densities supporting mixed development.

```{r echo=FALSE}
#Urbanized area proportions of activity by diversity group at each density group
#-------------------------------------------------------------------------------
#Calculate relationship for each urbanized area
Tmp_Ua_df <- split(D_df, D_df$UZA_NAME)
D2ActProp_Ua_D1D2 <- lapply(Tmp_Ua_df, function(x) {
  TotAct_D1D2 <- tapply(x$TOTACT, list(x$D1DGrp, x$D2Grp), sum)
  D2ActProp_D1D2 <- 
    sweep(TotAct_D1D2, 1, rowSums(TotAct_D1D2, na.rm = TRUE), "/")
  D2ActProp_D1D2[is.nan(D2ActProp_D1D2)] <- 0
  D2ActProp_D1D2
})
ActProp_Ua_D2 <- lapply(Tmp_Ua_df, function(x) {
  TotAct_D2 <- tapply(x$TOTACT, x$D2Grp, sum)
  TotAct_D2[is.na(TotAct_D2)] <- 0
  TotAct_D2 / sum(TotAct_D2)
})
rm(Tmp_Ua_df)
#Calculate relationship for each urbanized area size group
Tmp_Sz_df <- split(D_df, D_df$UZA_SIZE)
D2ActProp_Sz_D1D2 <- lapply(Tmp_Sz_df, function(x) {
  TotAct_D1D2 <- tapply(x$TOTACT, list(x$D1DGrp, x$D2Grp), sum)
  D2ActProp_D1D2 <- 
    sweep(TotAct_D1D2, 1, rowSums(TotAct_D1D2, na.rm = TRUE), "/")
  D2ActProp_D1D2[is.nan(D2ActProp_D1D2)] <- 0
  D2ActProp_D1D2
})
ActProp_Sz_D2 <- lapply(Tmp_Sz_df, function(x) {
  TotAct_D2 <- tapply(x$TOTACT, x$D2Grp, sum)
  TotAct_D2[is.na(TotAct_D2)] <- 0
  TotAct_D2 / sum(TotAct_D2)
})
rm(Tmp_Sz_df)
#Add urbanized area size group tabulations to urbanized area list
for (sz in Sz) {
  D2ActProp_Ua_D1D2[[sz]] <- D2ActProp_Sz_D1D2[[sz]]
  ActProp_Ua_D2[[sz]] <- ActProp_Sz_D2[[sz]]
}
rm(D2ActProp_Sz_D1D2, ActProp_Sz_D2)
#Compare diversity patterns for selected metropolitan areas
DispPal_ <- colorRampPalette(c("black", "yellow"))(10)
InitPar_ls <- par(mfrow = c(3,3), mar = c(4,4,3,3), oma = c(0, 0, 2.2, 0))
for (Ua in UzaToPlot_) {
  ImageDat_D1D2 <- D2ActProp_Ua_D1D2[[Ua]]
  image2D(ImageDat_D1D2, 
          x = 1:20, y = 1:5,
          zlim = c(0,1),
          xlab = "Density Group",
          ylab = "Diversity Group",
          col = DispPal_, NAcol = "black",
          main = Ua, axes = FALSE)
  axis(1)
  axis(2, at = 1:5, 
       labels = c("Prn\nHH", "Lrg\nHH", "Mix", "Lrg\nJOB", "Prn\nJOB"))
}
mtext("Activity Proportions by Diversity Group for Each D1D Group\nFor Selected Urbanized Areas", 
      outer = TRUE, line = -0.5)
par(InitPar_ls)
rm(Ua, InitPar_ls)
```

The following figure compares the diversity proportions by density level for the 6 urbanized area size groups. Except for the highest densities and largest size groups, there are no clear distinctions in these patterns. The highest density values differ because high density levels are not found in smaller urbanized areas. The more dispersed pattern in the very large urbanized area size group reflects the high densities and extensive areas found in the few urbanized areas in this group.

```{r echo=FALSE}
#Compare diversity patterns by urbanized area size group
DispPal_ <- colorRampPalette(c("black", "yellow"))(10)
InitPar_ls <- par(mfrow = c(2,3), mar = c(3,3,3,3), oma = c(0, 0, 2.2, 0))
for (sz in Sz) {
  ImageDat_D1D2 <- D2ActProp_Ua_D1D2[[sz]]
  image2D(ImageDat_D1D2, 
          x = 1:20, y = 1:5,
          zlim = c(0,1),
          xlab = "Density Group",
          ylab = "Diversity Group",
          col = DispPal_, NAcol = "black",
          main = sz, axes = FALSE)
  axis(1)
  axis(2, at = 1:5, 
       labels = c("Prn\nHH", "Lrg\nHH", "Mix", "Lrg\nJOB", "Prn\nJOB"))
}
mtext("Activity Proportions by Diversity Group for Each D1D Group\nBy Urbanized Area Size", 
      outer = TRUE, line = -0.5)
par(InitPar_ls)
rm(InitPar_ls, ImageDat_D1D2, DispPal_)
```

Once a diversity level has been assigned to each SimBzone it is necessary to determine the specific proportion of activity that is jobs vs. households. This is done by sampling from distributions of the job proportions of activity by diversity group. The following figure shows for urbanized areas as a whole the proportions of block group activity at each job proportion level, segmented by diversity level. Tabulations are made for each named urbanized area in the SLD as well as for each urbanized area size group. These are used as sampling distributions to determine the job proportion of activity of each SimBzone based on the diversity level of the SimBzone. Once the SimBone job proportion is determined, SimBzone activity is split into jobs and households. Before this is done, however, the numbers of jobs and households in each Azone portion of the urbanized area area compared to the Azone inputs and adjustments are made so the SimBzone splits to sum up to the Azone inputs.

```{r echo=FALSE}
#Calculate distribution of job proportions by diversity level
#------------------------------------------------------------
#The proportion of block group activity that is jobs is calculated as a more
#useful diversity measure to use to allocate SimBzone activity between jobs and
#households. For each diversity level, distributions of the jobs proportion of
#activity is calculated. These distributions will be sampling distributions.
TotAct_ <- D_df$TOTACT
PropEmp_ <- D_df$EMPTOT / TotAct_
EmpProp_D2_ls <- list()
for (d2 in D2) {
  DoSelect_ <- as.character(D2Grp_) == d2
  PropEmpToPlot_ <- rep(PropEmp_[DoSelect_], TotAct_[DoSelect_])
  Tmp_HS <- hist(PropEmpToPlot_, plot = FALSE)
  EmpProp_D2_ls[[d2]] <- list(
    Values = Tmp_HS$mids,
    Probs = Tmp_HS$counts / sum(Tmp_HS$counts)
  )
  rm(DoSelect_, PropEmpToPlot_, Tmp_HS)
}
rm(d2)
#Plot the distribution of proportion of employment by diversity level
Opar_ls <- par(mfrow = c(2,3), oma = c(0,0,3,0))
for (d2 in D2) {
  DoSelect_ <- as.character(D2Grp_) == d2
  PropEmpToPlot_ <- rep(PropEmp_[DoSelect_], TotAct_[DoSelect_])
  hist(PropEmpToPlot_, freq = FALSE,
       xlab = "Jobs Proportion",
       main = d2)
  rm(DoSelect_, PropEmpToPlot_)
}
mtext(
  text = paste0("Distribution Block Group Activity by Job Proportion of Activity\n",
                "By Jobs-Household Ratio Level"),
  side = 3, line = 0, outer = TRUE)
par(Opar_ls)
rm(PropEmp_, TotAct_, Opar_ls)
```

# Modeling Destination Accessibility as a Function of Activity Density
Destination accessibility groups are defined in the same way that activity density groups are defined. 20 levels of destination accessibility are defined.

```{r echo=FALSE}
#==================================================================================
#ANALYZE RELATIONSHIP OF DESTINATION ACCESSIBILITY (D5) WITH ACTIVITY DENSITY (D1D)
#==================================================================================

#Create destination accessibility (D5) levels
#--------------------------------------------
#Determine breaks by dividing log of distribution into 20 equal intervals
D5GrpBrk_ <- local({
  LogD5_ <- log(D_df$D5)
  Interval <- diff(range(LogD5_)) / 20
  LogBreaks_ <- min(LogD5_) + 0:20 * Interval
  exp(LogBreaks_)
})
#Make sure the minimum and maximum values bound the breaks
D5GrpBrk_[1] <- min(D_df$D5)
D5GrpBrk_[length(D5GrpBrk_)] <- max(D_df$D5)
#Identify the destination accessibility group for each block group
D5Grp_ <- cut(D_df$D5, breaks = D5GrpBrk_, include.lowest = TRUE)
D_df$D5Grp <- D5Grp_
#Calculate the average destination accessibility for each quantile
D5GrpAve_ <- tapply(D_df$D5, D5Grp_, mean)
#Calculate total activity in each qualtile
D5GrpTotAct_ <- tapply(D_df$TOTACT, D5Grp_, sum)
```

The following figure shows how the distribution of urbanized area activity varies by destination accessibility group.

```{r echo=FALSE}
#Plot the proportion of total activity by group
D5GrpPropAct_ <- D5GrpTotAct_ / sum(D5GrpTotAct_)
plot(1:20, D5GrpPropAct_,
     xlab = "Destination Accessibility (D5) Group",
     ylab = "Proportion of Activity (HHs & Jobs)",
     main = "Proportions of Activity by Destination Accesibility Group\nBlock Groups in All Urbanized Areas")
addSmoothLine(1:20, D5GrpPropAct_, lty = 2)
```

The following figure shows how average destination accessibility varies by density group.

```{r echo=FALSE}
#Plot the group D5 averages
plot(1:20, D5GrpAve_,
     xlab = "Destination Accessibility (D5) Group",
     ylab = "Average Destination Accessibility",
     main = "Average Destination Accessibility by Destination Accessibility Group\nBlocks Groups in All Urbanized Areas")
addSmoothLine(1:20, D5GrpAve_, lty = 2)
#Clean up
rm(D5GrpBrk_, D5Grp_, D5GrpAve_, D5GrpTotAct_, D5GrpPropAct_)
```

Tabulations of activity proportions by destination accessibility group are made for each density group. These tabulations are made for urbanized areas as a whole, for each urbanized area, and for each urbanized area size group. As with the diversity-density tabulations, these tabulations are used as sampling distributions to determine the destination accessibility level for each SimBzone based on the density level.

```{r echo=FALSE}
#Calculate activity jointly by density (D1D) and accessibility (D5) groupings
#----------------------------------------------------------------------------
#Calculate total activity by density group and accessibility group and the
#activity proportions by accessibility group for each density group
#Calculate for all urbanized areas
TotAct_D1D5 <- tapply(D_df$TOTACT, list(D_df$D1DGrp, D_df$D5Grp), sum)
TotAct_D1D5[is.na(TotAct_D1D5)] <- 0
D5ActProp_D1D5 <- sweep(TotAct_D1D5, 1, rowSums(TotAct_D1D5), "/")
#Calculate by urbanized area
D_Ua_df <- split(D_df, D_df$UZA_NAME)
TotAct_Ua_D1D5 <- lapply(D_Ua_df, function(x) {
  TotAct_D1D5 <- tapply(x$TOTACT, list(x$D1DGrp, x$D5Grp), sum)
  TotAct_D1D5
})
D5ActProp_Ua_D1D5 <- lapply(TotAct_Ua_D1D5, function(x) {
  D5ActProp_D1D5 <- sweep(x, 1, rowSums(x, na.rm = TRUE), "/")
  D5ActProp_D1D5
})
ActProp_Ua_D5 <- lapply(D_Ua_df, function(x) {
  TotAct_D5 <- tapply(x$TOTACT, x$D5Grp, sum)
  TotAct_D5[is.na(TotAct_D5)] <- 0
  TotAct_D5 / sum(TotAct_D5)
})
rm(D_Ua_df, TotAct_Ua_D1D5)
#Calculate by urbanized area size group
D_Sz_df <- split(D_df, D_df$UZA_SIZE)
TotAct_Sz_D1D5 <- lapply(D_Sz_df, function(x) {
  TotAct_D1D5 <- tapply(x$TOTACT, list(x$D1DGrp, x$D5Grp), sum)
  TotAct_D1D5
})
D5ActProp_Sz_D1D5 <- lapply(TotAct_Sz_D1D5, function(x) {
  D5ActProp_D1D5 <- sweep(x, 1, rowSums(x, na.rm = TRUE), "/")
  D5ActProp_D1D5
})
ActProp_Sz_D5 <- lapply(D_Sz_df, function(x) {
  TotAct_D5 <- tapply(x$TOTACT, x$D5Grp, sum)
  TotAct_D5[is.na(TotAct_D5)] <- 0
  TotAct_D5 / sum(TotAct_D5)
})
rm(D_Sz_df, TotAct_Sz_D1D5)
#Add size group tabulations to urbanized area list
for (sz in Sz) {
  D5ActProp_Ua_D1D5[[sz]] <- D5ActProp_Sz_D1D5[[sz]]
  ActProp_Ua_D5[[sz]] <- ActProp_Sz_D5[[sz]]
}
rm(D5ActProp_Sz_D1D5, sz)

#Define function to make an image of proportions of total activity by D1D and D5
#-------------------------------------------------------------------------------
imagePropAct <- function(D5PropAct_D1D5, ...) {
  ImgDat_D1D5 <- D5PropAct_D1D5
  ImgDat_D1D5[is.na(ImgDat_D1D5)] <- 0
  DispPal_ <- colorRampPalette(c("black", "yellow"))(20)
  image2D(ImgDat_D1D5, 
          x = 1:20, y = 1:20,
          zlim = c(0,1),
          col = DispPal_, NAcol = "black",
          xlab = "",
          ylab = "",
          ...)
  title(ylab = "D5 Group", line = 1)
  title(xlab = "D1D Group", line = 1)
}
```

The following figure shows the destination accessibility-density relationship for all urbanized areas. The relatively high correlation between block group values can be seen in the figure. Except for the lowest density, all the non-zero values are arranged near the diagonal.

```{r echo=FALSE}
#Plot relationships of D5 group proportions with D1D group
#---------------------------------------------------------
#Plot D5 group proportions for all urbanized areas
imagePropAct(D5ActProp_D1D5,
             main = "Activity Proportions by D5 Group for Each D1D Group\nAll Urbanized Areas")
```

The following figure shows the destination accessibility-density relationship for the same 9 urbanized areas profiled previously. Again, the strength of the relationship between destination accessibility and density can be seen.

```{r echo=FALSE}
#Plot D5 group proportions for 9 urbanized areas
InitPar_ls <- par(mfrow = c(3,3), mar = c(3,3,3,3), oma = c(0, 0, 2.2, 0))
for (Ua in UzaToPlot_) {
  imagePropAct(D5ActProp_Ua_D1D5[[Ua]], main = Ua)
}
mtext("Activity Proportions by D5 Group for Each D1D Group\nFor Selected Urbanized Areas", 
      outer = TRUE, line = -0.5)
par(InitPar_ls)
rm(Ua, InitPar_ls)
```

The following figure shows the relationship between destination accessibility and activity density by urbanized area size group.

```{r echo=FALSE}
#Plot D5 group proportions for 6 urbanized area size groups
InitPar_ls <- par(mfrow = c(3,2), mar = c(3,3,3,3), oma = c(0, 0, 2.2, 0))
for (sz in Sz) {
  imagePropAct(D5ActProp_Ua_D1D5[[sz]], main = sz)
}
mtext("Activity Proportions by D5 Group for Each D1D Group\nBy Urbanized Area Population Size Group", 
      outer = TRUE, line = -0.5)
par(InitPar_ls)
rm(InitPar_ls)
```

The average urbanized area destination accessibility by level is calculated for all urbanized areas, for each urbanized area, and for each urbanized area size category. Once the destination accessibility level for a SimBzone has been determined, the destination accessibility value for the SimBzone is taken from that average value for the level for the urbanized area. Missing urbanized area values are filled in from the overall urbanized area averages.

```{r echo=FALSE}
#Calculate average destination accessibility by destination accessibility group
#------------------------------------------------------------------------------
#Calculate for all urbanized areas
D5Ave_D5 <- unlist(lapply(split(D_df, D_df$D5Grp), function(x) {
  sum(x$TOTACT * x$D5) / sum(x$TOTACT)
}))
#Calculate by urbanized area
Tmp_Ua_df <- split(D_df, D_df$UZA_NAME)
D5Ave_Ua_D5 <- lapply(Tmp_Ua_df, function(x) {
  unlist(lapply(split(x, x$D5Grp), function(y) {
    sum(y$TOTACT * y$D5) / sum(y$TOTACT)
  }))
})
rm(Tmp_Ua_df)
#Calculate by urbanized area size group
Tmp_Sz_df <- split(D_df, D_df$UZA_SIZE)
D5Ave_Sz_D5 <- lapply(Tmp_Sz_df, function(x) {
  unlist(lapply(split(x, x$D5Grp), function(y) {
    sum(y$TOTACT * y$D5) / sum(y$TOTACT)
  }))
})
rm(Tmp_Sz_df)
#Add size group tabulations to urbanized area list
for (sz in Sz) {
  D5Ave_Ua_D5[[sz]] <- D5Ave_Sz_D5[[sz]]
}
rm(D5Ave_Sz_D5)
```

# Implementing Bzone Synthesis
The Bzone synthesis process is implemented in a function which does the following:

1) Azone level urbanized area allocations of households and jobs, and the land area of the urbanized area determine the basic constraints that the zone synthesis must meet.

2) For each Azone portion of the urbanized area SimBzones are created. The number of SimBzones is the whole number result (by rounding) of dividing the total activity allocated to the urbanized area in the Azone by 750. The remainder (positive or negative) is randomly allocated to the SimBzones.

3) Each SimBzone is assigned an activity density level. First a proportional allocation of activity to activity density levels is calculated such that the distribution averages to the average density for the urbanized area in the Azone. This is accomplished by starting with the urbanized area profile of the proportional distribution by activity density level and then adjusting those proportions in small increments until the average density for the urbanized area is achieved. The mean density profile for the urbanized area is used in this calculation. Missing values in the mean density profile are filled in with values from the profile for urbanized areas as a whole. The proportional allocation of activity by density level is used as a sampling distribution to assign each SimBzone a density level. This process is carried out separately for each Azone portion of the urbanized area. A final density adjustment is made so that the average activity density in each Azone urbanized area matches input assumptions.

4) Each SimBzone is assigned a diversity level by using the urbanized area profile of proportional allocations of activity by diversity level for each activity density level for sampling distributions. Missing values are filled in from the overall urbanized area profile. Once the diversity level has been chosen a value for the employment proportion of activity is picked for each SimBzone from the employment proportions distributions. The employment proportions are used to make an initial calculation of employment in each SimBzone. That employment is adjusted as necessary to match the urbanized area employment total in each Azone. The number of SimBzone households is then calculated.

5) SimBzones are assigned destination accessibility levels in a similar manner. The urbanized area profile of proportional allocations of activity by destination accessibility level are used as sampling distributions to determine destination accessibility levels. Missing values are filled in from the overall urbanized area profile. The specific destination accessibility values are the mean values computed for the levels for the urbanized area. If destination accessibility proportions are missing for one or more density levels, they are filled in from the profile for urbanized areas as a whole.

```{r echo=FALSE}
#Make a list containing all of the estimated distributions
#---------------------------------------------------------
Ds_ls <- list(
  D1DGrp_ls =  D1DGrp_ls,
  D1DGrp_Ua_ls = D1DGrp_Ua_ls,
  ActProp_Ua_D2 = ActProp_Ua_D2,
  ActProp_Ua_D5 = ActProp_Ua_D5,
  D2ActProp_D1D2 = D2ActProp_D1D2,
  D2ActProp_Ua_D1D2 = D2ActProp_Ua_D1D2,
  EmpProp_D2_ls = EmpProp_D2_ls,
  D5ActProp_D1D5 = D5ActProp_D1D5,
  D5ActProp_Ua_D1D5 = D5ActProp_Ua_D1D5,
  D5Ave_D5 = D5Ave_D5,
  D5Ave_Ua_D5 = D5Ave_Ua_D5
)

#Define function to create Bzones
#--------------------------------
createUbzBzones <- 
  function(TotHHs_Az, TotJobs_Az, ActDen_Az, UzaName, ActPerBzone = 750) {
    set.seed(1)
    Az <- names(TotHHs_Az)
    #----------------------------------------
    #Create zones to accommodate all activity
    #----------------------------------------
    TotAct_Az <- TotHHs_Az + TotJobs_Az
    SimBzone_Az_df <- lapply(TotAct_Az, function(x) {
      NumZones <- round(x / ActPerBzone)
      RemAct <- x - NumZones * ActPerBzone
      BzName_Bz <- paste0("B", 1:NumZones)
      BzAct_Bz <- rep(ActPerBzone, NumZones)
      if (RemAct != 0) {
        RemAllocation_tb <- table(sample(1:NumZones, abs(RemAct), replace = TRUE))
        Idx_ <- as.numeric(names(RemAllocation_tb))
        Vals_ <- as.vector(RemAllocation_tb)
        BzAct_Bz[Idx_] <- BzAct_Bz[Idx_] + Vals_ * sign(RemAct)
      }
      data.frame(
        Bzone = BzName_Bz,
        Activity = BzAct_Bz
      )
    })
    #Add Azone name onto front of Bzone name
    for (az in Az) {
      SimBzone_Az_df[[az]]$Bzone <- paste0(az, SimBzone_Az_df[[az]]$Bzone)
    }
    #--------------------------------
    #Assign density levels and values
    #--------------------------------
    #Calculate density distributions by Azone
    ActDen_Az_df <- lapply(ActDen_Az, function(x) {
      ActDen_df <- adjDenDist(
        UzaAveDensity_ = Ds_ls$D1DGrp_Ua_ls[[UzaName]]$AveDensity,
        DenDist_ = Ds_ls$D1DGrp_Ua_ls[[UzaName]]$PropActivity,
        Target = x,
        AveDensity_ = Ds_ls$D1DGrp_ls$AveDensity
      )
    })
    #Assign density level and average density to zones
    for (az in Az) {
      D1Grp_Bz <- sample(1:20, nrow(SimBzone_Az_df[[az]]), replace = TRUE,
                         ActDen_Az_df[[az]]$ActProp)
      ActDen_Bz <- ActDen_Az_df[[az]]$AveDensity[D1Grp_Bz]
      #Adjust density to meet target
      Activity_Bz <- SimBzone_Az_df[[az]]$Activity
      AveDensity <- sum(Activity_Bz) / sum(Activity_Bz * 1 / ActDen_Bz)
      ActDen_Bz <- ActDen_Bz * ActDen_Az[az] / AveDensity
      SimBzone_Az_df[[az]]$D1Grp <- D1Grp_Bz
      SimBzone_Az_df[[az]]$D1D <- ActDen_Bz
      rm(D1Grp_Bz, ActDen_Bz, Activity_Bz, AveDensity)
    }
    rm(ActDen_Az_df)
    #----------------------------------------------------------------------
    #Functions to assist diversity and destination accessibility assignment
    #----------------------------------------------------------------------
    #Define function to fill in missing activity proportions
    fillMissingUaActProp <- function(UaActProp_mx, ActProp_mx) {
      NaRows_ <- apply(UaActProp_mx, 1, function(x) all(is.na(x)))
      UaActProp_mx[NaRows_,] <- ActProp_mx[NaRows_,]
      UaActProp_mx[is.na(UaActProp_mx)] <- 0
      UaActProp_mx
    }
    #Define function to identify urbanized area size category
    idSizeCategory <- function(TotActivity) {
      SzBrk_ <- c(0, 5e4, 1e5, 5e5, 1e6, 5e6, 1e9)
      Sz <- c("small", "medium-small", "medium", "medium-large", "large", "very-large")
      as.character(cut(TotActivity, SzBrk_, labels = Sz, include.lowest = TRUE))
    }
    #--------------------------------------------------------
    #Assign diversity level and calculate households and jobs
    #--------------------------------------------------------
    #Create D2 activity proportions sampling matrix
    D2ActProp_D1D2 <- 
      fillMissingUaActProp(
        Ds_ls$D2ActProp_Ua_D1D2[[UzaName]], Ds_ls$D2ActProp_D1D2)
    #Assign diversity values to Bzones
    for (az in Az) {
      SimBzone_df <- SimBzone_Az_df[[az]]
      #Assign diversity levels
      D2Grp_Bz <- sapply(SimBzone_df$D1Grp, function(x) {
        sample(colnames(D2ActProp_D1D2), 1, prob = D2ActProp_D1D2[x,])
      })
      #Calculate numbers of jobs by Bzone
      EmpProp_Bz <- sapply(D2Grp_Bz, function(x) {
        Sample_ls <- Ds_ls$EmpProp_D2_ls[[x]]
        sample(Sample_ls$Values, 1, prob = Sample_ls$Probs)
      })
      Jobs_Bz <- unname(round(SimBzone_df$Activity * EmpProp_Bz))
      #Adjust jobs to match inputs
      TargetJobs <- TotJobs_Az[az]
      JobDiff <- TargetJobs - sum(Jobs_Bz)
      JobsProp_Bz <- Jobs_Bz / sum(Jobs_Bz)
      JobsAdj_Bx <- table(
        sample(1:nrow(SimBzone_df), abs(JobDiff), replace = TRUE, prob = JobsProp_Bz)
      )
      AdjJobs_Bz <- Jobs_Bz
      AdjIdx_ <- as.numeric(names(JobsAdj_Bx))
      AdjJobs_Bz[AdjIdx_] <- Jobs_Bz[AdjIdx_] + sign(JobDiff) * JobsAdj_Bx
      if(any(AdjJobs_Bz > SimBzone_df$Activity)) {
        AdjJobs_Bz[AdjJobs_Bz > SimBzone_df$Activity] <- 
          SimBzone_df$Activity[AdjJobs_Bz > SimBzone_df$Activity]
      }
      #Calculate numbers of households by Bzone
      HHs_Bz <- SimBzone_df$Activity - AdjJobs_Bz
      #Calculate Jobs - HHs ratio
      D2A_JPHH_Bz <- AdjJobs_Bz / HHs_Bz
      #Assign values to SimBzones
      SimBzone_Az_df[[az]]$D2Grp <- D2Grp_Bz
      SimBzone_Az_df[[az]]$Jobs <- AdjJobs_Bz
      SimBzone_Az_df[[az]]$HHs <- HHs_Bz
      SimBzone_Az_df[[az]]$`D2A_JPHH` <- D2A_JPHH_Bz
    }
    #--------------------------------------
    #Assign Destination Accessibility Level
    #--------------------------------------
    #Calculate total urbanized activity and identify size group
    # TotAct <- sum(TotHHs_Az) + sum(TotJobs_Az) 
    # UaSizeGrp <- idSizeCategory(TotAct)
    #Create D5 activity proportions sampling matrix
    # D5ActProp_D1D5 <- 
    #   fillMissingUaActProp(
    #     Ds_ls$D5ActProp_Ua_D1D5[[UzaName]], Ds_ls$D5ActProp_Ua_D1D5[[UaSizeGrp]])
    D5ActProp_D1D5 <- 
      fillMissingUaActProp(
        Ds_ls$D5ActProp_Ua_D1D5[[UzaName]], Ds_ls$D5ActProp_D1D5)
    #Create vector of D5 averages by activity density level
    D5Ave_D5 <- Ds_ls$D5Ave_Ua_D5[[UzaName]]
    #D5SzGrpAve_D5 <- Ds_ls$D5Ave_Ua_D5[[UaSizeGrp]]
    #D5Ave_D5[is.nan(D5Ave_D5)] <- D5SzGrpAve_D5[is.nan(D5Ave_D5)]
    D5Ave_D5[is.nan(D5Ave_D5)] <- Ds_ls$D5Ave_D5[is.nan(D5Ave_D5)]
    #Fill in missing values as moving average
    D5Ave_D5[is.nan(D5Ave_D5)] <- 0
    D5MovAve_D5 <- 
      (c(D5Ave_D5[-1], max(D5Ave_D5)) + c(min(D5Ave_D5), D5Ave_D5[1:19])) / 2
    D5Ave_D5[D5Ave_D5 == 0] <- D5MovAve_D5[D5Ave_D5 == 0]
    #Iterate through Azones to assign destination accessibility
    for (az in Az) {
      SimBzone_df <- SimBzone_Az_df[[az]]
      #Assign D5Grp
      D5Grp_Bz <- sapply(SimBzone_df$D1Grp, function(x) {
        sample(1:ncol(D5ActProp_D1D5), 1, prob = D5ActProp_D1D5[x,])
      })
      SimBzone_Az_df[[az]]$D5Grp <- D5Grp_Bz
      #Assign D5 value
      D5_Bz <- D5Ave_D5[D5Grp_Bz]
      SimBzone_Az_df[[az]]$D5 <- D5_Bz
    }
    #------------------
    #Return the results
    #------------------
    SimBzone_Az_df
  }
```

The zone synthesis process is tested by synthesizing Bzones for the 9 example metropolitan areas and comparing the activity proportions for the synthesized zone with the activity proportions for the metropolitan area block groups. In general, the synthesis process does a good job of replicating the observed values for these metropolitan areas. The process tends to overestimate the proportion of activity in higher density levels. This in turn results in an overestimation of the proportion of activity in 'primarily-job' oriented Bzones and in higher destination accessibility groups. This may be a result of the combination of applying the density adjustment function at the Azone level (i.e. county level for this test). The largest difference can be seen for the New York urbanized area.

```{r echo=FALSE}
#Define function to create test Azone inputs
#-------------------------------------------
#This function takes the data for an urbanized area and splits it at the county
#level considering counties to be Azones. It then produces each of the required
#inputs at the Azone level.
createAzoneTestData <- function(UzaName) {
  Uza_df <- D_df[D_df$UZA_NAME == UzaName,]
  Azone_Bg <- 
    paste0("A", sapply(Uza_df$GEOID10, function(x) substr(x, 1, 5)))
  TotJobs_Az <- tapply(Uza_df$EMPTOT, Azone_Bg, sum)
  TotHHs_Az <- tapply(Uza_df$HH, Azone_Bg, sum)
  TotAct_Az <- tapply(Uza_df$TOTACT, Azone_Bg, sum)
  TotAcLand_Az <- tapply(Uza_df$AC_LAND, Azone_Bg, sum)
  ActDen_Az <- TotAct_Az / TotAcLand_Az
  list(
    TotJobs_Az = TotJobs_Az,
    TotHHs_Az = TotHHs_Az,
    ActDen_Az = ActDen_Az
  )
}

#Define function to tabulate activity by group(s)
tabSimBzoneAct <- function(SimBzones_Az_df, VarNames, VarDimnames) {
  SimB_df <- do.call(rbind, SimBzones_Az_df)
  if (length(VarNames) == 1) {
    Act_ <- numeric(length(VarDimnames))
    names(Act_) <- VarDimnames
    TabAct_ <- tapply(SimB_df$Activity, SimB_df[[VarNames]], sum)
    Act_[names(TabAct_)] <- TabAct_
    Act_[is.na(Act_)] <- 0
    return(Act_)
  }
  if (length(VarNames) == 2) {
    Act_mx <- matrix(0, nrow = length(VarDimnames[[1]]), ncol = length(VarDimnames[[2]]))
    dimnames(Act_mx) <- VarDimnames
    TabAct_mx <- with(SimB_df, tapply(Activity, as.list(SimB_df[VarNames]), sum))
    Idx_mx <- as.matrix(expand.grid(rownames(TabAct_mx), colnames(TabAct_mx)))
    Act_mx[Idx_mx] <- as.vector(TabAct_mx)
    Act_mx[is.na(Act_mx)] <- 0
    return(Act_mx)
  }
  stop("Can't tabulate by more than 2 VarNames")
}

#Define function to summarize SimBzone activity
summarizeSimBzones <- function(SimBzones_Az_df) {
  D1Names <- as.character(1:20)
  D5Names <- as.character(1:20)
  D2Names <- 
    c("primarily-hh", "largely-hh", "mixed", "largely-job", "primarily-job")
  Act_D1 <- tabSimBzoneAct(SimBzones_Az_df, "D1Grp", D1Names)
  PropAct_D1 <- Act_D1 / sum(Act_D1)
  Act_D2 <- tabSimBzoneAct(SimBzones_Az_df, "D2Grp", D2Names)
  PropAct_D2 <- Act_D2 / sum(Act_D2)
  Act_D5 <- tabSimBzoneAct(SimBzones_Az_df, "D5Grp", D5Names)
  PropAct_D5 <- Act_D5 / sum(Act_D5)
  Act_D1D2 <- tabSimBzoneAct(SimBzones_Az_df, c("D1Grp", "D2Grp"), list(D1Names, D2Names))
  D2PropAct_D1D2 <- sweep(Act_D1D2, 1, rowSums(Act_D1D2), "/")
  D2PropAct_D1D2[is.na(D2PropAct_D1D2)] <- 0
  Act_D1D5 <- tabSimBzoneAct(SimBzones_Az_df, c("D1Grp", "D5Grp"), list(D1Names, D5Names))
  D5PropAct_D1D5 <- sweep(Act_D1D5, 1, rowSums(Act_D1D5), "/")
  D5PropAct_D1D5[is.na(D5PropAct_D1D5)] <- 0
  list(
    PropAct_D1 = PropAct_D1,
    PropAct_D2 = PropAct_D2,
    PropAct_D5 = PropAct_D5,
    PropAct_D1D2 = D2PropAct_D1D2,
    PropAct_D1D5 = D5PropAct_D1D5
  )
}

#Define function to collate observed urbanized area values
collateUbzSummaries <- function(Ds_ls, UaName) {
  PropAct_D1 <- Ds_ls$D1DGrp_Ua_ls[[UaName]]$PropActivity
  PropAct_D1[is.na(PropAct_D1)] <- 0
  PropAct_D2 <- Ds_ls$ActProp_Ua_D2[[UaName]]
  PropAct_D2[is.na(PropAct_D2)] <- 0
  PropAct_D5 <- Ds_ls$ActProp_Ua_D5[[UaName]]
  PropAct_D5[is.na(PropAct_D5)] <- 0
  PropAct_D1D2 <- D2ActProp_Ua_D1D2[[UaName]]
  PropAct_D1D2[is.na(PropAct_D1D2)] <- 0
  PropAct_D1D5 <- D5ActProp_Ua_D1D5[[UaName]]
  PropAct_D1D5[is.na(PropAct_D1D5)] <- 0
  list(
    PropAct_D1 = PropAct_D1,
    PropAct_D2 = PropAct_D2,
    PropAct_D5 = PropAct_D5,
    PropAct_D1D2 = PropAct_D1D2,
    PropAct_D1D5 = PropAct_D1D5
  )
}

#Comparison vector plot
vectorCompare <- function(Est_ls, Obs_ls, VectorName, ...) {
  Obs_ <- Obs_ls[[VectorName]]
  Est_ <- Est_ls[[VectorName]]
  Ylim_ <- range(c(Obs_, Est_))
  Xlab_ <- names(Obs_)
  Xvals_ <- 1:length(Xlab_)
  plot(Xvals_, Obs_, axes = FALSE, ylim = Ylim_, type = "p", 
       ylab = "Activity Proportion",
       ...)
  box()
  axis(2)
  axis(1, at = Xvals_, labels = Xlab_)
  lines(Xvals_, Obs_)
  points(Xvals_, Est_, col = "red")
  lines(Xvals_, Est_, col = "red")
}

#Function to compare observed and estimated density, diversity, and destination accessibility proportions
compareObsEst1 <- function(Obs_ls, Est_ls, UaName) {
  Opar_ls <- par(mfrow = c(2,2), oma = c(0,0,3,0))
  vectorCompare(Est_ls, Obs_ls, "PropAct_D1", 
                xlab = "Density (D1D) Category",
                main = "Density")
  legend("topleft", legend = c("Observed", "Simulated"), bty = "n", lty = 1, col = c("black", "red"))
  vectorCompare(Est_ls, Obs_ls, "PropAct_D2", 
                xlab = "Diversity (D2A_JPHH) Category",
                main = "Diversity")
  vectorCompare(Est_ls, Obs_ls, "PropAct_D5", 
                xlab = "Destination Accessibility (D5) Category",
                main = "Destination Accessibility")
  mtext(text = paste0(UaName, "\nComparison of Density, Diversity, and Destination Accessibility Activity Proportions"), 
        side = 3, outer = TRUE)
  par(Opar_ls)
}

#Function to compare joint distributions
compareObsEst2 <- function(Obs_ls, Est_ls, UaName) {
  Opar_ls <- par(mfrow = c(2,2), oma = c(0,0,3,0))
  Rng_ <- 
    range(c(as.vector(Obs_ls$PropAct_D1D2), as.vector(Est_ls$PropAct_D1D2)))
  DispPal_ <- colorRampPalette(c("black", "yellow"))(20)
  image2D(Obs_ls$PropAct_D1D2, 
          x = 1:20, y = 1:5,
          xlab = "D1D Group", ylab = "D2 Group",
          zlim = Rng_,
          col = DispPal_, NAcol = "black",
          main = paste("Observed Diversity Proportions\nBy Density group"),
          cex.main = 1)
  image2D(Est_ls$PropAct_D1D2, 
          x = 1:20, y = 1:5,
          xlab = "D1D Group", ylab = "D2 Group",
          zlim = Rng_,
          col = DispPal_, NAcol = "black",
          main = paste("Synthesized Diversity Proportions\nBy Density group"),
          cex.main = 1)
  Rng_ <- 
    range(c(as.vector(Obs_ls$PropAct_D1D5), as.vector(Est_ls$PropAct_D1D5)))
  DispPal_ <- colorRampPalette(c("black", "yellow"))(20)
  image2D(Obs_ls$PropAct_D1D5, 
          x = 1:20, y = 1:20,
          xlab = "D1D Group", ylab = "D5 Group",
          zlim = Rng_,
          col = DispPal_, NAcol = "black",
          main = paste("Observed Destination Accessibility Proportions\nBy Density group"),
          cex.main = 1)
  image2D(Est_ls$PropAct_D1D5, 
          x = 1:20, y = 1:20,
          xlab = "D1D Group", ylab = "D5 Group",
          zlim = Rng_,
          col = DispPal_, NAcol = "black",
          main = paste("Synthesized Destination Accessibility Proportions\nBy Density group"),
          cex.main = 1)
  mtext(text = paste0(UaName, "\nComparison of Diversity and Destination Accessibility Proportions by Density Group"), 
        side = 3, outer = TRUE)
  par(Opar_ls)
}

#Function to simulate for urbanized area and compare with observed values
simulateAndCompare <- function(UzaName) {
  UzaTestData_ls <- createAzoneTestData(UzaName)
  SimBzones_Az_df <- 
    createUbzBzones(
      TotHHs_Az = UzaTestData_ls$TotHHs_Az, 
      TotJobs_Az = UzaTestData_ls$TotJobs_Az, 
      ActDen_Az = UzaTestData_ls$ActDen_Az, 
      UzaName = UzaName, 
      ActPerBzone = 750)

  Est_ls <- summarizeSimBzones(SimBzones_Az_df)
  Obs_ls <- collateUbzSummaries(Ds_ls, UzaName)
  compareObsEst1(Obs_ls, Est_ls, UzaName)
  compareObsEst2(Obs_ls, Est_ls, UzaName)
}
```

The following 2 figures compare the simulated and observed values for the Atlanta, GA urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
#Simulate and compare several urbanized areas
simulateAndCompare("Atlanta, GA")
```

The following 2 figures compare the simulated and observed values for the Jacksonville, FL urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Jacksonville, FL")
```

The following 2 figures compare the simulated and observed values for the Cincinnati, OH urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Cincinnati, OH")
```

The following 2 figures compare the simulated and observed values for the Dallas--Fort Worth--Arlington, TX urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Dallas-Fort Worth-Arlington, TX")
```

The following 2 figures compare the simulated and observed values for the Baltimore, MD urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Baltimore, MD")
```

The following 2 figures compare the simulated and observed values for the Denver--Aurora, CO urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Denver-Aurora, CO")
```

The following 2 figures compare the simulated and observed values for the Portland, OR urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("Portland, OR")
```

The following 2 figures compare the simulated and observed values for the San Francisco--Oakland, CA urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("San Francisco-Oakland, CA")
```

The following 2 figures compare the simulated and observed values for the New York--Newark, NY urbanized area. The first figure compares the overall density, diversity, and destination accessibility distributions. The second figure compares the joint diversity and density distribution and the joint destination accessibility and density distribution.

```{r echo=FALSE}
simulateAndCompare("New York-Newark, NY")
```

### 2020-08-26 sources/tools/helpers.R
Re-re-removed obsolete helpers.R

@@ -1,55 +1,0 @@
-# Author: Jeremy Raw
-
-# VisionEval helper tools
-# NOTE: these all depend on having set up the VisionEval environment
-# See VisionEval.R in the installation runtime (VE-Installer/boilerplate)
-# In particular, functions call models at locations relative to ve.root
-
-tool.contents <- c("vegui","verpat","verspm","vestate")
-
-# Function starts the VEGUI
-vegui <- function(root=ve.root) {
-  library("shiny")
-  full.path <- file.path(root,"VEGUI")
-  owd <- setwd(full.path) 
-  runApp('../VEGUI')
-  setwd(owd)
-}
-
-# The following two functions run the command line model versions per the
-# Getting Started document.  Optional "scenarios" argument, if TRUE, will
-# run the scenarios version of the test models.
-verpat <- function(scenarios=FALSE,baseyear=FALSE,root=ve.root) {
-  if ( ! scenarios ) {
-    if ( ! baseyear ) {
-      full.path <- file.path(root,"models/VERPAT")
-    } else {
-      full.path <- file.path(root,"models/BaseYearVERPAT")
-    }
-  } else {
-    full.path <- file.path(root,"models/VERPAT_Scenarios")
-  }
-  owd <- setwd(full.path)
-  source("run_model.R")
-  setwd(owd)
-}
-
-verspm <- function(scenarios=FALSE,root=ve.root) {
-  if ( ! scenarios ) {
-    full.path <- file.path(root,"models/VERSPM")
-    test.dir <- file.path(full.path,"Test1") # Older structure for VERSPM
-    if ( dir.exists(test.dir) ) full.path <- test.dir
-  } else {
-    full.path <- file.path(root,"models/VERSPM_Scenarios")
-  }
-  owd <- setwd(full.path)
-  source("run_model.R")
-  setwd(owd)
-}
-
-vestate <- function(root=ve.root) {
-  full.path <- file.path(root,"models/VE-State")
-  owd <- setwd(full.path)
-  source("run_model.R")
-  setwd(owd)
-}

### 2020-08-26 build/VisionEval-dev.R
Made VE_EXPRESS the default for ve.build() [Much faster build]

@@ -53,7 +53,8 @@ evalq(
       targets=character(0),
       r.version=paste(R.version[c("major","minor")],collapse="."),
       config=character(0), # defaults to config/VE-config.yml
-      flags=character(0),
+      express=TRUE, # if FALSE, will run devtools::check; otherwise build as rapidly as possible
+      flags=character(0), # flags for VE_CONFIG or VE_EXPRESS will override the direct flag 
       use.git=FALSE
     ) {
       owd<-setwd(file.path(ve.root,"build"))
@@ -70,6 +71,11 @@ evalq(
             if ( file.exists(config) ) {
               Sys.setenv(VE_CONFIG=config)
             }
+          }
+          if ( express ) {
+            Sys.setenv(VE_EXPRESS="YES")
+          } else {
+            Sys.unsetenv("VE_EXPRESS")
           }
           # Force our own version of VE_BRANCH
           Sys.setenv(VE_BRANCH=getLocalBranch(ve.root,use.git))

### 2020-08-27 sources/framework/visioneval/R/query.R
Removed obsolete ModelState management causing spurious warnings

@@ -330,7 +330,6 @@ readDatastoreTables <- function(Tables_ls, Group, QueryPrep_ls) {
         }
       }
     }
-    rm(ModelState_ls, pos = 1)
     Out_ls[[tb]] <- data.frame(Out_ls[[tb]])
   }
   #Identify missing datasets

### 2020-08-27 build/travis/install-deps.R
Fixed dependency needed for Travis

@@ -1,9 +1,15 @@
 #!/bin/env Rscript
 
 # Load and install R package dependencies
+# These get put into ve-lib on Travis (since we don't have a separate
+# dev-lib there). They also become part of the package cache.
 
 if ( ! suppressWarnings(require("yaml")) ) {
   cat("Installing YAML package\n")
+  install.packages("yaml", repos="https://cloud.r-project.org", dependencies=NA)
+}
+if ( ! suppressWarnings(require("roxygen2")) ) {
+  cat("Installing roxygen2 package\n")
   install.packages("yaml", repos="https://cloud.r-project.org", dependencies=NA)
 }
 

### 2020-08-27 sources/framework/visioneval/R/query.R
Added query functionality to VEModel

@@ -801,7 +801,7 @@ summarizeDatasets <-
       Data_ls <- readDatastoreTables(Tables_ls, Group, QueryPrep_ls)
       #Stop if any of the datasets are missing
       if (length(Data_ls$Missing[[Table]]) != 0) {
-        MissingDsets_ <- paste(Data_ls$Missing[[Table]], collapse = ", ")
+        MissingDsets_ <- paste(names(Data_ls$Missing[[Table]]), collapse = ", ")
         Msg <- paste("The following datasets are not present in the",
                      Table, "table", "in the", Group, "group:", MissingDsets_)
         stop(Msg)

### 2020-08-27 sources/models/VERSPM/Query-Spec.R

@@ -33,7 +33,7 @@
 #           units of the corresponding dataset.
 #    "By" - a vector of Datasets along which to break out the Measure. Usually just "Marea".
 #           If doing Breaks, a vector of two, with the first one being the sub-breaks (e.g.
-#           PctMedianIncome).
+#           Income).
 #    "Table" - a vector of one or more tables that will be joined; If Table has more than one
 #           entry, you must specify "Key" to say how they are joined.
 #  OPTIONAL
@@ -208,33 +208,6 @@ PMSpecifications <- list(
     ),
     Units = "Persons",
     Description = "Number of persons residing in mixed use in urban area in RVMPO county"
-  ),
-
-  #Urban Population in mixed use in RVMPO county, broken out by income
-  #-------------------
-  list(
-    Name = "UrbanHhPopLowMFI_Ma",
-    Summarize = list(
-      Expr = "sum(HhSize[LocType == 'Urban'])",
-      Units_ = c(
-        HhSize = "",
-        LocType = "",
-        PctMedianIncome = "",
-        Marea = ""
-      ),
-      By = c(
-        "PctMedianIncome",
-        "Marea"),
-      Breaks = list(
-        PctMedianIncome = c(0.30, 0.50, 0.80, 1.00, 1.20, 2.00)
-      ),
-      BreakNames = list(
-        PctMedianIncome = c("030pct", "050pct", "080pct", "100pct", "120pct","200pct")
-      ),
-      Table = "Household"
-    ),
-    Units = "Persons",
-    Description = "Number of persons in households residing in urban area by Percent Median Family Income"
   ),
 
   #Urban Population in mixed use in RVMPO county, broken out by income

### 2020-08-27 sources/tools/Queries/Measures_Demo.R

@@ -1,9 +1,9 @@
 #Measures_Demo.R
 
-#---------------
-#Packages to Use
-#---------------
-library(visioneval)
+#-----------------------------------------
+# visioneval has the query functions in it
+#-----------------------------------------
+requireNamespace("visioneval")
 
 #---------------------------
 #Prepare for datastore query
@@ -12,7 +12,7 @@ library(visioneval)
 #1) Path to datastore
 #2) Functions for reading data given the datastore type
 #3) Datastore listing 
-QPrep_ls <- prepareForDatastoreQuery(
+QPrep_ls <- visioneval::prepareForDatastoreQuery(
   DstoreLocs_ = "Datastore", 
   DstoreType = "RD")
 
@@ -23,7 +23,7 @@ QPrep_ls <- prepareForDatastoreQuery(
 #Archive is organized by group. Within each group folder is a set of CSV files,
 #one for each table in the group. Each CSV file lists the datasets included in
 #the table giving the dataset name, data type, units, and description.
-documentDatastoreTables(
+visioneval::documentDatastoreTables(
   SaveArchiveName = "DatastoreDocumentation", 
   QueryPrep_ls = QPrep_ls)
 
@@ -62,7 +62,7 @@ TablesRequest_ls <- list(
 #Call the readDatastoreTables function using the list of requested tables and
 #datasets
 TableResults_ls <-
-  readDatastoreTables(
+  visioneval::readDatastoreTables(
     Tables_ls = TablesRequest_ls,
     Group = "2020",
     QueryPrep_ls = QPrep_ls
@@ -186,7 +186,7 @@ TableResults_ls$Missing
 #Total household DVMT is in the Household table while the proportion of
 #household DVMT by vehicle access type is in the Vehicle table. The two tables
 #are joined using the HhId.
-summarizeDatasets(
+visioneval::summarizeDatasets(
   Expr = "sum((Dvmt * DvmtProp)[LocType == 'Urban' & VehicleAccess %in% c('LowCarSvc', 'HighCarSvc')])",
   Units_ = c(
     Dvmt = "",
@@ -204,4 +204,3 @@ summarizeDatasets(
   Group = "2020",
   QueryPrep_ls = QPrep_ls
 )
-

### 2020-08-27 sources/tools/Queries/QueryDocs.R

@@ -1,232 +1,0 @@
-# This needs to become a vignette, if it's not entirely redundant
-# with the function's Roxygen header.
-
-#================
-#QueryDatastore.R
-#================
-
-#<doc>
-## QueryDatastore Module
-#### February 13, 2020
-#
-#This module defines functions used to query the Datastore and produce summary measures from the datastore. This documentation explains the overall approach to querying and summarization enabled by these functions, and the higher level functions to produce summaries. Refer to the function documentation for explanations of the lower level functions and details on all functions.
-#
-### Summary
-#
-#These functions enable users to specify datasets to be read from the datastore and calculations to be performed on those datasets.
-#- Calculations are specified as standard R language expressions (see below)
-#- Calculations may be grouped: for example, measures calculated by Azone
-#- Groupings may be specified: for example, income groups may be specified
-#- Region summaries can be produced for a set of measures specified in a json-formatted definitions file
-#
-#There are some limitations to the calculations that may be performed, however:
-#- All datasets identified in a calculation must be in the same table. The functions at the present time do not allow you to use datasets in different tables (e.g. Households and Vehicles) in a calculation.
-#- Summarization functions are limited to:
-#  - sum: sum the values in the dataset,
-#  - count: count the number of values in the dataset,
-#  - mean: calculate the arithmetic average of values in the dataset,
-#  - wtmean: calculate the weighted mean of values in the dataset
-#- Operators in the calculation expression are limited to '+', '-', '*', and '/'
-#
-#Before any datasets can be retrieved and summarized, the 'prepareForDatastoreQuery' function must be called in order to establish information about the datastore location(s), type, and structure(s). This information is used by query and summarization functions. This function has 2 arguments:
[…]

### 2020-08-27 sources/tools/Queries/Sample-Query-Script.R

@@ -1,425 +1,0 @@
-# Sample_Measures_Script.R
-# This script will process a list of measure specifications across the Years in a Scenario Datastore
-# and generate a .csv file that contains the measures.
-
-# Default/Example measures in Test-Spec.R will work with output from the sample VE_RSPM. Adjust the
-# Test-Spec.R file for your own MArea and Azone geography and it should work with your local version
-# of VE_RSPM.
-
-# CONFIGURE THE FOLLOWING PARAMETERS FOR THE SCRIPT RUN:
-
-# SpecFile - (optional) file name containing measure specifications
-#               default/examples: PM-Spec.R in folder with this script
-# ScenarioRoot - the directory relative to which to look for scenarios
-# Scenarios - vector of directories to find the Datastore from which to extract the measures
-#            (each Scenario is sought as subdirectory of ScenarioRoot)
-# OutputFile - (optional) file name in which to put the measures (default is long but informative)
-#              Can be a string or it can contain a "%s" two letter parameter that gets replaced
-#              with "basename(Scenario)
-# Years - which years in the Datastore to process measures for
-# Geography - which geography (Type/Value) to create measures for
-# Median Income - (optional) value from which to build PctMedianIncome for each Household to use in making measures
-
-# SpecFile contains a list of specifications for the measures to generate (see the sample file for format)
-SpecFile <- "Sample-VERSPM-Spec.R"
-#   (default is "PM-Spec.R" in the same folder as this script)
-# SpecFile contains a definition of PMSpecifications (a list of specifications; see the sample file for format)
-# SpecFile should contain measure specs for whatever Geography you specified abaove (in the By and Units Summarize fields)
-# For different geography types, make another PM-Spec-Azone.R (e.g.) and in the summarize blocks,
-# replace 'By = "Marea"' with 'By="Azone"' and replace 'Marea = ""' with 'Azone = ""' in the Summarize Units
-
-# ScenarioRoot is a path (absolute or relative) to a directory that contains the Scenarios
-ScenarioRoot <- file.path(ve.root,"ModelState")
-
-# 'Scenarios' is a vector of names of subfolders of 'ScenarioRoot' to be a directory containing a
[…]

### 2020-08-27 sources/tools/models.R

@@ -667,6 +667,373 @@ ve.model.extract <- function(
   invisible(results)
 }
 
+# Query.R
+# This script provides illustrative functions for doing Datastore summary queries, using a SpecFile
+
+###########################################################################
+# Required libraries
+# Need to affix namespace resolution operator to use functions
+
+requireNamespace("stringr")
+
+###########################################################################
+# FUNCTION DEFINITIONS - helpers
+###########################################################################
+
+###########################################################################
+# FUNCTION: makeMeasure
+#
+# Process a measureSpec for a Year and Geography
+# Return the measureName, with the side effect that the value(s) of the measure
+# are placed into measureEnv (whence they will later be summarized)
+#
+makeMeasure <- function(measureSpec,thisYear,Geography,QPrep_ls,measureEnv) {
+  GeoValue <- Geography["Value"]
+  measureName <- measureSpec$Name
+
+  # Skip or include measures based on presence of required Dataset
+  if ( "Require" %in% names(measureSpec) ) {
+    if ( ! visioneval::isDatasetPresent(measureSpec$Require["Dataset"], measureSpec$Require["Table"], thisYear, QPrep_ls) ) {
+      return(paste(measureName,"(SKIPPED due to Require:)"))
+    }
+  } else
+  if ( "RequireNot" %in% names(measureSpec) ) {
+    if ( visioneval::isDatasetPresent(measureSpec$Require["Dataset"], measureSpec$Require["Table"], thisYear, QPrep_ls) ) {
+      return(paste(measureName,"(SKIPPED due to RequireNot:)"))
+    }
+  }
+
+  # Compute the measure based on the measureSpec     
+  if ( "Function" %in% names(measureSpec) ) {
+    measure <- eval(parse(text=measureSpec$Function), envir=measureEnv)
+    names(measure) <- measureName
+  } else
+  if ( "Summarize" %in% names(measureSpec) ) {
+    sumSpec <- measureSpec$Summarize
+    if ( ! "By" %in% names(sumSpec) ||
+         ! Geography["Type"] %in% sumSpec$By ) {
+      stop(paste("Script wants Geography Type ",Geography["Type"]," in 'By' but got ",sumSpec$By,"",sep="'"))
+    }
+    usingBreaks <- "Breaks" %in% names(sumSpec) && ! is.null(sumSpec$Breaks)
[…]

### 2020-08-27 sources/tools/models.R
First cut at allowing regional measures, not just Marea/Azone

@@ -688,7 +688,14 @@ requireNamespace("stringr")
 # are placed into measureEnv (whence they will later be summarized)
 #
 makeMeasure <- function(measureSpec,thisYear,Geography,QPrep_ls,measureEnv) {
-  GeoValue <- Geography["Value"]
+  if ( Geography["Type"] != "Region" ) {
+    GeoValue <- Geography["Value"]
+    byRegion <- FALSE
+  } else {
+    GeoValue <- ""
+    byRegion <- TRUE # skip processing "By" specifications
+    # Yields one less dimension on results from summarizeDatasets
+  }
   measureName <- measureSpec$Name
 
   # Skip or include measures based on presence of required Dataset
@@ -710,35 +717,45 @@ makeMeasure <- function(measureSpec,thisYear,Geography,QPrep_ls,measureEnv) {
   } else
   if ( "Summarize" %in% names(measureSpec) ) {
     sumSpec <- measureSpec$Summarize
-    if ( ! "By" %in% names(sumSpec) ||
-         ! Geography["Type"] %in% sumSpec$By ) {
-      stop(paste("Script wants Geography Type ",Geography["Type"]," in 'By' but got ",sumSpec$By,"",sep="'"))
+    if ( ! byRegion ) {
+      if ( ! "By" %in% names(sumSpec) ||
+           ! Geography["Type"] %in% sumSpec$By ) {
+        stop(paste("Script wants Geography Type ",Geography["Type"]," in 'By' but got ",sumSpec$By,"",sep="'"))
+      }
     }
     usingBreaks <- "Breaks" %in% names(sumSpec) && ! is.null(sumSpec$Breaks)
     usingKey <- "Key" %in% names(sumSpec) && ! is.null(sumSpec$Key)
     measure <- visioneval::summarizeDatasets(
         Expr = sumSpec$Expr,
         Units = sumSpec$Units,
-        By_ = sumSpec$By,
+        By_ = if ( ! byRegion || usingBreaks ) sumSpec$By else NULL,
         Breaks_ls = if ( usingBreaks) sumSpec$Breaks else NULL,
         Table = sumSpec$Table,
         Key = if ( usingKey ) sumSpec$Key else NULL,
         Group = thisYear,
         QueryPrep_ls = QPrep_ls
       )
-    if ( ! usingBreaks ) {
-      measure <- measure[GeoValue]
+    if ( ! byRegion && ! usingBreaks ) {
[…]

### 2021-04-20 sources/models/VEScenario/run_model.R
Add Sample VEScenario Model

@@ -1,0 +1,39 @@
+#===========
+#run_model.R
+#===========
+
+#This script demonstrates the VisionEval framework for a demonstration of VEScenario model
+
+#Load libraries
+#--------------
+cat('run_model.R: script entered\n')
+library(visioneval)
+cat('run_model.R: libraries loaded\n')
+
+planType <- 'callr'
+ptm <- proc.time()
+#Initialize model
+#----------------
+initializeModel(
+  ModelScriptFile = "run_model.R",
+  ParamDir = "defs",
+  RunParamFile = "run_parameters.json",
+  GeoFile = "geo.csv",
+  ModelParamFile = "model_parameters.json",
+  LoadDatastore = FALSE,
+  DatastoreName = NULL,
+  SaveDatastore = TRUE
+)
+#Run all demo module for all years
+#---------------------------------
+for(Year in getYears()) {
+  runModule(
+    ModuleName = "RunScenarios",
+    PackageName = "VEScenario",
+    RunFor = "AllYears",
+    RunYear = Year
+  )
+}
+cat('run_model.R: run complete.\n')
+proc.time() - ptm
+

### 2021-04-20 sources/models/VEScenario/tutorial_VEScenario.Rmd
Create tutorial_VEScenario.Rmd

---
title: "Introduction to VERRSPM_Scenario"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to VETRSPM_Scenario}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Strategic planning often requires the assessment of large numbers of future scenarios, each assessing a different combination of prioritizations, policy decisions, and constraints. The VEScenarios module provides the capability to quickly set up, run, and visualize large numbers of VERSPM scenarios using a baseline scenario combined with multiple changes to model inputs.

Both VERSPM and VERPAT models are capable of running multiple scenarios efficiently.

### Scenario manager database
The first step to run VeScenario is to build the scenario manager database. 
The database should be a csv file in standard format defined and has the following attributes:

strategy_label: A concise headline label for a Strategy used in downstream visualizations, e.g., Community Design.
strategy_description: A verbose description of the Strategy explaining its meaning, e.g., “Local policies to enable shorter trips and travel by alternate modes.” 
strategy_level: An integer index unique to each strategy level, e.g., “1”. 
category_name: A single unique letter used to define the relevant Category and used to name scenario folders, e.g., “L”. 
category_label: A concise headline label for the Category used in downstream visualizations, e.g., “Land Use”. 
category_description: A verbose description of the Category explaining its meaning, e.g., “Distribution of population and employment by place type”. 
category_instructions: Additional meta-data describing the specific way in which the relevant model will impact this category, e.g., “The form in which development occurs (density, regional accessibility, mixed use, etc.) represented by the distribution of population and employment by place type”. 
policy_name: A single number used to define the relevant Policy and used to name scenario folders, e.g., “1”
policy_label: A concise headline label for the Policy used in downstream visualizations, e.g., “Base”. The keywords Base or Baseline indicated that the files from the Scenario Baseline Input Folder (see below) are not expected to be edited.  
policy_description: A verbose description of the Policy explaining its meaning, e.g., “Maintain current distribution”. 
input_label: A numeric index uniquely identifying an input file needed to implement the Policy within the model. 
input_filename: The name of the file that implements the policy.

A sample files is located in `runtime/VEScenario/scenario_inputs/verspm_scenario_manager.csv`. You can 
add as many sceanrios as you want but the format of strategies and scenarios should be standard

First, you should define the strategies you want to run. These are the default list of VERSPM_Scenarios strategies:

* **Community Design** :  Policies that seek to enable shorter trips and alternate modes such as promotion of mixed use land use, transit service, bicycling, and parking management. ( group of L, B, P , T ) 
* **Marketing/Incentive** : Policies that improve driving efficiency such as eco driving, and Intelligent Transportation System efforts, as well as programs that reduce auto demand such as carsharing, and home or work-based transportation demand management.( group of D, E ) 
* **Pricing : Policies** that move towards true cost pricing such as road user fees to pay for the cost of operating, maintaining and improving roads, pay-as-you-drive (PAYD) insurance, and environmental impact fees such as a carbon tax. ( C )
* **Vehicles/Fuels** : Factors representing changes to future vehicles and fuels ( group of V, F ) 
* **Income** : Context factor on the assumed growth of statewide average per capita income ( I )

Each strategy covers couple of categories in a lower level.

These are the current list of VERSPM_Scenarios categories:

  - B - Bicycles (`azone_prop_sov_dvmt_diverted.csv`) : Network improvements, incentives, and technologies that encourage bicycling and other light-weight vehicle travel

  - C - Vehicle Travel Cost (`azone_hh_veh_own_taxes.csv`) and (`region_prop_externalities_paid.csv`) : Combination of fuel prices and charges to pay for roadway costs and possibly externalities

  - D - DemandManagement (`bzone_travel_demand_mgt.csv`) : Programs to encourage less private vehicle travel

  - E - Driving Efficiency (`marea_operations_deployment.csv`) and (`marea_speed_smooth_ecodrive.csv`) and (`other_ops_effectiveness.csv`) : Driving efficiency by increasing implementation of ITS

  - F - Technology Mix and CI (`marea_transit_powertrain_prop.csv`) and (`region_carsvc_powertrain_prop.csv`) and (region_comsvc_powertrain_prop.csv`) : Vehicle technology mix and carbon intensity of fuels.

  - G - Fuel Price (`azone_fuel_power_cost.csv`) : Real fuel price in 2010 USD

  - I - Income (`azone_fuel_power_cost.csv`) : Real average household income in 2010 USD

  - L - LandUse (`azone_hhsize_targets.csv`) and (`bzone_urban_du_proportions.csv`) : Distribution of population and employment by place type

  - P - Parking (`bzone_parking.csv`) : The extent of paid parking and its price

  - T - Transit (`marea_transit_service.csv`) : The extent and frequency of transit service

  - V - Vehicle Characteristics (`azone_hh_veh_mean_age.csv`) and (`azone_lttrk_prop`): The combination of fuel prices and vehicle travel charges to pay for roadways and to pay for externalities such as carbon pricing

 For each category, you should define the levels you would like to run. For each category level you need to change a set of inputs based on the scenario definition.

 The input column should be populated with the input file names for each scenario level.

You can compeletly delete the default list and add new strategies for your own specific scenarios.

After creating this database when you run a sample visioneval model, a directory with all the scneario levels and required inputs will be created . see `runtime/VEScenario/scenario_inputs` for an example of the directory structure.
## Running VEScenario
After creating scenario manager file, you need to run VEScenario::NewScenarioSetInterior() function to 
create the scenario directories and populate with required inputs
                                                                                                        ```{r, echo=FALSE, message=FALSE, warning=FALSE}
VEScenario::NewScenarioSetInterior(ScenarioInputFolder=NULL,ScenarioManagerFile,ModelFolder,ScenarioBaselineInputFolder=NULL)
```                                                                                                     This function requires 4 inputs by user:

** ScenarioInputFolder — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
May be NULL, in which case ScenarioManagerFile is interpreted as the full path to the file, and ** ScenarioManagerFile is computed as basename(normalizePath(ScenarioInputFolder))
** ScenarioManagerFile — the file path of the CSV file described above, which is expected to be in the ScenarioInputFolder. 
** ModelFolder — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).
** ScenarioBaselineInputFolder — the location on the computer of the hard disk where the inputs used for the baseline scenarios are located.
May be NULL, in which case it is constructed as file.path(ModelFolder,”inputs”)

After running the code above, a new scenario set with all the folder and subfolders will be created and 
all the required inputs will be copied to the relevant scenario folder. However all the inputs are from the base mode and needs to be modified for different scenario levels. You will see all the inputs has the prefix of UNEDITED- to signal to the user that this file needs to be edited. After modifyin all the files with the correct inputs you need to remove the UNEDITED prefix from all the files names.

The next step is to build all the possible scenarios based on the scenario set by running the following code

```{r, echo=FALSE, message=FALSE, warning=FALSE}
VEScenario::BuildScenariosInterior(ScenarioInputFolder,ScenarioOutputFolder,ModelFolder)
```

This function requires 3 inputs by user:

** ScenarioInputFolder — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
** ScenarioOutputFolder — file path identifying the location on the computer hard disk where the built scenarios will be located
** ModelFolder — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).

After running this step, you can see all the built scenarios in the outputfolder. This function Copies the Base Model input files to each scenario combination folder. It also copies the run_model.R and defs folder from the ModelFolder into each of the scenario folders 
It overwrites the specific scenario input_filename files from the ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute a standalone VERPAT or VERPSM model run with the combinations of scenario input files.

The last step is to run the scenarios using VEScenario::RunScenarios module. You first need
to copy the RunScenario sample model located in the  `runtime/models/VEScenario`. From the defs folder ,
model_parameters.json needs to be modifed to point to ScenarioOutputFolder mentioned in BuildScenariosInterior step.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
VEScneario_test <- openModel("VEScenario")
VEScneario_test $run()
```

### 2021-04-20 sources/modules/VEScenario/R/BuildScenarios.R
Update VEScenario Package

@@ -120,65 +120,67 @@ visioneval::savePackageDataset(BuildScenariosSpecifications, overwrite = TRUE)
 #=======================================================
 #SECTION 3: DEFINE FUNCTIONS THAT IMPLEMENT THE SUBMODEL
 #=======================================================
-
-
 #Main module function builds scenarios
 #------------------------------------------------------------------
-#' Function to build scenarios.
+#' Interior Function to build scenarios.
 #'
-#' \code{BuildScenarios} builds structure to run mulitple scenarios.
+#' \code{BuildScenariosInterior} builds structure to run mulitple scenarios.
 #'
-#' This function builds scenarios from scenario input levels. The folder names
-#' in the scenario input folder is used for naming scenarios. The function
-#' creates folder for each scneario, copy the model files to the scenario folder,
-#' and replaces the model files with relevant scenario inputs.
+#'Check if the folder structure expected by VEScenario, which is every unique combination of 
+#'category_name and policy_name, for example B1C1D1E1F1G1I1L2P1T2V2, exists. If it does not exist, 
+#'create it. Copies the Base Model input files to each scenario combination folder.
+#'Copies the run_model.R and defs folder from the ModelFolder into each of the scenario folders 
+#'(for example B1C1D1E1F1G1I1L2P1T2V2) Overwrites the specific scenario input_filename files from the 
+#'ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute 
+#'a standalone VERPAT or VERPSM model run with the combinations of scenario input files.
+#'At this stage the user should have modified the UNEDITED- file names created by the NewScenarioSet 
+#'method. If these file names remain, the method should issue an error and identify the unedited files.
+#'If any of these actions fail (e.g. because a path is wrong or input files are missing), throw an error
 #'
-#' @param L A list containing the components listed in the Get specifications
-#' for the module.
+#' @param ScenarioInputFolder a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located
+#' @param ScenarioOutputFolder file path identifying the location on the computer hard disk where the built scenarios will be located
[…]

### 2021-04-20 sources/modules/VEScenario/R/NewScenarioSet.R

@@ -25,19 +25,19 @@ NewScenarioSetSpecifications <- list(
   # Specify Input data
   # Specify data to be loaded from the datastore
   Get = items(
-
-  item(
-    NAME = "ModelFolder",
-    TABLE = "Model",
-    GROUP = "Global",
-    TYPE = "character",
-    UNITS = "NA",
-    SIZE = 20,
-    PROHIBIT = "NA",
-    ISELEMENTOF = ""
-  )
-  ,
-  
+    
+    item(
+      NAME = "ModelFolder",
+      TABLE = "Model",
+      GROUP = "Global",
+      TYPE = "character",
+      UNITS = "NA",
+      SIZE = 20,
+      PROHIBIT = "NA",
+      ISELEMENTOF = ""
+    )
+    ,
+    
     item(
       NAME = "ScenarioInputFolder",
       TABLE = "Model",
@@ -47,33 +47,33 @@ NewScenarioSetSpecifications <- list(
       PROHIBIT = "NA",
       ISELEMENTOF = "",
       DESCRIPTION = "Scenario Input folder"
-    
+      
+    )
+    ,
+    item(
+      NAME = "ScenarioManagerFile",
+      TABLE = "Model",
+      GROUP = "Global",
+      TYPE = "character",
+      UNITS = "NA",
+      PROHIBIT = "NA",
+      ISELEMENTOF = "",
+      DESCRIPTION = "name of scenario management csv database"
+      
+    )
+  ),
+  Set = items(
+    item(
+      NAME = "Scenarioset",
+      TABLE = "Model",
+      GROUP = "Global",
+      TYPE = "integer",
+      UNITS = "NA",
+      PROHIBIT = "NA",
+      ISELEMENTOF = "",
+      DESCRIPTION = "Returns 1 if scenario set build is complete"
+    )
   )
-  ,
-  item(
-    NAME = "ScenarioManagerFile",
-    TABLE = "Model",
-    GROUP = "Global",
-    TYPE = "character",
-    UNITS = "NA",
-    PROHIBIT = "NA",
-    ISELEMENTOF = "",
-    DESCRIPTION = "name of scenario management csv database"
-    
-  )
-),
-Set = items(
-  item(
-    NAME = "Scenarioset",
-    TABLE = "Model",
-    GROUP = "Global",
-    TYPE = "integer",
-    UNITS = "NA",
-    PROHIBIT = "NA",
-    ISELEMENTOF = "",
[…]

### 2021-04-20 sources/modules/VEScenario/R/RunScenarios.R

@@ -39,16 +39,6 @@ RunScenariosSpecifications <- list(
   # Specify data to be loaded from the datastore
   Get = items(
     item(
-      NAME = "ScenarioInputFolder",
-      TABLE = "Model",
-      GROUP = "Global",
-      TYPE = "character",
-      UNITS = "NA",
-      SIZE = 20,
-      PROHIBIT = "NA",
-      ISELEMENTOF = ""
-    ),
-    item(
       NAME = "ScenarioOutputFolder",
       TABLE = "Model",
       GROUP = "Global",
@@ -411,8 +401,9 @@ RunScenarios <- function(L){
   # -------------
   # Set input directory
   ModelPath <- getwd()
-  ScenarioInputPath <- file.path(ModelPath, L$Global$Model$ScenarioOutputFolder)
-
+  ScenarioInputPath <- L$Global$Model$ScenarioOutputFolder
+  ScenarioInputPath <- L$Global$Model$ScenarioOutputFolder
+  
   # Gather Scenario file structure and names
   ScenariosPath_ar <- list.dirs(ScenarioInputPath, recursive = FALSE)
   ScenarioNames_ar <- basename(ScenariosPath_ar)
@@ -435,8 +426,7 @@ RunScenarios <- function(L){
   libs <- .libPaths() # Set .libPaths(libs) in call to child process
   
   # Update the Scenario Progress Report
-  Scenarios_df  <- read.csv(file.path(ModelPath,
-                                      L$Global$Model$ScenarioOutputFolder,
+  Scenarios_df  <- read.csv(file.path(ScenarioInputPath,
                                       "ScenarioProgressReport.csv"),
                             stringsAsFactors = FALSE)
   RunScenariosFlag_ar <- vector(mode="logical",
@@ -506,8 +496,7 @@ RunScenarios <- function(L){
                        return(output)
                      },
                      label = ScenarioName_,
-                     globals = TRUE,
-                     seed = TRUE),
+                     globals = TRUE),
                      callback = function(asyncResults){
                        # asyncResults is: list(asyncTaskName,
                        #                        taskResult,
@@ -521,8 +510,7 @@ RunScenarios <- function(L){
                        taskResult <- asyncResults[["taskResult"]]
                        ScenarioName_ <- gsub("Scenario_", "", taskName_)
                        # Check if the log file exists
-                       LogFilePath_ <- list.files(file.path(ModelPath,
[…]

### 2021-04-20 sources/models/VEScenarios/tutorial_VEScenario.Rmd
Change models/VEScenario -> models/VEScenarios
- Add VEScenarios to the build components
- Formatting and spelling fixes throughout
- Still not completing runs through the tutorial

--- 2021-04-20 sources/models/VEScenario/tutorial_VEScenario.Rmd
@@
 
-After creating this database when you run a sample visioneval model, a directory with all the scneario levels and required inputs will be created .
-see `runtime/VEScenario/scenario_inputs` for an example of the directory structure.
+After creating this database when you run a sample visioneval model, a directory with all the scenario levels and required inputs will be created .
+see `runtime/models/VEScenario/scenario_inputs` for an example of the directory structure.
+
 ## Running VEScenario
@@
 create the scenario directories and populate with required inputs
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-VEScenario::NewScenarioSetInterior(ScenarioInputFolder=NULL,ScenarioManagerFile,ModelFolder,ScenarioBaselineInputFolder=NULL)
-```                                                                                                     This function requires 4 inputs by user:
+ScenarioManagerFile = "models/VEScenario/scenario_inputs/verspm_scenario_manager"
+ModelFolder = "models/VERSPM"
+VEScenario::NewScenarioSetInterior(ScenarioManagerFile,ModelFolder,ScenarioInputFolder=NULL,ScenarioBaselineInputFolder=NULL)
+```
+
+This function requires 4 inputs by user:
 
 ** ScenarioInputFolder — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
-May be NULL, in which case ScenarioManagerFile is interpreted as the full path to the file, and ** ScenarioManagerFile is computed as basename(normalizePath(ScenarioInputFolder))
-** ScenarioManagerFile — the file path of the CSV file described above, which is expected to be in the ScenarioInputFolder. 
+In case ScenarioManagerFile is in this folder, it can be NULL.
+** ScenarioManagerFile — the file path of the CSV file described above, which is expected ( but not required) to be in the ScenarioInputFolder. 
 ** ModelFolder — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).
-** ScenarioBaselineInputFolder — the location on the computer of the hard disk where the inputs used for the baseline scenarios are located.
-May be NULL, in which case it is constructed as file.path(ModelFolder,”inputs”)
+** ScenarioBaselineInputFolder — the location on the computer of the hard disk where the inputs used for the baseline scenarios are located. In case the folder it inside the ModelFolder it can be NULL
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
+ScenarioInputFolder = "models/VEScenario/scenario_inputs"
+ScenarioOutputFolder ="models/VEScenario/scenarios"
 VEScenario::BuildScenariosInterior(ScenarioInputFolder,ScenarioOutputFolder,ModelFolder)
@@
 
-** ScenarioInputFolder — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
+** ScenarioInputFolder — described above
+** ModelFolder — Described above
 ** ScenarioOutputFolder — file path identifying the location on the computer hard disk where the built scenarios will be located
-** ModelFolder — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
+ve.run()
 VEScneario_test <- openModel("VEScenario")

### 2021-04-20 sources/modules/VEScenario/R/BuildScenarios.R

@@ -145,9 +145,7 @@ visioneval::savePackageDataset(BuildScenariosSpecifications, overwrite = TRUE)
 #' @name BuildScenariosInterior
 #' @import jsonlite
 #' @export
-BuildScenariosInterior <- function(ScenarioInputFolder , ScenarioOutputFolder , ModelFolder  ){
-  
-  
+BuildScenariosInterior <- function(ScenarioInputFolder, ScenarioOutputFolder, ModelFolder){
   
   ScenarioInputPath <- ScenarioInputFolder
   # Gather Scenario file structure and names
@@ -171,8 +169,8 @@ BuildScenariosInterior <- function(ScenarioInputFolder , ScenarioOutputFolder , 
       file.remove(ScenarioOutputFolder)
     }
     dir.create(ScenarioOutputFolder)
-  }
-  if(file.exists(file.path(ScenarioInputPath,"scenario_config.json"))){
+  } 
+  if(file.exists(file.path(ScenarioInputPath, "scenario_config.json"))){
     file.copy(file.path(ScenarioInputPath, "scenario_config.json"),
               ScenarioOutputFolder)
     ScenarioConfig_ls <- fromJSON(file.path(ScenarioInputPath, "scenario_config.json"))

### 2021-04-20 sources/modules/VEScenario/R/NewScenarioSet.R

@@ -120,7 +120,7 @@ visioneval::savePackageDataset(NewScenarioSetSpecifications, overwrite = TRUE)
 #' @name NewScenarioSetInterior
 #' @import jsonlite 
 #' @export
-   NewScenarioSetInterior <- function(ScenarioInputFolder=NULL , ScenarioManagerFile, ModelFolder, ScenarioBaselineInputFolder=NULL   ){
+   NewScenarioSetInterior <- function(ScenarioInputFolder=NULL, ScenarioManagerFile, ModelFolder, ScenarioBaselineInputFolder=NULL){
      assign("%>%",getFromNamespace("%>%","magrittr"))
      
          if (is.null( ScenarioInputFolder)) {
@@ -137,7 +137,7 @@ visioneval::savePackageDataset(NewScenarioSetSpecifications, overwrite = TRUE)
         stop("Base scenario folder directory is missing")
         
       } else {
-        ScenarioBaselineInputFolder = file.path(ModelFolder,"inputs")
+        ScenarioBaselineInputFolder = file.path(ModelFolder, "inputs")
       }
     }
     
@@ -255,7 +255,7 @@ visioneval::savePackageDataset(NewScenarioSetSpecifications, overwrite = TRUE)
     #   input_form_file_name:
     #     Standard VisionEval scenario configuration CSV file, which can have any name and
     #     file location that you wish. 
-    ve.scenario_management.make_directory_structure <- function(target_root_dir , input_form_csv){
+    ve.scenario_management.make_directory_structure <- function(target_root_dir, input_form_csv){
       dir = grepl(basename(ScenarioManagerFile),list.files(target_root_dir, full.names = TRUE))
       unlink(list.files(target_root_dir, full.names = TRUE)[!dir], recursive = TRUE)
       form_df <- .ReadScenarioFormFromCSV(input_form_csv)
@@ -270,7 +270,7 @@ visioneval::savePackageDataset(NewScenarioSetSpecifications, overwrite = TRUE)
         if (!dir.exists(level_one_dir)) dir.create(level_one_dir)
         if (!dir.exists(level_two_dir)) dir.create(level_two_dir)
         inputs_list <-  unlist(strsplit(inputs_requierd, ","))
-        #check if inputs exist in parent directry
+        #check if inputs exist in parent directory
         for (i in 1:length(inputs_list)) {
           
[…]

### 2021-04-20 sources/models/VEScenarios/tutorial_VEScenario.Rmd
Running scenario manager
- Make output folder just 'scenarios' again; this works
- Remove land use from scenario manager file, was identical to bicycle
- Fix errors in the tutorial
- Restore seed = T in RunScenarios.R

@@
 
+```{r, setup, include=FALSE}
+# When running RMarkdown file, the working directory is the location of the file itself.
+# We want to run this from the root of the VisionEval installation. It may be 've.runtime' when testing a developer-built version, or ve.root # when using an end user installed version.
+use_dir = ifelse(grepl(ve.runtime, getwd()), ve.runtime, ve.root)
+
+knitr::opts_knit$set(root.dir = use_dir)
+```
+
 ## Overview
@@
 ### Scenario manager database
-The first step to run VeScenario is to build the scenario manager database. 
+The first step to run VEScenario is to build the scenario manager database. 
 The database should be a csv file in standard format defined and has the following attributes:
@@
 
-A sample files is located in `runtime/VEScenario/scenario_inputs/verspm_scenario_manager.csv`. You can 
-add as many sceanrios as you want but the format of strategies and scenarios should be standard
+A sample files is located in `models/VEScenarios/scenario_inputs/verspm_scenario_manager.csv`. You can 
+add as many scenarios as you want but the format of strategies and scenarios should be standard.
 
@@
 
-You can compeletly delete the default list and add new strategies for your own specific scenarios.
+You can completely delete the default list and add new strategies for your own specific scenarios.
 
-After creating this database when you run a sample visioneval model, a directory with all the scenario levels and required inputs will be created .
-see `runtime/models/VEScenario/scenario_inputs` for an example of the directory structure.
+After creating this database when you run a sample VisionEval model, a directory with all the scenario levels and required inputs will be created; see `models/VEScenarios/scenario_inputs` for an example of the directory structure.
 
-## Running VEScenario
+## Running VEScenarios
+
 After creating scenario manager file, you need to run VEScenario::NewScenarioSetInterior() function to 
-create the scenario directories and populate with required inputs
+create the scenario directories and populate with required inputs:
 
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-ScenarioManagerFile = "models/VEScenario/scenario_inputs/verspm_scenario_manager"
+ScenarioManagerFile = "models/VEScenarios/scenario_inputs/verspm_scenario_manager.csv"
 ModelFolder = "models/VERSPM"
-VEScenario::NewScenarioSetInterior(ScenarioManagerFile,ModelFolder,ScenarioInputFolder=NULL,ScenarioBaselineInputFolder=NULL)
+VEScenario::NewScenarioSetInterior(ScenarioManagerFile,
+                                  ModelFolder,
+                                  ScenarioInputFolder = NULL,
+                                  ScenarioBaselineInputFolder = NULL)
 ```
@@
 
-After running the code above, a new scenario set with all the folder and subfolders will be created and 
+After running the code above, a new scenario set with all the folder and sub-folders will be created and 
 all the required inputs will be copied to the relevant scenario folder.
@@
 You will see all the inputs has the prefix of UNEDITED- to signal to the user that this file needs to be edited.
-After modifyin all the files with the correct inputs you need to remove the UNEDITED prefix from all the files names.
+After modifying all the files with the correct inputs you need to remove the UNEDITED prefix from all the files names.
 
-The next step is to build all the possible scenarios based on the scenario set by running the following code
+The next step is to build all the possible scenarios based on the scenario set by running the following code:
 
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-ScenarioInputFolder = "models/VEScenario/scenario_inputs"
-ScenarioOutputFolder ="models/VEScenario/scenarios"
-VEScenario::BuildScenariosInterior(ScenarioInputFolder,ScenarioOutputFolder,ModelFolder)
+ScenarioInputFolder = "models/VEScenarios/scenario_inputs"
+ScenarioOutputFolder ="models/VEScenarios/scenarios"
+
+if(grepl('Unedited', list.files(ScenarioInputFolder, all.files = T, recursive = T))){
+  stop('Please finish editing the input files before continuing. At least one is still `Unedited`.')
+}
+
+VEScenario::BuildScenariosInterior(ScenarioInputFolder,
+                                   ScenarioOutputFolder,
+                                   ModelFolder)
 ```
@@
 
-After running this step, you can see all the built scenarios in the outputfolder.
+After running this step, you can see all the built scenarios in the output folder.
 This function Copies the Base Model input files to each scenario combination folder.
-It also copies the run_model.R and defs folder from the ModelFolder into each of the scenario folders
-It overwrites the specific scenario input_filename files from the ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute a standalone VERPAT or VERPSM model run with the combinations of scenario input files.
+It also copies the `run_model.R` and `defs` folder from the ModelFolder into each of the scenario folders
+It overwrites the specific scenario `input_filename` files from the ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute a standalone VERPAT or VERPSM model run with the combinations of scenario input files.
 
 The last step is to run the scenarios using VEScenario::RunScenarios module. You first need
-to copy the RunScenario sample model located in the  `runtime/models/VEScenario`. From the defs folder ,
-model_parameters.json needs to be modifed to point to ScenarioOutputFolder mentioned in BuildScenariosInterior step.
+to copy the RunScenario sample model located in the  `models/VEScenarios`. From the defs folder,
+model_parameters.json needs to be modified to point to ScenarioOutputFolder mentioned in BuildScenariosInterior step.
 
@@
 ve.run()
-VEScneario_test <- openModel("VEScenario")
-VEScneario_test $run()
+VEScnearios_test <- openModel("VEScenarios")
+VEScnearios_test$run()
 ```

### 2021-04-20 sources/modules/VEScenario/R/RunScenarios.R

@@ -496,7 +496,8 @@ RunScenarios <- function(L){
                        return(output)
                      },
                      label = ScenarioName_,
-                     globals = TRUE),
+                     globals = TRUE,
+                     seed = TRUE),
                      callback = function(asyncResults){
                        # asyncResults is: list(asyncTaskName,
                        #                        taskResult,

### 2021-04-20 sources/models/VEScenarios/tutorial_VEScenario.Rmd
Run the Rmarkdown without evaluation
- User will need to step through the code themselves to modify the scenarios before BuildScenarios step

@@
 ---
-title: "Introduction to VERRSPM_Scenario"
+title: "Introduction to VEScenarios"
 date: "`r Sys.Date()`"
@@
 vignette: >
-  %\VignetteIndexEntry{Introduction to VETRSPM_Scenario}
+  %\VignetteIndexEntry{Introduction to VEScenarios}
   %\VignetteEngine{knitr::rmarkdown}
@@
 
-```{r, setup, include=FALSE}
+```{r, setup1, include=FALSE, eval=T}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+
+```{r, setup2, include=FALSE}
 # When running RMarkdown file, the working directory is the location of the file itself.
@@
 Strategic planning often requires the assessment of large numbers of future scenarios, each assessing a different combination of prioritizations, policy decisions, and constraints.
-The VEScenarios module provides the capability to quickly set up, run, and visualize large numbers of VERSPM scenarios using a baseline scenario combined with multiple changes to model inputs.
+The VEScenario module provides the capability to quickly set up, run, and visualize large numbers of VisionEval scenarios using a baseline scenario combined with multiple changes to model inputs.
 
@@
 ### Scenario manager database
-The first step to run VEScenario is to build the scenario manager database. 
+
+The first step to run VEScenarios is to build the scenario manager database. 
 The database should be a csv file in standard format defined and has the following attributes:
 
-strategy_label: A concise headline label for a Strategy used in downstream visualizations, e.g., Community Design.
-strategy_description: A verbose description of the Strategy explaining its meaning, e.g., “Local policies to enable shorter trips and travel by alternate modes.” 
-strategy_level: An integer index unique to each strategy level, e.g., “1”. 
-category_name: A single unique letter used to define the relevant Category and used to name scenario folders, e.g., “L”. 
-category_label: A concise headline label for the Category used in downstream visualizations, e.g., “Land Use”. 
-category_description: A verbose description of the Category explaining its meaning, e.g., “Distribution of population and employment by place type”. 
+- strategy_label: A concise headline label for a Strategy used in downstream visualizations, e.g., Community Design.
+- strategy_description: A verbose description of the Strategy explaining its meaning, e.g., “Local policies to enable shorter trips and travel by alternate modes.” 
+- strategy_level: An integer index unique to each strategy level, e.g., “1”. 
+- category_name: A single unique letter used to define the relevant Category and used to name scenario folders, e.g., “L”. 
+- category_label: A concise headline label for the Category used in downstream visualizations, e.g., “Land Use”. 
+- category_description: A verbose description of the Category explaining its meaning, e.g., “Distribution of population and employment by place type”. 
 category_instructions: Additional meta-data describing the specific way in which the relevant model will impact this category, e.g., “The form in which development occurs (density, regional accessibility, mixed use, etc.) represented by the distribution of population and employment by place type”.
-policy_name: A single number used to define the relevant Policy and used to name scenario folders, e.g., “1”
+- policy_name: A single number used to define the relevant Policy and used to name scenario folders, e.g., “1”
 policy_label: A concise headline label for the Policy used in downstream visualizations, e.g., “Base”.
 The keywords Base or Baseline indicated that the files from the Scenario Baseline Input Folder (see below) are not expected to be edited.
-policy_description: A verbose description of the Policy explaining its meaning, e.g., “Maintain current distribution”. 
-input_label: A numeric index uniquely identifying an input file needed to implement the Policy within the model. 
-input_filename: The name of the file that implements the policy.
+- policy_description: A verbose description of the Policy explaining its meaning, e.g., “Maintain current distribution”. 
+- input_label: A numeric index uniquely identifying an input file needed to implement the Policy within the model. 
+- input_filename: The name of the file that implements the policy.
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r new_scenario_set}
 ScenarioManagerFile = "models/VEScenarios/scenario_inputs/verspm_scenario_manager.csv"
@@
 
-** ScenarioInputFolder — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
+- **ScenarioInputFolder** — a file path identifying the location on the computer hard disk where the inputs for the scenarios will be located.
 In case ScenarioManagerFile is in this folder, it can be NULL.
-** ScenarioManagerFile — the file path of the CSV file described above, which is expected ( but not required) to be in the ScenarioInputFolder. 
-** ModelFolder — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).
-** ScenarioBaselineInputFolder — the location on the computer of the hard disk where the inputs used for the baseline scenarios are located. In case the folder it inside the ModelFolder it can be NULL
+- **ScenarioManagerFile** — the file path of the CSV file described above, which is expected ( but not required) to be in the ScenarioInputFolder. 
+- **ModelFolder** — The path to the folder containing the Base Model run_model.R  script that will run each of the constructed scenarios plus its other elements (inputs and defs sub-folders).
+- **ScenarioBaselineInputFolder** — the location on the computer of the hard disk where the inputs used for the baseline scenarios are located.
+In case the folder it inside the ModelFolder it can be NULL
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r build_scenarios}
 ScenarioInputFolder = "models/VEScenarios/scenario_inputs"
-ScenarioOutputFolder ="models/VEScenarios/scenarios"
+ScenarioOutputFolder = "models/VEScenarios/scenarios"
 
@@
 
-** ScenarioInputFolder — described above
-** ModelFolder — Described above
-** ScenarioOutputFolder — file path identifying the location on the computer hard disk where the built scenarios will be located
+- **ScenarioInputFolder** — described above
+- **ModelFolder** — Described above
+- **ScenarioOutputFolder** — file path identifying the location on the computer hard disk where the built scenarios will be located
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r run_scenarios}
 ve.run()

### 2021-05-11 sources/models/VEScenarios/tutorial_VEScenario.Rmd
final changes to VEScenario package

@@
 
-if(grepl('Unedited', list.files(ScenarioInputFolder, all.files = T, recursive = T))){
+if(any(grepl('Unedited', list.files(ScenarioInputFolder, all.files = T, recursive = T)))){
   stop('Please finish editing the input files before continuing. At least one is still `Unedited`.')

### 2021-05-11 sources/modules/VEScenario/R/BuildScenarios.R

@@ -389,7 +389,7 @@ BuildScenarios <- function(L){
   ScenarioOutputFolder <- file.path(RunDir, L$Global$Model$ScenarioOutputFolder)
   ModelFolder <- file.path(RunDir, L$Global$Model$ModelFolder)
 
-  BuildScenarios_interior(ScenarioInputFolder , ScenarioOutputFolder , ModelFolder) 
+  BuildScenariosInterior(ScenarioInputFolder , ScenarioOutputFolder , ModelFolder) 
   Out_ls <- initDataList()
   Out_ls$Global$Model <- list(CompleteBuild = 1L,
                               InputLabels = InputLabels_ar)

### 2021-05-11 sources/modules/VEScenario/R/NewScenarioSet.R

@@ -325,7 +325,7 @@ NewScenarioSet <- function(L){
   ModelFolder <- file.path(RunDir, L$Global$Model$ModelFolder)
   ScenarioBaselineInputFolder <- file.path(ModelFolder,"inputs")
   
-  NewScenarioSet_interior(ScenarioInputFolder , ScenarioManagerFile, ModelFolder, ScenarioBaselineInputFolder)   
+  NewScenarioSetInterior(ScenarioInputFolder , ScenarioManagerFile, ModelFolder, ScenarioBaselineInputFolder)   
   Out_ls <- initDataList()
   Out_ls$Global$Model <- list(Scenarioset = 1L)
   return(Out_ls)  

### 2021-05-20 sources/tools/tableau/make-tableau.Rmd
Initial files for tableau export/visualization, scenario manager
- Make output folder just 'scenarios' again; this works
- Remove land use from scenario manager file, was identical to bicycle
- Fix errors in the tutorial
- Restore seed = T in RunScenarios.R

---
title: "Make Tableau from Tara's Files"
output: html_notebook
---

# Overhead
```{r overhead, include = FALSE}
packages_vector <- c("tidyverse")

need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]

if (length(need_to_install)) install.packages(need_to_install)

for (package in packages_vector) {
  library(package, character.only = TRUE)
}

source("tableau_scenario_viewer.R")
```

# Remote I-O
```{r remote-io}
input_dir <- "../input/"
output_dir <- "../output/"

input_viewer_dir <- paste0(input_dir, "VERSPM/")

output_file_name <- paste0(output_dir, "for-tableau.csv")

```

# Make Scenario database
```{r}
output_df <- ve.tableau_scenario_viewer.MakeCSV(input_viewer_dir)
```

# Write Data
```{r write-data}
write_csv(output_df, path = output_file_name)
```

### 2021-09-01 sources/modules/VEModel/inst/models/VE-State/scripts-base/run_model.R
Updated standard models

@@ -1,27 +1,6 @@
-#===========
-#run_model.R
-#===========
+# Full VE-State model (takes a long time to run)
+message('VE-State: Model script entered\n')
 
-#This script demonstrates the VisionEval framework for the VE-State model.
-cat('run_model.R: script entered\n')
-#Load libraries
-#--------------
-library(visioneval)
-cat('run_model.R: library visioneval loaded\n')
-
-#Initialize model
-#----------------
-initializeModel(
-  ModelScriptFile = "run_model.R",
-  ParamDir = "defs",
-  RunParamFile = "run_parameters.json",
-  GeoFile = "geo.csv",
-  ModelParamFile = "model_parameters.json",
-  LoadDatastore = FALSE,
-  DatastoreName = NULL,
-  SaveDatastore = TRUE
-  )  
-cat('run_model.R: initializeModel completed\n')
 #Run all demo module for all years
 #---------------------------------
 for(Year in getYears()) {
@@ -64,5 +43,4 @@ for(Year in getYears()) {
   runModule("CalculateComEnergyAndEmissions",   "VETravelPerformance",   RunFor = "AllYears",    RunYear = Year)
   runModule("CalculatePtranEnergyAndEmissions", "VETravelPerformance",   RunFor = "AllYears",    RunYear = Year)
 }
-cat('run_model.R: run complete.\n')
-
+message('VE-State: Run Complete\n')

### 2021-09-01 sources/modules/VEModel/inst/models/VE-State/scripts-staged/run_stage_1.R

@@ -3,26 +3,7 @@
 #===================
 #This run_model.R script runs all of the VE modules that synthesize households for a scenario. 
 
-cat('run_model.R: Stage 1 script entered\n')
-
-#Load libraries
-#--------------
-library(visioneval)
-
-#Initialize model
-#----------------
-initializeModel(
-  ModelScriptFile = "run_model.R",
-  ParamDir = "../Stage-1/defs",
-  RunParamFile = "run_parameters.json",
-  GeoFile = "geo.csv",
-  ModelParamFile = "model_parameters.json",
-  LoadDatastore = FALSE,
-  DatastoreName = NULL,
-  SaveDatastore = TRUE
-  )  
-cat('run_model.R: initializeModel completed\n')
-
+message('VE-Stage Staged: Stage 1 script entered\n')
 #Run modules to synthesize households for all years
 #--------------------------------------------------
 for(Year in getYears()) {
@@ -31,5 +12,4 @@ for(Year in getYears()) {
   runModule("AssignLifeCycle",                 "VESimHouseholds",       RunFor = "AllYears",    RunYear = Year)
   runModule("PredictIncome",                   "VESimHouseholds",       RunFor = "AllYears",    RunYear = Year)
 }
- 
-cat('run_model.R: Stage 1 complete.\n')
+message('VE-Stage Staged: Stage 1 complete.\n')

### 2021-09-01 sources/modules/VEModel/inst/models/VE-State/scripts-staged/run_stage_2.R

@@ -3,26 +3,7 @@
 #===================
 # This run_model.R script runs all of the VE modules to model a land use and transportation scenario
 
-cat('run_model.R: Stage 2 script entered\n')
-
-#Load libraries
-#--------------
-library(visioneval)
-
-#Initialize model
-#----------------
-initializeModel(
-  ModelScriptFile = "run_model.R",
-  ParamDir = "../stage-2/defs",
-  RunParamFile = "run_parameters.json",
-  GeoFile = "geo.csv",
-  ModelParamFile = "model_parameters.json",
-  LoadDatastore = TRUE,
-  DatastoreName = "../stage-1/Datastore",
-  SaveDatastore = TRUE
-  )
-cat("run_model.R: Stage 2 Initialization Complete")
-
+message('VE-Stage Staged: Stage 2 script entered\n')
 #Run core models for land use and travel
 #---------------------------------------
 for(Year in getYears()) {
@@ -48,5 +29,4 @@ for(Year in getYears()) {
   runModule("CalculateVehicleTrips",           "VEHouseholdTravel",     RunFor = "AllYears",    RunYear = Year)
   runModule("DivertSovTravel",                 "VEHouseholdTravel",     RunFor = "AllYears",    RunYear = Year)
 }
-cat('run_model.R: Stage 2 run complete.\n')
-
+message('VE-State Staged: Stage 2 complete.\n')

### 2021-09-01 sources/modules/VEModel/inst/models/VE-State/scripts-staged/run_stage_3.R

@@ -3,31 +3,11 @@
 #===================
 # This run_model.R script runs all of the VE modules to model the adopted plans powertrains and fuels scenario
 
-cat('run_model.R: Stage 3 script entered\n')
-
-#Load libraries
-#--------------
-library(visioneval)
-
-#Initialize model
-#----------------
-initializeModel(
-  ModelScriptFile = "run_model.R",
-  ParamDir = "../stage-3/defs",
-  RunParamFile = "run_parameters.json",
-  GeoFile = "geo.csv",
-  ModelParamFile = "model_parameters.json",
-  LoadDatastore = TRUE,
-  DatastoreName = "../stage-2/Datastore",
-  SaveDatastore = TRUE
-  )  
-
+message('VE-Stage Staged: Stage 3 script entered\n')
 #Run all demo module for all years
 #---------------------------------
-# requirePackage("VEHouseholdTravel")
 for(Year in getYears()) {
   runModule("CalculateCarbonIntensity",        "VEPowertrainsAndFuels", RunFor = "AllYears",    RunYear = Year)
   runModule("AssignHhVehiclePowertrain",       "VEPowertrainsAndFuels", RunFor = "AllYears",    RunYear = Year)
 }
-
-cat('run_model.R: Stage 3 script complete\n')
+message('VE-State Staged: Stage 3 complete\n')

### 2021-09-01 sources/modules/VEModel/inst/models/VE-State/scripts-staged/run_stage_4.R

@@ -3,26 +3,8 @@
 #===================
 # This run_model.R script runs all of the VE modules to model a travel performance scenario (e.g. pricing)
 
-cat('run_model.R: Stage 4 script entered\n')
-
-#Load libraries
-#--------------
-library(visioneval)
-
-#Initialize model
-#----------------
-initializeModel(
-  ModelScriptFile = "run_model.R",
-  ParamDir = "../stage-4/defs",
-  RunParamFile = "run_parameters.json",
-  GeoFile = "geo.csv",
-  ModelParamFile = "model_parameters.json",
-  LoadDatastore = TRUE,
-  DatastoreName = "../stage-3/Datastore",
-  SaveDatastore = TRUE
-  )  
-
-#Run all demo module for all years
+message('VE-Stage Staged: Stage 4 script entered\n')
+#Run VMT Feedback loop and final measures
 #---------------------------------
 for(Year in getYears()) {
   for (i in 1:2) {
@@ -37,5 +19,4 @@ for(Year in getYears()) {
   runModule("CalculateComEnergyAndEmissions",   "VETravelPerformance",   RunFor = "AllYears",    RunYear = Year)
   runModule("CalculatePtranEnergyAndEmissions", "VETravelPerformance",   RunFor = "AllYears",    RunYear = Year)
 }
-
-cat('run_model.R: Stage 4 script complete\n')
+message('VE-Stage Staged: Stage 4 script complete\n')

### 2021-09-02 sources/modules/VEModel/extensions/Model Run/Single_Scenarios_Results.Rmd
Rationalizing stage construction; starting on scenarios

---
title: "VE-RSPM Single Scenarios Results"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet -->

# Metropolitan Data Visualization

Visualize daily VMT for all scenarios. This is done by combining seven DVMT outputs by location and type of mode. Percent change is calculated relative to the base scenario value.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

# This sections sets the location of the VisionEval installation for each of the Volpe team members.
# The result is that the object `ve.runtime` is the full path to where VisionEval is installed, and the results from the single scenarios runs are located.

current_file_directory <- dirname(rstudioapi::getActiveDocumentContext()$path)
ve.runtime <- dirname((dirname(current_file_directory)))

#ve.runtime <- file.path(dirname('~/'), 'Desktop', 'VE_Tool_Test')

if(!dir.exists(ve.runtime)){
  stop('Please specify `ve.runtime` path in the setup chunk of code. \n This will be the path to your installed version of VisionEval')
}

# Make a directory to save JPEG outputs as well
fig_save = file.path(ve.runtime, 'models', 'VERSPM_Scenarios', 'Single_Scenarios_Figures')
if(!dir.exists(fig_save)) { dir.create(fig_save) }

knitr::opts_knit$set(root.dir = ve.runtime)

library(tidyverse)
library(plotly) #install.packages('plotly')
library(DT) # install.packages('DT')

overview_file = file.path(ve.runtime, 
                          'models', 'VERSPM_Scenarios',
                          'Single_Scenarios_Overview_Template.xlsx')

if(!file.exists(overview_file)){
  stop('Please copy over the file `Single_Scenarios_Overview_Template.xlsx` to your VisionEval directory, \n and place it in the `models/VERSPM_Scenarios` location')
}

# Read in scenario descriptions
scenario_desc <- readxl::read_excel(file.path(ve.runtime,
                                              'models', "VERSPM_Scenarios",
                                              'Single_Scenarios_Overview_Template.xlsx'))

# clean up trailing whitespace
scenario_values <- scenario_desc %>%
  mutate(`Model Run Name` = gsub('\\s+', '', `Model Run Name`))

# scenario_values <- scenario_values %>%
#   mutate(ModelNameLabel = ifelse(scenario_values$Level == 'NA' |
#         str_detect(scenario_values$Folder,'C0\\d'), scenario_values$`Model Run Name`,
#         paste0(scenario_values$`Model Run Name`, scenario_values$Level)))
# 
# scenario_values <- scenario_values %>%
#   mutate(`ModelNameLabel` = gsub('\\s+', '', `ModelNameLabel`))

```

## DVMT Plots

### Percent change

The following looks at percent change in each case relative to the baseline scenario for 2038.

```{r dvmt_rel}

marea_results <- read.csv(file.path(ve.runtime, 'models', 'VERSPM_Scenarios','Single_Scenarios_Metrics_Marea.csv'))

marea_results <- marea_results %>% 
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>% 
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>% 
  left_join(scenario_values, by = c("modelName" = "Model Run Name"))

marea_results <- marea_results %>%   
  mutate(modelName = ifelse(modelName == 'VERSPM_base_model', 
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))

#info for regional dvmt graphic

dvmt <- marea_results %>%
  select(modelName, contains('Dvmt')) %>%
  rowwise() %>%
  mutate(regionalDVMT = sum(c_across(contains('Dvmt'))))

dvmtbase = as.numeric(dvmt %>% filter(modelName == '0 - Base') %>% select(regionalDVMT))
#dvmtbase = as.numeric(dvmt %>% filter(modelName == 'VERSPM_NVTA') %>% select(regionalDVMT))

dvmt <- dvmt %>%
  mutate(regDVMTAbsDif = regionalDVMT - dvmtbase,
         regDVMTPerChange = (regDVMTAbsDif/dvmtbase)*100,
         color = ifelse(regDVMTPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(regDVMTPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(regDVMTAbsDif, 2),
                                                              scientific = F, 
                                                              big.mark = ','), 'mi/day'))

gp1 <- ggplot(dvmt, aes(x = modelName, y = regDVMTPerChange, fill = color,
                        text = pct_text)) + 
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Regional DVMT", x = "Scenario", y = "Percent Change") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the output figure to disk for inclusion in a manuscript
ggsave(file.path(fig_save, 'DVMT_Rel_Change.jpeg'))

# And plot dynamically for viewing.
ggplotly(gp1, tooltip = 'text')
```

### Absolute change

```{r dvmt_abs}
gp2 <- ggplot(dvmt, aes(x = modelName, y = regDVMTAbsDif, fill = color,
                        text = mi_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Absolute Change Regional DVMT (mi/day)", x = "Scenario", y = "Absolute Change")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'DVMT_Abs_Change.jpeg'))

ggplotly(gp2, tooltip = 'text')
```

## Energy Usage plots

### Truck energy usage relative change

Next we look a the relative change between scenarios for truck energy usage.

```{r truck_energy_rel}

#info for commercial heavy truck gge (sum columns k-m)
# use `matches()` to use regular expressions first to find variable names that start with
# ComSvc or HvyTrk, and then a second one to find only those which also have GGE at the end
gge <- marea_results %>%
  select(modelName, matches('^(ComSvc|HvyTrk)')) %>%
  select(modelName, matches('GGE$')) %>%
  rowwise() %>%
  mutate(ComHvyTruckGGE = sum(c_across(contains('GGE'))))

ggebase = as.numeric(gge %>% filter(modelName == '0 - Base') %>% select(ComHvyTruckGGE))
#ggebase = as.numeric(gge %>% filter(modelName == 'VERSPM_NVTA') %>% select(ComHvyTruckGGE))

gge <- gge %>%
  mutate(ComHvyTruckGGEAbsDif = ComHvyTruckGGE - ggebase,
         ComHvyTruckGGEPerChange = (ComHvyTruckGGEAbsDif/ggebase)*100,
         sign = ifelse(ComHvyTruckGGEPerChange > 0, "pos","neg"),
         pct_text = paste('Scenario:', modelName, '\n', round(ComHvyTruckGGEPerChange, 2), '%'),
         mi_text = paste('Scenario:', modelName, '\n', format(round(ComHvyTruckGGEAbsDif, 2),
                                                              scientific = F,
                                                              big.mark = ','), 'Gasoline gallone equiv.'))

# View(gge)

gp3 <- ggplot(gge, aes(x = modelName, y = ComHvyTruckGGEPerChange, fill = sign,
                       text = pct_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Percent Change Commercial Service and  Heavy Truck GGE", x = "Scenario", y = "Percent Change")+
  theme(axis.text.x = element_text(angle = 45))

ggsave(file.path(fig_save, 'Truck_GGE_Pct_Change.jpeg'))

ggplotly(gp3, tooltip = 'text')

```

## Metropolitan area data table

```{r marea_table}
marea_table <- marea_results %>% select(-names(marea_results)[names(marea_results) %in% names(scenario_values)])

val_length <- sapply(marea_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))

datatable(marea_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Complete metropolitan area results for NVTA scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2)

```

# Household Data Visualization

In addition to examining the outputs which are reported at the entire metropolitan area level, we also investigated the outputs rolled up across all households.

## Energy consumption - gasoline equivalents

```{r household_prep, cache = F}
hhfile <- readr::read_csv(file.path(ve.runtime, 'models','VERSPM_Scenarios' ,'Single_Scenarios_Metrics_Hh.csv'))

# Join in the scenario names
hhfile <- hhfile %>%
  mutate(`modelName` = gsub('\\s+', '', `modelName`)) %>%
  mutate(model_run_no = str_extract(modelName, '[A-Z]\\d{1,3}$')) %>%
  left_join(scenario_values, by = c("modelName" = "Model Run Name"))

hhfile <- hhfile %>%
  mutate(modelName = ifelse(modelName == 'VERSPM_base_model',
                            '0 - Base',
                            paste(model_run_no, `Short Name`, sep = ' - ')))

# Calculate mode split
hhfile <- hhfile %>%
  rowwise() %>%
  mutate(total_trips = sum(WalkTrips, BikeTrips, TransitTrips, VehicleTrips),
         pct_Walk = WalkTrips / total_trips,
         pct_Bike = BikeTrips / total_trips,
         pct_Transit = TransitTrips / total_trips,
         pct_Vehicle = VehicleTrips / total_trips)

scenario_values <- hhfile %>%
  filter(modelName != '0 - Base') %>% # exclude the base scenario
  group_by(modelName) %>%
  summarize(sum_Hh_GGE = sum(DailyGGE),
            sum_Hh_kWh = sum(DailyKWH),
            sum_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

base_values <- hhfile %>%
  filter(modelName == "0 - Base") %>%
  group_by(modelName) %>%
  summarize(base_Hh_GGE = sum(DailyGGE),
            base_Hh_kWh = sum(DailyKWH),
            base_Hh_CO2e = sum(DailyCO2e),
            median_Income = median(Income),
            median_OwnCost = median(OwnCost, na.rm = T),
            median_pct_Walk = median(pct_Walk),
            median_pct_Bike = median(pct_Bike),
            median_pct_Transit = median(pct_Transit),
            median_pct_Vehicle = median(pct_Vehicle)
            )

# We use base_values$... to compare the scenario values with the corresponding base values data frame

scenario_values <- scenario_values %>%
  mutate(GGEPerChange = ((sum_Hh_GGE - base_values$base_Hh_GGE) / base_values$base_Hh_GGE) * 100,
         KWHPerChange = ((sum_Hh_kWh - base_values$base_Hh_kWh) / base_values$base_Hh_kWh) * 100,
         CO2PerChange = ((sum_Hh_CO2e - base_values$base_Hh_CO2e) / base_values$base_Hh_CO2e) * 100,
         gge_color= ifelse(GGEPerChange > 0, "pos","neg"),
         gge_pct_text = paste('Scenario:', modelName, '\n', round(GGEPerChange, 2), '%'),
         gge_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_GGE, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'Gasoline gallone equiv.'),
         kwh_color= ifelse(KWHPerChange > 0, "pos","neg"),
         kwh_pct_text = paste('Scenario:', modelName, '\n', round(KWHPerChange, 2), '%'),
         kwh_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_kWh, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'kWh'),
         co2_color= ifelse(CO2PerChange > 0, "pos","neg"),
         co2_pct_text = paste('Scenario:', modelName, '\n', round(CO2PerChange, 2), '%'),
         co2_text = paste('Scenario:', modelName, '\n', format(round(sum_Hh_CO2e, 2),
                                                               scientific = F,
                                                               big.mark = ','), 'g CO[2] equivalent'))

gp4 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_GGE,
                                   text = gge_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional daily gasoline gallon equivalents", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Total.jpeg'))

ggplotly(gp4, tooltip = 'text')
```

```{r householde_ggepct}
gp5 <- ggplot(scenario_values, aes(x = modelName, y = GGEPerChange,
                                   text = gge_pct_text, fill = gge_color)) +
  geom_bar(stat="identity") + theme_minimal() +
  scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional daily gasoline gallon equivalents % Change", x = "Scenario", y = "GGE/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_GGE_Percent_Change.jpeg'))

ggplotly(gp5, tooltip = 'text')
```

### Energy consumption - electricity

```{r household_kwh}
gp6 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_kWh,
                            text = kwh_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional KWH", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Total.jpeg'))

ggplotly(gp6, tooltip = 'text')
```

```{r household_kwh_pct}
gp7 <- ggplot(scenario_values, aes(x = modelName, y = KWHPerChange,
                            text = kwh_pct_text, fill = kwh_color)) +
  geom_bar(stat="identity") + theme_minimal() +
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  labs(title = "Regional KWH Percent Change", x = "Scenario", y = "kWh/Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_kWh_Percent_Change.jpeg'))

ggplotly(gp7, tooltip = 'text')
```

### Emissions

```{r household_emissions}
gp8 <- ggplot(scenario_values, aes(x = modelName, y = sum_Hh_CO2e,
                            text = co2_text)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e", x = "Model Name", y ="g CO[2] equivalent / Day")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_CO2e_Total.jpeg'))

ggplotly(gp8, tooltip = 'text')
```

```{r household_emissions_pct}
gp9 <- ggplot(scenario_values, aes(x = modelName, y = CO2PerChange,
                            text = co2_pct_text, fill = co2_color)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Regional CO[2]e Percent Change", x = "Model Name", y ="g CO[2] equivalent / Day")+
     scale_fill_manual(values = c('midnightblue', 'tomato'),
                    guide = F) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Regional_CO2e_Percent_Change.jpeg'))

ggplotly(gp9, tooltip = 'text')

```

### Mode Choice
Mode choice across the various scenarios can also be explored.

```{r mode_choice}
# ranking by  % vehicle trips

mode_choice <- scenario_values %>%
  select(modelName, contains('median_pct')) %>%
  pivot_longer(cols = contains('median_pct'),
               names_to = 'Mode',
               values_to = 'median_pct') %>%
  mutate(Mode = gsub('median_pct_', '', Mode))

gp10 <- ggplot(mode_choice, aes(x = modelName, y = median_pct, fill = Mode)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Percent of household trips by mode", x = "Model Name", y ="Median % of trips by each mode") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Mode_Choice.jpeg'))

ggplotly(gp10)

```

### Ownership Costs

Looking at impact on household vehicle ownership costs.

```{r own_cost}
gp11 <- ggplot(scenario_values, aes(x = reorder(modelName, median_OwnCost), y = median_OwnCost)) +
  geom_bar(stat="identity") + theme_minimal() +
  labs(title = "Median vehicle ownership costs", x = "Model Name", y ="$ / household") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave(file.path(fig_save, 'Ownership_costs.jpeg'))

ggplotly(gp11)
```

## Household level data table

```{r hh_table}
scenario_table <- scenario_values %>% select(names(scenario_values)[!grepl('_text', names(scenario_values)) &
                                                 !grepl('_color', names(scenario_values))])

val_length <- sapply(scenario_table, FUN = function(x) ifelse(is.numeric(x), max(nchar(as.character(round(x, 0)))),
                                                           max(nchar(x))))

datatable(scenario_table,
          rownames = F,
          filter = 'top',
          extensions = c('Buttons','ColReorder', 'Select'),
 options = list(
   pageLength = 10,
   dom = 'BSfrtip',
   colReorder = TRUE,
   select = TRUE,
   search = list(regex = TRUE), # add regex searching to all search boxes
   buttons = c('csv', 'excel', 'colvis'),
   autoWidth = TRUE,
   scrollX = TRUE, scrollY = 600)
  ,
 caption = "Summary of household level results for Single scenarios." ) %>%
  formatRound(which(val_length >= 5 & !grepl('model', names(val_length))), 0) %>%
  formatRound(which(val_length < 5 & !grepl('model', names(val_length))), 2)

```

```{r valid_table}

library(DT)

datatable(dvmt %>% select(-regDVMTAbsDif, -color, -pct_text, -mi_text) %>%
            relocate(regionalDVMT, .after = modelName),
           rownames = F,
          filter = 'top',
          extensions = c('Buttons'),
          options = c(
            pageLength = 10,
            dom = 'BSfrtip',
            colReorder = TRUE,
            select = TRUE,
            search = list(regex = TRUE), # add regex searching to all search boxes
            buttons = list('copy', 'print',
                           list(
                             extend = 'collection',
                             buttons = c('csv', 'excel', 'pdf'),
                             text = 'Download')),
            autoWidth = TRUE,
            scrollX = TRUE, scrollY = 600),
           caption = "VMT values from NVTA scenarios" ) %>%
  formatRound(2:9, 0) %>%
  formatStyle('regionalDVMT',
              backgroundColor = 'dodgerblue')

```

```{r save}
save(file = 'Final_Single_Scenarios_results.RData',
     list = c('marea_results',
              'scenario_desc',
              'scenario_values',
              'hhfile'))

# to confirm the total CO2 values from household file:
# co2compile <- hh_compiled %>% group_by(modelName) %>% summarize(sum_CO2 = sum(DailyCO2e))
# View(co2compiled)

```

### 2021-09-02 sources/framework/visioneval/R/datastore.R

@@ -1639,7 +1639,7 @@ getFromDatastore <- function(ModuleSpec_ls, RunYear, Geo = NULL, GeoIndex_ls = N
 #' @param envir An environment from which to extract G / ModelState_ls
 #' @return A logical value which is TRUE if the data are successfully saved to
 #' the datastore.
-#' @exports
+#' @export
 setInDatastore <- function(Data_ls, ModuleSpec_ls, ModuleName, Year, Geo = NULL, GeoIndex_ls = NULL, envir=modelEnvironment()) {
   # Note: since we're writing, we'll look directly at G$Datastore rather than environment in
   # functions like checkTableExistence... DatastorePath is only used when reading...

### 2021-09-02 sources/framework/visioneval/R/visioneval.R

@@ -129,6 +129,7 @@ loadParamFile <- function(Param_ls=NULL,ModelDir=NULL) {
   ParamFile <- getRunParameter("RunParamFile",Param_ls)
   ParamPath <- findRuntimeInputFile(ParamFile,"ParamDir",Param_ls=Param_ls,StopOnError=FALSE)
   if ( ! is.na(ParamPath) ) {
+    writeLog(paste("Found run_parameters.json:",ParamPath),Level="info")
     Param_ls <- loadConfiguration(ParamPath=ParamPath,keep=Param_ls)
     ParamPath <- dirname(ParamPath)
   } else {

### 2021-09-02 sources/modules/VEModel/R/models.R

@@ -273,11 +273,14 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
       resultsDir <- visioneval::getRunParameter("ResultsDir",Param_ls=modelParam_ls)
     )
   )
+  writeLog(paste("Model Results go in:",self$modelResults),Level="info")
 
   # Check for LoadModel and LoadStage in model parameters
   if ( "LoadModel" %in% names(modelParam_ls ) ) {
+    writeLog("Processing LoadModel directive",Level="info")
     baseModel <- VEModel$new( modelPath=modelParam_ls$LoadModel )
     if ( baseModel$valid() ) {
+      writeLog(paste("LoadModel =",baseModel$modelName),Level="info")
       if ( ! "LoadStage" %in% names(modelParam_ls) ) {
         loadStage <- names(baseModel$modelStages)[length(baseModel$modelStages)]
       } else loadStage <- modelParam_ls$LoadStage
@@ -288,7 +291,7 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
       )
       modelParam_ls$LoadDatastore <- TRUE
     } else {
-      writeLog(paste("LoadModel is invalid:",modelParam_ls$LoadModel),Level="warn")
+      writeLog(paste("LoadModel present but invalid:",modelParam_ls$LoadModel),Level="warn")
       modelParam_ls$LoadDatastore <- FALSE
     }
   }
@@ -309,6 +312,7 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
       InputPath=c( modelParam_ls$ModelDir, modelParam_ls$InputPath )
     )
   }
+  # Cull input paths (directory must exist and contain InputDir (default: "inputs")
   modelParam_ls <- visioneval::addRunParameter(
     Param_ls=modelParam_ls,
     Source="VEModel::findModel",
@@ -317,6 +321,10 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
       InputDir=visioneval::getRunParameter("InputDir",Param_ls=modelParam_ls)
     )
   )
+  writeLog("Input Paths:",Level="info")
+  for ( p in modelParam_ls$InputPath ) {
+    writeLog(paste("Input Path:",p),Level="info")
+  }
 
   # Locate ParamPath for overall model
   # If it doesn't exist at the level of the model, could still set for individual stages
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/Make_Bzone_Dwelling_Units.R

@@ -1,0 +1,205 @@
+#===========
+#Make_Bzone_Dwelling_Units.R
+#===========
+#===========
+# Description: 
+#     This script will create the bzone_dwelling_units.csv file and save in the input subfolder
+#===========
+
+
+source("Data Prep/config.R")
+
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+
+# Load Census Data --------------
+# add in census api key
+fileName <- 'census_api.txt'
+
+# check if census_api.txt is in the working directory
+if(!file.exists(file.path(working_dir,fileName))){ 
+  stop(paste('Census API Key needed in as a plain text file in /n', file.path(working_dir), '/n go to https://api.census.gov/data/key_signup.html /n and save the API key as `census_api.txt`'))
+}
+
+mystring <- read_file(file.path(working_dir, fileName))
+api_key <- gsub('/r/n', '', mystring) #clean up in case text file has breaks
+
+# load census api key (get one here: https://api.census.gov/data/key_signup.html)
+census_api_key(api_key, install = TRUE)
+readRenviron("~/.Renviron") # will check R environment for past api keys
+
+
+# load census tract data
+# Dwell units uses census table S1101
+# Full table found at this link: https://api.census.gov/data/2016/acs/acs5/subject/groups/S1101.html
+
+# Download with geography
+dwell_units_raw <- get_acs(geography = "tract", table = "S1101",
+                           state = state, county = counties, geometry = TRUE)
+
+#filter for our columns
+dwell_units_016 <- dwell_units_raw %>% filter(variable=="S1101_C01_016") # percent one unit households
+dwell_units_017 <- dwell_units_raw %>% filter(variable=="S1101_C01_017") # percent two unit households
+dwell_units_018 <- dwell_units_raw %>% filter(variable=="S1101_C01_018") # percent mobile home households
+dwell_units_total <- dwell_units_raw %>% filter(variable=="S1101_C01_001") # total households
+
+
+# Make sure all have the same GEOID
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_bzone_hh_inc_qrtl_prop.R

@@ -1,0 +1,294 @@
+#===========
+#make_bzone_hh_inc_qrtl_prop.R
+#===========
+#===========
+# Description: 
+#     This script will create the make_bzone_hh_inc_qrtl_prop.csv file and save in the input subfolder
+#===========
+
+
+source("Data Prep/config.R")
+
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+
+# Load Census Data --------------
+# add in census api key
+fileName <- 'census_api.txt'
+
+# check if census_api.txt is in the working directory
+if(!file.exists(file.path(working_dir,fileName))){ 
+  stop(paste('Census API Key needed in as a plain text file in /n', file.path(working_dir), '/n go to https://api.census.gov/data/key_signup.html /n and save the API key as `census_api.txt`'))
+}
+
+mystring <- read_file(file.path(working_dir, fileName))
+api_key <- gsub('/r/n', '', mystring) #clean up in case text file has breaks
+
+# load census api key (get one here: https://api.census.gov/data/key_signup.html)
+census_api_key(api_key, install = TRUE)
+readRenviron("~/.Renviron") # will check R environment for past api keys
+
+# load census tract data
+# HH income quartiles uses census table B19001
+# Full table found at this link: https://api.census.gov/data/2016/acs/acs5/groups/B19001.html
+
+# Download Census table
+hh_income_county <- get_acs(geography = "county", table = "B19001",
+                         state = state, geometry = FALSE) %>% filter(GEOID %in% counties_geoids)
+
+
+var_list <- c("B19001_002", "B19001_003", "B19001_004", "B19001_005", "B19001_006", "B19001_007", 
+              "B19001_008", "B19001_009", "B19001_010", "B19001_011","B19001_012", "B19001_013", 
+              "B19001_014", "B19001_015", "B19001_016", "B19001_017")
+
+
+estimate_list <- c(10000,14999,19999,24999,29999,34999,39999,44999,49999,59999,74999,99999,124999,149999,199999,200000)
+
+
+
+percentile_q1_num_list = 0
+percentile_q2_num_list = 0
+percentile_q3_num_list = 0
+percentile_q4_num_list = 0
+total_hh_list = 0
+tract_list = 0
+county_list = 0
+
+
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_bzone_lat_lon.R

@@ -1,0 +1,66 @@
+#===========
+#make_bzone_lat_lon.R
+#===========
+#===========
+# Description: 
+#     This script will create the bzone_lat_lon.csv input file and save in the input subfolder
+#===========
+
+
+source("Data Prep/config.R")
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+
+
+# Find centroid of bzone geometries --------------
+
+bzone_geometry$centroids <- bzone_geometry %>%
+  st_centroid() %>% 
+  st_transform(., 4326) %>% # move to standard lat lon WGS84 projection
+  st_geometry()
+
+
+# Create input dataframe --------------
+
+bzone_lat_lon <- bzone_geometry %>%
+  mutate(Geo = Bzone,
+        "Longitude" = st_coordinates(bzone_geometry$centroids)[,1] ,
+         "Latitude" = st_coordinates(bzone_geometry$centroids)[,2]) %>%
+  select(Geo, Latitude, Longitude ) %>%
+  st_set_geometry(., NULL)
+  
+
+
+# Duplicate 2019 data for 2045 ----------
+
+bzone_lat_lon_copy <- bzone_lat_lon
+bzone_lat_lon$Year <- years[1]
+bzone_lat_lon_copy$Year <- years[2]
+
+# Make final input file, save  ----------
+bzone_lat_lon_final <- rbind(bzone_lat_lon, bzone_lat_lon_copy)  
+write.csv(bzone_lat_lon_final, file.path(proj_dir, 'inputs/bzone_lat_lon.csv'), row.names = FALSE) #save as csv in final directory
+
+
+
+################################################################################
+###################### Data Quality Checks #####################################
+
+
+# This section makes a plot to check if the centroids visually look like they are in the correct spot. 
+bzone_geometry <- st_transform(bzone_geometry, 4326)
+plot(st_geometry(bzone_geometry))
+plot(st_set_geometry(bzone_geometry, 'centroids')[, 0], add = T, col = 'red', pch = 19)
+

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_default_inputs.R

@@ -1,0 +1,42 @@
+#===========
+#make_default_inputs.R
+#===========
+
+# This script will all the input files where default inputs are acceptable
+# This script will make the following files:
+#       - region_base_year_dvmt.csv
+#       - bzone_urban-mixed-use_prop.csv
+#       - region_hh_driver_adjust_prop.csv
+#       - azone_hh_veh_mean_age.csv
+#       - azone_carsvc_characteristics.csv
+#       - azone_hh_veh_own_taxes.csv
+#       - azone_payd_insurance_prop.csv
+#       - region_carsvc_powertrain_prop.csv
+#       - region_comsvc_powertrain_prop.csv
+#       - region_ave_fuel_carbon_intensity.csv
+#       - region_hvytrk_powertrain_prop.csv
+#       - azone_charging_availability.csv
+#       - marea_speed_smooth_ecodrive.csv
+#       - marea_congestion_charges.csv
+#       - azone_veh_use_taxes.csv
+#       - azone_fuel_power_cost.csv
+#       - azone_vehicle_access_times.csv
+#       - region_co2e_costs.csv
+#       - region_prop_externalities_paid.csv
+#       - region_road_cost.csv
+#       - region_comsvc_lttrk_prop.csv
+#       - region_comsvc_veh_mean_age.csv
+
+# Others to use defaults but not made in this script:
+#       - other_ops_effectiveness.csv
+
+
+# Relies on create_geo.R to be run first, to make the template files.
+# Then check to see if the named file (e.g., azone_hh_veh_age_mean.csv) has already been updated (does it have the default Rogue Valley MPO geo names, or not?). If not, proceed
+# Then join in the default columns, repeating for the base and future years for each zone as necessary.
+
+# File paths
+proj_dir = '//vntscex/dfs/Projects/PROJ-HW32A1/Task 2.9 - SHRP/SHRP2 C10-C04-C05-C16/Implementation/VisionEval/VDOT_Case_Study/'
+
+input = file.path(proj_dir, 'Data_to_process')
+

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_marea_lane_miles.R

@@ -1,0 +1,151 @@
+#===========
+#Make marea_lane_miles.R
+#===========
+#===========
+# Description: 
+#     This script will create the Make marea_lane_miles.csv input file and save in the input subfolder
+#     https://github.com/VisionEval/VisionEval/blob/master/sources/modules/VETransportSupply/inst/module_docs/AssignRoadMiles.md#marea_lane_milescsv
+#===========
+
+
+source("Data Prep/config.R")
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+library(ggplot2)
+# geo, year, FwyLaneMi, ArtLaneMi
+
+
+# Check to see if geo.csv exists; read marea from geo
+
+geo_file = file.path(proj_dir, 'defs', 'geo.csv')
+
+if(file.exists(geo_file)){
+  geo <- read.csv(geo_file)
+} else {
+  source('Data Prep/create_geo.R')
+}
+
+marea_name = unique(geo$Marea)
+
+if(length(marea_name) > 1){
+  stop('Only implemented for one Marea currently')
+}
+
+
+# Get HPMS shapefile and drop in working directory
+# files available at this location: https://www.fhwa.dot.gov/policyinformation/hpms/shapefiles_2017.cfm
+
+
+
+# Use 2017 layer
+# Downloaded from HPMS website
+year = 2017
+
+
+hpms_zip_file = file.path(working_dir,  
+                          paste0(state, year, '.zip'))
+
+
+if(!file.exists(hpms_zip_file)){
+  stop(paste('Get the', state, year, 'file from https://www.fhwa.dot.gov/policyinformation/hpms/shapefiles_2017.cfm'))
+}
+
+
+rdata_file = paste0('HPMS_', state, '_', year, '.RData')
+
+if(!file.exists(rdata_file)){
+  hpms <- st_read(file.path(proj_dir, 'Data to Process', paste0(state, year, '.shp'))) # load HPMS
+  save(list = c('hpms'), file = rdata_file)
+  
+} else {
+  load(rdata_file)
+}
+
+
+# Merge geometry with bzones
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_placetypes_USA.R

@@ -1,0 +1,235 @@
+#===========
+#make_placetypes_USA.R
+#===========
+#===========
+# Description: 
+
+# This script will create all data inputs that use Placetypes_USA/EPA Smart Locations Database as the data source
+# Input files: 
+#       - bzone_employment.csv 
+#       - bzone_unprotected_area.csv 
+#       - bzone_network_design.csv 
+#       - bzone_transit_service.csv 
+
+# Note that 5 marea input files are also listed here but these are not necessary for the final inputs:
+#       - marea_d4bpo4_adj.csv
+#       - marea_mix_targets.csv 
+#       - marea_travel-demand-mgt_by_area-type.csv
+#       - marea_parking-avail_by_area-type.csv
+#       - marea_parking-cost_by_area-type.csv
+#===========
+
+#load 2010 EPA smart location dataset
+setwd(dirname(rstudioapi::getSourceEditorContext()$path))
+load("./data/Sld_df.Rda")
+
+
+source("Data Prep/config.R")
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+
+
+# Load Census Data --------------
+# add in census api key
+fileName <- 'census_api.txt'
+
+# check if census_api.txt is in the working directory
+if(!file.exists(file.path(working_dir,fileName))){ 
+  stop(paste('Census API Key needed in as a plain text file in /n', file.path(working_dir), '/n go to https://api.census.gov/data/key_signup.html /n and save the API key as `census_api.txt`'))
+}
+
+mystring <- read_file(file.path(working_dir, fileName))
+api_key <- gsub('/r/n', '', mystring) #clean up in case text file has breaks
+
+# load census api key (get one here: https://api.census.gov/data/key_signup.html)
+census_api_key(api_key, install = TRUE)
+readRenviron("~/.Renviron") # will check R environment for past api keys
+
+# load census tract data
+# Dwell units uses census table S1101
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/make_template_files.R

@@ -1,0 +1,44 @@
+#===========
+#make_template_files.R
+#===========
+#===========
+# Description: 
+#     This file can be used to create template files for azone and bzone input files
+#     Note that this won't make a specific input file but can be useful reference to make files and check for any errors
+#===========
+source("Data Prep/config.R")
+
+# Install/Load libraries --------------
+source("Data Prep/get_packages.R")
+
+library(dplyr)
+library(tidyverse)
+library(tidycensus)
+library(viridis)
+library(leaflet)
+library(readr)
+library(sp)
+library(sf)
+library(rgeos)
+library(rlist)
+
+
+# Make file templates
+
+years = c(2019, 2045)
+
+
+azone_template = expand.grid(unique(geo$Azone), years)
+colnames(azone_template) = c('Geo', 'Year')
+
+write.csv(azone_template, 
+          file = file.path(working_dir, 'azone_template.csv'),
+          row.names = F)
+
+bzone_template = expand.grid(unique(geo$Bzone), years)
+colnames(bzone_template) = c('Geo', 'Year')
+
+write.csv(bzone_template, 
+          file = file.path(working_dir, 'bzone_template.csv'),
+          row.names = F)
+

### 2021-09-02 sources/modules/VEModel/extensions/Data Prep/references/income_quartile_script_odot.R

@@ -1,0 +1,90 @@
+#income_quartile_script.R
+#------------------------
+
+#Load income group index
+IncGroupIdx_df <- read.csv("data/census_block_group/inc_bracket_columns.csv", as.is = TRUE)
+#Load block group household income data
+ColsToSelect_ <-
+  c(IncGroupIdx_df$Column)
+IncCategoriesByBG_df <-
+  read.csv("data/census_block_group/ACS_13_5YR_B19101_with_ann.csv", 
+           as.is = TRUE, row.names = 2)[-1, ColsToSelect_]
+names(IncCategoriesByBG_df) <- c(IncGroupIdx_df$Name)
+IncCategoriesByBG_df[] <- lapply(IncCategoriesByBG_df, function(x) as.numeric(x))
+IncCategoriesByBG_df <- IncCategoriesByBG_df[Bg,]
+
+#Calculate quartiles
+IncCategories_ <- colSums(IncCategoriesByBG_df)
+barplot(IncCategories_, horiz = TRUE)
+#Number of households in each quartile
+QrtlNumHh <- sum(IncCategories_) / 4
+#Identify where quartiles fall between income categories
+QrtlNumHhCumSum_ <- QrtlNumHh * 1:3
+IncCategoriesCumSum_ <- cumsum(IncCategories_)
+#Find quartile breaks
+findPosBetween <- function(Val, Vals_) {
+  WhichLowVal <- max(which(Val > Vals_))
+  WhichHighVal <- min(which(Val < Vals_))
+  Prop <- 
+    unname(
+      (Val - Vals_[WhichLowVal]) / (Vals_[WhichHighVal] - Vals_[WhichLowVal])
+    )
+  c(LowIdx = WhichLowVal,
+    HighIdx = WhichHighVal,
+    Prop = Prop)
+}
+QrtlBreaks_ls <- 
+  lapply(QrtlNumHhCumSum_, findPosBetween, IncCategoriesCumSum_)
+#Calculate the number of households in each quartile in each block group
+calcBGHhByQuartile <- 
+  function(NumByInc_, QrtlBreaks_) {
+    NumHhByQrtl_ <- numeric(4)
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Extract_HH_Metrics.R

@@ -1,0 +1,126 @@
+# Extract metrics from a Portland Metro VERSPM model and save the results
+# Assumptions: 
+# 1. You have completed one or more VERSPM models
+# 2. These models are in the `models` directory of your VisionEval installation
+# 3. You launched VisionEval.Rproj to open this session in RStudio
+# 4. You want to extract a select set of household level metrics 
+
+# Setup ----
+library(dplyr)   # install.packages(dplyr)
+library(stringr) # install.packages(stringr)
+
+# User Input ----
+# Input the name of the completed VERSPM model
+
+modelName <- 'VERSPM'  # could be RefCase or similar
+
+
+# Extract ----
+# Define a function to extract metrics from each model folder
+
+extract_scenario_metrics <- function(modelName){
+  # Will return an error if the model doesn't exist yet
+  mod <- openModel(modelName) 
+  
+  # Household level
+  # First clear the selections
+  mod$tables <- ''
+  mod$fields <- ''
+  
+  mod$tables <- 'Household'
+  mod$fields <- c('Bzone',
+                  'Dvmt',
+                  'DailyCO2e'
+                  
+                  # Other options include:
+                  # 'Income',
+                  # 'OwnCost',
+                  # 'WalkTrips',
+                  # 'VehicleTrips',
+                  # 'BikeTrips',
+                  # 'TransitTrips',
+                  # 'DailyKWH',
+                  # Daily GGE
+                  )
+  
+  cat('Extracting \t', mod$groupsSelected, '\t',
+      mod$tablesSelected, '\n', 
+      paste(mod$fieldsSelected, collapse = '\t'))
+  
+  hh_results <- mod$extract(saveTo = F, quiet = T)
+  
+  hh_results
+}
+
+
+
+results <- extract_scenario_metrics(modelName)
+
+# This looks for a four-digit number in the names of the results tables, e.g. _2010_, and returns just the year 2010
+years_run = str_match(names(results), '(?:_)(\\d{4})(?:_)')[,2]
+
+hh_compiled <- vector() # make one big data frame for all years
+
+# Also have individual year data frames, e.g. hh_2010
+
+for(y in years_run){
+  
+  assign(paste0('hh_', y), results[[grep(y, names(results))]]) 
+  
+  hh_compiled <- rbind(hh_compiled, data.frame(Year = y, get(paste0('hh_', y))))
+  
+  # Save to csv as well
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Model Run/Create_Single_Scenarios.R

@@ -1,0 +1,98 @@
+# Launch from a VisionEval.Rproj file in an installed version of VisionEval
+
+
+library(tidyverse)
+library(tools)
+
+# User Input Requirements
+
+# scenario_inputs_path: path where scenarios are stored i.e. B-V 
+# This uses the default VERSPM_Scenarios folder. To run using your own set of scenarios and replace this path with the name of your scenarios directory.
+scenario_inputs_path <- file.path("models","VERSPM_Scenarios","scenario_inputs");
+
+# base_model_path: path where base model is stored i.e., VERSPM_base_model
+base_model_path <- file.path("models", "VERSPM_Scenarios", "VERSPM_base_model")
+
+
+# Little to no changes should be required below
+
+#get the list of all scenarios to be created
+scenario_directories <- list.dirs(scenario_inputs_path,full.names = FALSE,
+                                  recursive = FALSE)
+
+
+# iterate through all scenarios
+# Create a dataframe of scenario names
+modelNames <- vector()
+
+
+# Append base model to the list of models to be run. `files` is empty for the base model, because no files are changed. 
+modelNames <- rbind(modelNames, data.frame(name = basename(base_model_path),
+                                           files = "",
+                                           location = file.path(ve.runtime, base_model_path),
+                                           status = "Not Run"))
+
+
+for (scenario in scenario_directories){
+  cat('Preparing Scenario', scenario, '\n\n')
+  scenario_path <- file.path(scenario_inputs_path,scenario)
+  print(scenario_path)
+  single_scenario_levels <- list.files(scenario_path, full.names = FALSE, recursive = FALSE,
+                                       pattern ="[2-9]")
+  
+  
+  # run through each case excluding the 1 case
+  # changing pattern regex above specifies what levels the script should 
+  #   build folders for
+  
+  for (level in single_scenario_levels){
+    cat('\tPreparing level', level, '\t')
+    
+    # create the model name
+    base <- openModel(base_model_path)
+    modelName <- paste0('VERSPM_',scenario,level)
+    toChange <- list.files(file.path(scenario_inputs_path,scenario,level),full.names = FALSE)
+    cat('Modifying file(s):\n', paste(toChange, collapse = '\n'), '\n')
+    
+    #create the model run folder
+    if(!dir.exists(file.path(ve.runtime, "models", "VERSPM_Scenarios", modelName))){
+      runningModel <- base$copy(modelName)
+    }
+    
[…]

### 2021-09-02 sources/modules/VEModel/extensions/Model Run/Run_Single_Scenarios.R

@@ -1,0 +1,27 @@
+
+# Run the models created by Create_Single_Scenarios.R script
+# utilize Single_Scenarios_Status.csv file for model name and path
+
+# read in csv which contains models to be run
+csvpath <- file.path(ve.runtime,"models","VERSPM_Scenarios","Single_Scenarios_Status.csv")
+scenarios <- read.csv(csvpath)
+
+#Iterate through and run each model in the CSV
+for(i in 1:nrow(scenarios)){
+  
+  name <- scenarios[i, "name"]
+  modelpath <- scenarios[i,"location"]
+  
+  cat('Running', name, '\n\n')
+  
+  model <- openModel(modelpath)
+  model$run()
+  
+  # Update status when complete
+  scenarios[i, 'status'] <- model$status
+  
+}
+
+# Update all model status
+write.csv(scenarios, csvpath, row.names = F)
+

### 2021-09-02 sources/modules/VEModel/tests/test.R

@@ -134,63 +134,69 @@ test_classic <- function(modelName="VERSPM-Classic",clear=TRUE,log="info") {
   return(rs)
 }
 
-test_all_install <- function() {
-
-  owd <- setwd("models")
-  print(files<-list.dirs("models",full.names=TRUE,recursive=FALSE))
-  if ( length(files)>0 ) unlink(files,recursive=TRUE)
-  setwd(owd)
-  
-  return (
-    list(
-      # VERSPM variants
-      vr.classic = installModel("VERSPM",variant="classic",confirm=FALSE,overwrite=TRUE),
-      vr.base = installModel("VERSPM",variant="base",confirm=FALSE,overwrite=TRUE),
-      vr.year = installModel("VERSPM",variant="year",confirm=FALSE,overwrite=TRUE),
-      vr.pop = installModel("VERSPM",variant="pop",confirm=FALSE,overwrite=TRUE),
-
-      # VERPAT variants
-      vp.base = installModel("VERPAT",variant="base",confirm=FALSE,overwrite=TRUE),
-
-      # VE-State variants
-      vs.base = installModel("VE-State",variant="base",confirm=FALSE,overwrite=TRUE),
-      vs.staged = installModel("VE-State",variant="staged",confirm=FALSE,overwrite=TRUE)
-    )
-  )
-}
-
-test_install <- function(modelName="VERSPM",variant="base",installAs=NULL,log="info") {
+test_install <- function(modelName="VERSPM",variant="base",installAs="",log="info") {
 
   if ( ! missing(log) ) logLevel(log)
-
   if ( ! nzchar(variant) ) variant <- ""
-  if ( missing(installAs) || is.null(installAs) ) {
+  if ( missing(installAs) || ! nzchar(installAs) ) {
     if ( nzchar(variant) && nzchar(modelName) ) {
       installAs <- paste0("test-",modelName,"-",variant)
     }
   }
 
-  if ( nzchar(installAs) ) {
+  if ( nzchar(modelName) && nzchar(variant) && nzchar(installAs) ) {
     if ( dir.exists(file.path("models",installAs)) ) {
       testStep(paste0("Clearing previous installation at ",installAs))
       unlink(file.path("models",installAs), recursive=TRUE)
     }
 
     testStep(paste("Installing",modelName,"model from package as",installAs))
-    rs <- installModel("VERSPM",installAs,variant,log=log,confirm=FALSE)
[…]

### 2021-12-07 sources/framework/VEModel/inst/module_docs/Next-Generation.Rmd
Moved VEModel to 'framework' from 'modules'

---
title: "VisionEval Next Generation Walkthrough"
output: html_document
author: "Jeremy Raw"
---

```{ r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# VisionEval Next Generation

## Introduction

[VisionEval](https://visioneval.org) is a strategic modeling system. The "Next Generation" of
VisionEval (aka VNG, for "VisionEval Next Generation") adds a new package `VEModel` that works with an extensively refactored version of the
VisionEval framework (in the package `visioneval`) to provide a streamlined command-line interface
(CLI) for inspecting and running VisionEval models and for extracting and querying the model
results.

This document provides a walk-through of `VEModel` capabilities.

## Installing

The VNG system can be installed from Github (currently unreleased) or from a VisionEval binary
installer, just like earlier versions of VisionEval.

From within the Github source tree, it is also possible to launch a full R environment for live
testing of the `VEModel` package (a similar facility is available for the `visioneval` framework).
Start the R environment of your choice (RGUI or RStudio) and set the working directory to
sources/modules/VEModel, then `source("tests/test.R")` which will run `pkg.load` to install VEModel
and will also load some test functions that illustrate the capabilities that are also reviewed in
this document.

Processing this R Markdown file reproduces most of the `test.R` features. In particular it creates a
full VisionEval runtime space in which you can subsequently do interactive testing of VEModel
capabilities. You must have an installed "ve-lib" folder that can be placed on the R path. This
document presumes that your working directory is writeable and that "ve-lib" (all the visioneval
packages and dependencies) can be found on the R_LIBS_USER path (set explicitly as a system environment
variable, or provided through `.Renviron`). If this R Markdown file does not run for you, make sure
that your `.libPaths()` can find `ve-lib`. All the remaining work will be done in the current
directory, which should ideally be an empty directory but which may already contain a `visioneval.cnf` file
and a `models` sub-directory (more on both of those below), either from pre-creating those, or by
previously running this R Markdown in that working directory.

### 2021-12-07 build/scripts/build-inventory.R
Fixed obsolete function call in build/scripts/build-inventory.R

@@ -73,7 +73,7 @@ cat("Parsing model scripts for use of modules\n")
 model.path <- file.path(ve.runtime,"models")
 scan.models <- dir(file.path(model.path,ve.models),pattern="run_model\\.R",recursive=TRUE,full.names=TRUE)
 models <- lapply(scan.models,FUN=function(model) {
-  script <- visioneval::parseModelScript(model,TestMode=TRUE)$ModuleCalls_df
+  script <- visioneval::parseModelScript(model)$ModuleCalls_df
   df <- data.frame(MODULE=script$ModuleName,PACKAGE=script$PackageName,MODEL=TRUE)
   names(df)[3] <- basename(dirname(model))
   return(df)

### 2021-12-17 sources/framework/VEModel/R/environment.R
Developing documentation/implementation of scenarios

@@ -450,7 +450,7 @@ uniqueSources <- function(Param_ls,shorten=NULL) {
     writeLog("'sources' attribute is null in uniqueSources",.traceback(1),Level="info")
     sources <- "NULL"
   } else {
-    sources <- unique(sources$Source)
+    sources <- unique(sources)
     if ( ! is.null(shorten) ) sources <- sub(paste0(shorten,"/"),"",sources)
   }
   return(sources)
@@ -467,10 +467,12 @@ getModelIndex <- function(reset=FALSE) {
   # Uses the package global ve.env to cache models and variants
   if ( ! reset && "modelIndex" %in% names(ve.env) ) return(ve.env$modelIndex)
 
-  # WARNING: if using pkgload, this will still reach for the installed package version
-  pkgs <- utils::installed.packages(fields=c("Package","LibPath"))
-  VE.pkgs <- pkgs[ grep("^VE",pkgs[,"Package"]),"LibPath"] # named character vector of paths
-  modelPaths <- dir(file.path(VE.pkgs,names(VE.pkgs),"models"),pattern="model-index.cnf",recursive=TRUE,full.names=T)
+  # Hack for developing VEModel with pkgload
+  pkgs <- utils::installed.packages(fields="Package")
+  VE.pkgs <- grep("^VE",pkgs[,"Package"],value=TRUE)
+  if ( ! "VEModel" %in% VE.pkgs ) VE.pkgs <- c(VE.pkgs,VEModel="VEModel")
+  modelPaths <- sapply(VE.pkgs, function(p) system.file("models",package=p))
+  modelPaths <- dir(modelPaths,pattern="model-index.cnf",recursive=TRUE,full.names=T)
   modelIndex <- list()
   for ( confPath in modelPaths ) {
     confPackage <- basename(sub("[/\\]models[/\\].*","",confPath))

### 2021-12-17 sources/framework/VEModel/R/models.R

@@ -199,11 +199,11 @@ cullInputPath <- function(InputPath) {
   InputPath <- InputPath[ nzchar(InputPath) & InputPath != "." ]
 
   # Normalize remaining InputPath elements, if any, and remove duplicates
-  writeLog(paste("Culling Input Path:\n",paste(InputPath,collapse="\n")),Level="info")
+  writeLog(paste("Culling Input Path:\n",paste(InputPath,collapse="\n")),Level="debug")
   InputPath <- unique(normalizePath(InputPath,winslash="/",mustWork=FALSE))
 
   InputPath <- InputPath[dir.exists( InputPath )]
-  writeLog(paste("InputPath length after culling:",length(InputPath)),Level="info")
+  writeLog(paste("InputPath length after culling:",length(InputPath)),Level="debug")
 
   return(InputPath)
 }
@@ -273,7 +273,7 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
     # Load default parameter or get from larger runtime environment
     modelParam_ls <- visioneval::addRunParameter(
       Param_ls=modelParam_ls,
-      visioneval::getRunParameter("ResultsDir",Param_ls=Param_ls,Source=TRUE)
+      visioneval::getRunParameter("ResultsDir",Param_ls=Param_ls)
     )
   }
 
@@ -462,27 +462,6 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
         }
       )
     }
-
-    scenarioStages <- list()
-    scenarios <- self$scenarios(fromFile=fromFile)  # re-create VEModelScenario object from file
-    scenarioStages <- scenarios$stages()            # scenario stages may be an empty list
-
-    # It is possible for a model to ONLY have scenarios (if they are "manually" created)
-    # Each "scenario" in that case must be a complete model run
-    # Usually in such cases, it may be easier just to make them Reportable modelStages
-    if ( ! is.list(modelStages) ) {
-      if ( is.list(scenarioStages) && length(scenarioStages) > 0 ) {
-        modelStages <- scenarioStages
-      } else {
-        # If no stages remain, model is invalid
-        writeLog("No model stages found!",Level="error")
-        return(self)
-      }
-    } else if ( is.list(scenarioStages) && length(scenarioStages) > 0 ) {
[…]

### 2021-12-17 sources/framework/VEModel/R/query.R

@@ -102,6 +102,15 @@ ve.query.init <- function(
     stop(writeLogMessage(paste("Could not make a VEQuery from OtherQuery:",class(OtherQuery))))
   }
 
+  # Set up model and model path
+  if ( is.null(ModelPath) ) {
+    if ( ! is.null(Model) ) {
+      ModelPath <- Model$modelPath
+    } else {
+      ModelPath <- getRuntimeDirectory()
+    }
+  }
+
   # Add remaining defaults for saving the query out again
   self$attach(OtherQuery=OtherQuery,ModelPath=ModelPath,QueryDir=QueryDir,QueryName=QueryName)
 
@@ -823,8 +832,8 @@ ve.query.reload <- function( Results ) {
 }
 
 # Attach a Model and pre-load QueryResults if present
-ve.query.model <- function( Model ) {
-  if ( ! missing(Model) ) {
+ve.query.model <- function( Model=NULL ) {
+  if ( ! is.null(Model) ) {
     self$Model <- Model  # update attached model
     self$outputfile()    # Set file name for query results output
     self$results()       # cache results if available
@@ -904,7 +913,9 @@ ve.query.run <- function(
 
   if ( ! Force && length(Results)>0 ) {
     # Reload cached results (updates self$QueryResults and check for validity)
-    # private$reload returns all the Results, whether or not they have query results
+    # private$reload returns all the Results, whether or not they
+    # have query results
+    writeLog("Using cached query results",Level="warn")
     upToDate <- sapply( private$reload(Results) ,
       function(r) {
         if (
@@ -938,7 +949,7 @@ ve.query.run <- function(
   # Run the query on any out-of-date results
   # ResultsToUpdate is a list of VEResults
   if ( length(ResultsToUpdate) > 0 ) {
-    writeLog(paste("Updating Results:",sapply(ResultsToUpdate,function(x)x$Name),collapse=", "),Level="info")
+    writeLog(paste("Updating Results:",sapply(ResultsToUpdate,function(x)x$Name),collapse=", "),Level="warn")
     doQuery(
       Results=ResultsToUpdate,         # list of VEResults objects for which to generate results
       Specifications=self$getlist(),   # A list of VEQuerySpec

### 2021-12-17 sources/framework/VEModel/R/results.R

@@ -124,7 +124,7 @@ ve.results.index <- function() {
   InputDir[ is.na(InputDir) ] <- ""
   File <- sapply(ds$attributes, attributeGet, "FILE",simplify=TRUE) # should yield a character vector
   File[ is.na(File) ] <- ""
-  scenario <- visioneval::getRunParameter("Scenario",Default="Unknown Scenario",Param_ls=private$RunParam_ls)
+  scenario <- rep(visioneval::getRunParameter("Scenario",Default="Unknown Scenario",Param_ls=private$RunParam_ls),length(Description))
 
   splitGroupTableName <- strsplit(ds$groupname, "/")
   if ( length(Description) != length(splitGroupTableName) ) stop("Inconsistent table<->description correspondence")

### 2021-12-17 sources/framework/VEModel/R/scenarios.R

@@ -25,7 +25,9 @@ ve.scenario.init <- function( baseModel=NULL, fromFile=FALSE ) {
 }
 
 # Load scenario's visioneval.cnf (constructing self$RunParam_ls and self$loadParam_ls)
+# Then build whatever model stages are defined there
 ve.scenario.load <- function(fromFile=FALSE) {
+
   if ( ! fromFile && ! is.null(self$modelStages) ) return(NULL) # do not reload model stages
 
   # Reload scenario configuration file and then build the scenario stages
@@ -36,12 +38,13 @@ ve.scenario.load <- function(fromFile=FALSE) {
   # Layer in the base run parameters as basis for scenarios
 
   # If StartFrom is defined in the scenario RunParam_ls, use the parameters from that stage as the
-  # basis for the current scenarios. Otherwise, just load the parameters from the base model.
+  # basis for the current scenarios. Otherwise, just load the parameters from the base model,
+  # expecting to fill in the required missing ones here.
   if ( "StartFrom" %in% names(self$loadedParam_ls) ) {
     startFrom <- self$baseModel$modelStages[[ self$loadedParam_ls$StartFrom ]]
     baseParam_ls <- startFrom$RunParam_ls
     # Drop keys that we will force stages here to define
-    baseParam_ls <- baseParam_ls[ - which( names(baseParam_ls) %in% c("Scenario","Description") ) ]
+    baseParam_ls <- baseParam_ls[ - which( names(baseParam_ls) %in% c("Scenario","Description","ModelStages") ) ]
   } else {
     baseParam_ls <- self$baseModel$RunParam_ls
   }
@@ -51,23 +54,22 @@ ve.scenario.load <- function(fromFile=FALSE) {
 
   # Load different types of scenarios and build ModelStages for them
   writeLog("Loading model Scenarios",Level="info")
-  self$modelStages <- NULL
+
   # Explicit ModelStages are defined first (available to use as StartFrom for Category stages)
   # These will use the top-level StartFrom for the scenarios
-  if ( "ModelStages" %in% names(modelParam_ls) ) {
+  if ( "ModelStages" %in% names(self$loadedParam_ls) ) {
     writeLog(paste("Parsing explicit Scenarios from",self$scenarioDir),Level="info")
-    # TODO: overlay self$RunParam_ls on self$baseModel$RunParam_ls
-    modelStages <- lapply(names(modelParam_ls$ModelStages), # Use pre-defined structures
[…]

### 2021-12-17 sources/framework/VEModel/tests/test.R

@@ -189,7 +189,6 @@ test_all_install <- function() {
   )
   variants <- lapply(variants,function(v) v$Variant)
 
-  browser()
   installed <- sapply(
     models,
     function(m) {
@@ -1318,6 +1317,33 @@ test_setup <- function(model=NULL) {
   cat("Setting source...\n")
   print(setup$setting("Seed",defaults=TRUE,source=TRUE)) # Show source for Seed parameter: now default
   unlink(conf.file) # Don't leave the runtime visioneval.cnf around
+}
+
+test_load_scenarios <- function(useStages=TRUE, run=NULL, log="info") {
+  if ( ! missing(log) ) logLevel(log)
+
+  testStep(paste("Installing scenario model"))
+  if ( dir.exists(modelPath <- file.path("models","VERSPM-scenario")) ) unlink(modelPath,recursive=TRUE)
+  print(modelPath)
+
+  # useStages==TRUE will do the model stage scenario, otherwise combinations
+  scenarioVariant <- if ( useStages) "scenarios-ms" else "scenarios-cat"
+  testStep(paste("Installing opening variant:",scenarioVariant))
+
+  # Just install and then open it
+  mod <- installModel("VERSPM-scenario",variant=scenarioVariant,modelName="VERSPM",log=log,confirm=FALSE)
+
+  if ( missing(run) || is.null(run) ) run <- useStages # don't run combinatorial scenarios unless explicitly requested
+
+  if ( run ) {
+    mod$run()
+    qr <- mod$query("VERSPM-scenarios")
+    qr$run()
+    qr$extract()
+  }
+
+  print(mod,scenarios=TRUE)
+  return(invisible(mod))
 }
 
 # Now set it all up

### 2021-12-17 sources/framework/visioneval/R/initialization.R

@@ -1803,7 +1803,7 @@ loadModelParameters <- function(FlagChanges=FALSE,envir=modelEnvironment()) {
 #' @return A list of parsed parameters for each of the VisionEval model elements found in the script.
 #' @export
 parseModelScript <- function(FilePath) {
-  writeLog(c("Parsing model script",FilePath),Level="info")
+  writeLog(c("Parsing model script",FilePath),Level="debug")
   if (!file.exists(FilePath)) {
     Msg <- c(
       paste0("Specified model script file does not exist."),
@@ -1842,7 +1842,7 @@ parseModelScript <- function(FilePath) {
   if ( length(InitParams_ls) > 0 ) InitParams_ls <- InitParams_ls[[1]] # Ignore more than one
   RequiredVEPackages <- sapply(extractElement("requirePackage"),function(x)x$requirePackage$Package) # Vector of package names
 
-  writeLog("Done parsing model script",Level="info")
+  writeLog("Done parsing model script",Level="debug")
   return(
     list(
       AllCalls_ls        = Elements,

### 2022-01-03 build/scripts/build-config.R
Updated build scripts to improve runtime file copying

@@ -215,8 +215,6 @@ evalq(
   }
 
   # Helper function to compare package path (source) to a built target (modification date)
-  # TODO: reimplement to use fileSnapshot, and pass a list of files as "srcpath" rather than the
-  # directory - then we can do .Rbuildignore processing to ignore presence of non-copied files.
   newerThan <- function( srcpath, tgtpath, pkg.files=character(0), quiet=TRUE ) {
     # Compare modification time for a set of files to a target file
     #
@@ -245,8 +243,6 @@ evalq(
       }
       tgtpath <- file.path(tgtpath,tgtfiles)
     }
-    # TODO: implement with fileSnapshot and changedFiles? A bit hard because we're only interested
-    # in a subset of the files and fs$info keeps track of them by full path name...
     if ( length(tgtpath) < 1 ) {
       if (!quiet) cat("Newer: target files do not exist\n")
       return(TRUE)

### 2022-01-03 build/scripts/build-runtime.R

@@ -12,49 +12,8 @@ source(file.path(ve.installer,"scripts","get-runtime-config.R"))
 
 message("========== BUILD RUNTIME ENVIRONMENT (scripts, models) ==========")
 
-# Copy the runtime boilerplate
-
-# Get the boilerplate files from boilerplate.lst
-# boilerplate.lst just contains a list of the files to copy to runtime separated by
-# whitespace (easiest just to do one file/directory name per line.
-
-# Copy the boilerplate files, checking to see if what we expected was there.
-# WARNING: this won't work with boilerplate **directories**, only **files**
-
-# cat("Copying boilerplate files...\n")
-# # Set the boilperplate folder
-# ve.boilerplate <- file.path(ve.installer,"boilerplate")
-# 
-# # bp.file.list <- scan(file=file.path(ve.boilerplate, "boilerplate.lst"),
-#                      quiet=TRUE, what=character())
-# 
-# bp.files <- file.path(ve.boilerplate, bp.file.list)
-# if ( length(bp.files) > 0 ) {
-#   any.newer <- FALSE
-#   for ( f in bp.files ) {
-#     any.newer <- any( any.newer, newerThan(f,file.path(ve.runtime,basename(f))) )
-#   }
-#   success <- FALSE
-#   if ( any.newer ) {
-#     # there may be nothing to recurse into...
-#     cat(paste("Copying Boilerplate File: ", bp.files),sep="\n")
-#     success <- suppressWarnings(file.copy(bp.files, ve.runtime, recursive=TRUE))
-#     if ( ! all(success) ) {
-#       cat("WARNING!\n")
-#       print(paste("Failed to copy boilerplate: ", basename(bp.files[!success])))
-#       cat("which may not be a problem (e.g. .Rprofile missing)\n")
-#       cat(".Rprofile is principally intended to house default ve.remote for online installer.\n")
-#       cat("If something else is missing, you should revisit boilerplate/boilerplate.(bash.)lst\n")
-#     }
-#   } else {
-#     cat("Boilerplate runtime is up to date\n")
-#   }
-# } else {
-#   stop("No boilerplate files defined to setup runtime!")
-# }
-
 # Get the VisionEval runtime files
-# This will process the 'runtime'
+# This will process the 'runtime' boilerplate (e.g. VisionEval.R startup script)
 
[…]

### 2022-01-03 sources/modules/VETravelDemandMM/inst/tutorial/Response.Rmd
Updated vignettes and tutorial directory structures

---
title: "Response to Pull Request Comments"
author: "Liming Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Response to Pull Request Comments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Below are responses to the comments to the original pull request at https://github.com/gregorbj/VisionEval/wiki/Example-Review#feedback

1. Travis returns a bunch of R package warnings that should be addressed,  such as ‘DoPredictions: no visible binding for global variable ‘model’.

    **Fixed with commit 24502591e49646a936e0accdd6132c5698060a1e**.

1. The dependent data packages (NHTS2009 and SLD) should be submitted.

    **They will be included in a new/separate pull request after this PR is merged.**

1. The overall VE project travis.yaml script should be updated to include the new module and dependent packages in the automatic testing.

    **Done with commit e643502731e0d26f4fe3ddeefbfd97005c3c8c33**

1. The estimation methods should be revised to follow VE conventions. Some of these suggested revisions need to be discussed with the project team since we’re still figuring out how we all work together and incorporate additions.

    **I revised the variable names in estimation code in `data-raw`/ to follow VE conventions (commit 1b498d5760fac88c383705fd1c7c4bbf04a58080). There may be other places that I am happy to work with the review team and incorporate any apporpriate changes.**

1. update travis automated testing script to test new package

    **Done with commit e643502731e0d26f4fe3ddeefbfd97005c3c8c33**

2. revise the documentation/software to let the user know that the NHTS2009, SLD, confidential data for estimation, and estimation script are exceptions to the guidelines for various reasons.

    **A Data section is added to the Introduction vignette describing the data sources and the fact that confidential NTHS information is not included in the package (commit 6da6125cd696b17cc43302ee2a7df2899f0af871).**

3. add proof of ODOT release of ownership

    **Need Tara and Tony's help here.**

4. vignette and/or cheat sheet summarizing estimated functions and dependent variables

    **A `Variables Used in Models` subsection is added to the "Introduction to VETravelDemandMM" linking to Tara's cheat sheet (commit 6da6125cd696b17cc43302ee2a7df2899f0af871).**

1. For the software revisions, I recommend splitting any functions which alter the VE framework (the ‘helper functions’) as a separate pull request, as Brian Gregor mentioned.

    **I believe there is some confusion here - there is no function in the package that alters the VE framework. We did talk about some functions that are potentially helpful/useful to other packages, which may be better living upstream in the VE framework code. However, we didn't decide where it should be. These functions are all in R/DoPredictions.R (and, potentially, data-raw/EstModels.R), which could be easily moved to another place once we decide it. I am not sure it needs to be a separate pull request. Again I am happy to work with the review team and incorporate any apporpriate changes.**

1. The documentation is really quite thorough, and the inclusion of the submitted manuscript (while I haven’t read it) seems like an excellent addition.

1. I do have one minor comment about the documentation that I noticed in the Overview document of the package: The transit and walk TRFL models are called with the function R/PredictTransitTFL.R, not R/PredictTransitPMT.R (similar for the walk models).

    **Fixed with commit 9232f91bbb47dfa10b2569623c997fd79bb5562f**

### 2022-01-03 sources/framework/VEModel/R/models.R
Visualizer is working

@@ -483,7 +483,6 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
   scenarioStages <- scenarios$stages()            # scenario stages may be an empty list
 
   if ( length(scenarioStages) > 0 ) { # some scenarios are defined
-
     # It is possible for a model to ONLY have scenarios (if they are "manually" created)
     # Each "scenario" in that case must be a complete model run
     # Usually in such cases, it may be easier just to make them Reportable modelStages
@@ -496,7 +495,20 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
         return(self)
       }
     } else if ( is.list(scenarioStages) && length(scenarioStages) > 0 ) {
-      sapply(self$modelStages, function(s) s$Reportable <- scenarios$reportable(s$Name))
+      sapply(self$modelStages,
+        function(s) {
+          # Locate the StartFrom stage (if any) among base model stages
+          s$Reportable <- scenarios$reportable(s$Name)
+          if ( s$Reportable ) {
+            elements <- scenarios$Elements # may be NULL
+            if ( ! is.null(elements) ) elements <- sapply(elements,function(e) e$Name)
+            s$elements(
+              Names=elements,
+              update="model$configure (scenario StartFrom)" # Source for run parameter
+            ) # push default elements into StartFrom stage
+          }
+        }
+      )
       self$modelStages <- c( self$modelStages, scenarioStages )
     }
   }
@@ -504,6 +516,40 @@ ve.model.configure <- function(modelPath=NULL, fromFile=TRUE) {
   # Link the stages
   writeLog("Initializing Model Stages",Level="info")
   self$modelStages <- self$initstages( self$modelStages )
+
+  # Check for scenario element consistency (this should be taken care
+  # of automatically when ScenarioElements are loaded and built).
+  stageNames <- names(self$modelStages)
+  checkElements <- function(base,check) {
+    return(
+      length(base)>0 &&
+      ( length(base) == length(check) ) &&
+      ! is.null( names(base) ) &&
+      ! is.null( names(check) ) &&
+      all(names(base) %in% names(check))
+    )
+  }
+  scenarioElements <- character(0)
[…]

### 2022-01-03 sources/framework/VEModel/R/query.R

@@ -121,7 +121,7 @@ ve.query.init <- function(
   }
 
   # Evaluate what is present
-  self$model(Model) # Requires QueryName; also looks up possibly existing QueryResults
+  self$model(Model) # Requires QueryName; also looks up possibly existing VEQueryResults
   self$check()
   invisible(self$valid())
 }
@@ -532,7 +532,7 @@ ve.query.print <- function(details=FALSE) {
   if ( self$valid() ) {
     cat("Valid query with",length(private$QuerySpec),"elements\n")
     if ( length(private$QuerySpec) ) {
-      if ( is.list(private$QueryResults) && length(private$QueryResults)>0 ) cat("Results are available.\n")
+      if ( is.list(self$QueryResults) && length(self$QueryResults)>0 ) cat("Results are available.\n")
       print(self$names())
       if ( details ) for ( spec in private$QuerySpec ) print(spec)
     }
@@ -592,6 +592,7 @@ defaultMetadata <- c("Units","Description")
 
 # make a data.frame of all (and only) the valid query results
 ve.query.extract <- function(Results=NULL, Measures=NULL, Years=NULL,GeoType=NULL,GeoValues=NULL, metadata=TRUE, data=TRUE, exportOnly=FALSE) {
+  # "Results" is a list of VEResults (or a VEResultsList) from a VEModel
   # Visit each of the valid results in the Model (or Results list) and add its years as columns
   #  to the resulting data.frame, then return the accumulated result
   # if "metadata" is true, include default metadata in the output table
@@ -602,15 +603,15 @@ ve.query.extract <- function(Results=NULL, Measures=NULL, Years=NULL,GeoType=NUL
   # exportOnly, if TRUE, only includes metrics with the Export attribute set (conventionally to
   #   TRUE, but we're considering only present/mssing)
 
-  Results <- self$results(Results) # Results will contain list of valid query results
+  Results <- self$results(Results) # generate list of valid VEQueryResults
   if ( length(Results)==0 ) {
     stop(
       writeLog("No query results available; run the query first",Level="error")
     )
   }
 
-  valueList <- lapply(Results,function(r) structure(r$Results$Values,ScenarioName=r$Source$Name) )
-  valueList <- valueList[ ! sapply(valueList,is.null) ]
[…]

### 2022-01-03 sources/framework/VEModel/R/results.R

@@ -79,6 +79,8 @@ ve.results.modelstate <- function(ModelState_ls=NULL) {
 
 ve.results.index <- function() {
   # Load model state from self$resultsPath
+  # Note that if the model is using a non-standard ModelState name,
+  # we might not find it here. ModelState name should be set globally.
   FileName=normalizePath( file.path(
     self$resultsPath, # Should already include ResultsDir
     visioneval::getModelStateFileName()
@@ -256,7 +258,6 @@ addDisplayUnits <- function(GTN_df,Param_ls) {
   # get here with displayUnits being GTN_df, augmented by matching DisplayUnits
   return(displayUnits) # Minimally includes Group, Table, Name, DisplayUnits, DisplayUnitsFile
 }
-  
 
 ve.results.inputs <- function( fields=FALSE, module="", filename="" ) {
   # fields=TRUE, show all Datasets that originated as file inputs (lists all columns within input files)
@@ -279,6 +280,18 @@ ve.results.inputs <- function( fields=FALSE, module="", filename="" ) {
 
   ret.value <- unique(self$modelIndex[ filter, ret.fields ])
   return( ret.value[order(ret.value$InputDir,ret.value$File),] )
+}
+
+# Return a named list of ScenarioElements and Levels
+ve.results.elements <- function() {
+  # Get scenario element names plus level values for associated model stage
+  # Model stage should always have scenario elements (if not, re-run)
+  elements <- private$RunParam_ls$ScenarioElements
+  if ( !is.character(elements) ) {
+    writeLog("No scenario elements in results!",Level="error")
+    return(list())
+  }
+  return(as.list(elements)) # converted named vector to named list
 }
 
 ve.results.units <- function(selected=TRUE,display=NULL) {
@@ -546,6 +559,7 @@ VEResults <- R6::R6Class(
     queryprep=ve.results.queryprep,  # For query or other external access
     print=ve.results.print,          # summary of model results (index)
     units=ve.results.units,          # Set units on field list (modifies self$modelIndex)
+    elements=ve.results.elements,    # Get scenario elements (named list of scenario levels) for this scenario
     ModelState=ve.results.modelstate # Set/Get the model state for these results
   ),
   private = list(

### 2022-01-03 sources/framework/VEModel/R/scenarios.R

@@ -64,6 +64,7 @@ ve.scenario.load <- function(fromFile=FALSE) {
       function(stage) {
         obj <- self$loadedParam_ls$ModelStages[[stage]] # Get the stageParam_ls structure
         if ( ! "ScenarioElements" %in% names(obj) ) obj$ScenarioElements <- list() # mark as scenario
+        obj$IsScenario = TRUE # Force scenario marker into stageParam_ls
         writeLog(paste("Model Stage:",stage),Level="info")
         VEModelStage$new(
           Name=stage,
@@ -74,14 +75,16 @@ ve.scenario.load <- function(fromFile=FALSE) {
         )
       }
     )
+    names(modelStages) <- names(self$loadedParam_ls$ModelStages) # so we can look up the categoryStartFrom
   } else modelStages <- list()
   self$modelStages <- modelStages
 
   # Trigger for combination scenarios is presence of ScenarioElements
-  if ( "ScenarioElements" %in% names(modelParam_ls) ) {
+  # Must be in the scenario's visioneval.cnf (not inherited)
+  if ( "ScenarioElements" %in% names(self$loadedParam_ls) ) {
     writeLog(paste("Parsing Category combination Scenarios from",self$scenarioDir),Level="info")
 
-    self$Elements <- modelParam_ls$ScenarioElements
+    self$Elements <- self$loadedParam_ls$ScenarioElements
 
     # Add the fully built input path to each Level in the Element.
     # Those will be composed via the Category Levels into the InputPath vector for finding
@@ -108,11 +111,11 @@ ve.scenario.load <- function(fromFile=FALSE) {
     # just building the structures and throwing errors if we can't find what we need.
 
     # Compose the ScenarioCategories
-    if ( "ScenarioCategories" %in% names(modelParam_ls) ) {
+    if ( "ScenarioCategories" %in% names(self$loadedParam_ls) ) {
       # All the Category/Scenario stages will StartFrom the CategorySetting/StartFrom, which is
       # expected to be an explicit ModelStage in ScenarioDir. That explicit stage will use the
       # overall StartFrom, if any.
-      if ( ! "Category" %in% names(modelParam_ls$ScenarioCategories) ) {
+      if ( ! "Category" %in% names(self$loadedParam_ls$ScenarioCategories) ) {
         stop(
           writeLog("No ScenarioCategories present in model scenario configuration",Level="error")
         )
[…]

### 2022-01-03 sources/framework/VEModel/inst/visualizer/hide-visualizer.js

@@ -1,3 +1,4 @@
+scentitle = "Not displayed in original viewer version",
   catconfig = [
     {
       "NAME": ["Community Design"],

### 2022-01-03 sources/framework/VEModel/inst/visualizer/js/VisualVE.js

@@ -29,18 +29,27 @@ $(document).ready( function() {
 
 // Define the VisualVE (Visualizer) function that operates on loaded data
 // Expect catconfig, scenconfig, outputconfig and VEdata defined globally
+
 VisualVE = function( ) {
-  // scenconfig from scenario-cfg.js
-  var categdata = catconfig;       // from category-cfg.js = Scenario Group elements ("Community Design" etc.) - configures Scenario Input Levels
+  var scenarioData = scenconfig;
+  var categoryData = catconfig;       // from category-cfg.js = Scenario Group elements ("Community Design" etc.) - configures Scenario Input Levels
   var outputData = outputconfig;   // from output-cfg.js = Output Measure elements ( "DVMT Per Capita", etc.) - configures Model Outputs
   // VEdata from verspm.js
 
-  var categoryInputChartCount = categdata.length;
-  var outputMeasureChartCount = outputData.length;    
+  var categoryInputChartCount = categoryData.length;
+  var outputMeasureChartCount = outputData.length;
+
+  $('#ScenarioDisplayTitle').html(scentitle)
+
+  // Check if category Levels are 0-based or 1-based
+  // TODO: Ultimately rewrite description appending not to rely on numeric levels
+  // Later use of levels only checks them for "sameness", and they could be text
+  var levelOffset = 0; // Integer value of first actual scenarioData level (for compatibility with old samples)
+  if ( parseInt(categoryData[0].LEVELS[0].NAME[0])>0 ) levelOffset = 1;
 
   var scenData = [];
   var DataObjects = [];
-  $.each(catconfig,function(index,values) {
+  $.each(categoryData,function(index,values) {
       var scenarioName = values.NAME;         
       var scenarioDescription = values.DESCRIPTION;           
       var scenarioLevels = values.LEVELS;
@@ -65,10 +74,15 @@ VisualVE = function( ) {
       $('#scen' + index + '-popover-content').append('<p><strong>' + scenarioName + ':</strong> ' + scenarioDescription + '</p>');
       $.each(scenarioLevels, function(i,val){
           $.each(val.INPUTS,function(k,v){
-              $.each(scenconfig, function(idx,scenario){
+              $.each(scenarioData, function(idx,scenario){
                   if(v.NAME[0] == scenario.NAME[0]) {
                       $('#scen' + index + '-popover-content')
[…]

### 2022-01-04 sources/framework/VEModel/tests/test.R
Updated test architecture to support runtime; added scenario tests

@@ -1,27 +1,5 @@
-# Test.R
-# Comprehensively test VEModel and related interfaces
-# Also provides working examples of the API
+# Load required packages
 
-# You should run this from the VEModel package root.
-# Make sure your .libPaths includes the built ve-lib (e.g. by
-# creating a suitable .REnviron file).
-# If you're looking at this file in a built runtime, the individual test_*
-# functions are code you can try interactively to test the framework.
-# You should NOT run the "rewind" or "setup" functions...
-# The test functions are a more detailed and up-to-date version of what the
-# walkthrough does: those are the target of test-based development for the
-# framework.
-
-# starting in the root of the package, just do this
-#  source("tests/test_run.r")
-# It will create a temporary folder to use as a runtime
-# and list the available test functions.
-
-# function: pseudo_package
-# 
-if ( ! requireNamespace("pkgload",quietly=TRUE) ) {
-  stop("Missing required package: 'pkgload'")
-}
 if ( ! requireNamespace("visioneval",quietly=TRUE) ) {
   stop("Missing required package: 'visioneval'")
 }
@@ -38,77 +16,14 @@ if ( ! requireNamespace("future.callr",quietly=TRUE) ) {
   stop("Missing required package: 'future.callr'")
 }
 
+if ( ! "package:VEModel" %in% search() ) {
+  # Will be there already if running from test_setup.R inside VEModel package
+  message("Loading built VEModel package")
+  require("VEModel",quietly=TRUE)
+}
+
 logLevel <- function(log="info") {
-  initLog(Save=FALSE,Threshold=log)
-}
-
-setup <- function(ve.runtime=NULL) {
-  # Creates or uses a fresh minimal runtime environment as a sub-directory of "tests"
-  # Set VE_RUNTIME to some other location if desired (does not need to have a runtime
-  # there yet, and in fact it's better if it doesn't).
-  if ( ! is.character(ve.runtime) ) {
-    ve.runtime <- Sys.getenv("VE_RUNTIME",unset=NA)
-    if ( ! is.na(ve.runtime ) ) {
-      if ( ! dir.exists(ve.runtime) ) {
-        ve.runtime <- NA
-      }
-    }
-    if ( is.na(ve.runtime) ) {
-      ve.runtime <- grep("^(tests/)runtime.*",list.dirs("tests"),value=TRUE)[1]
-      if ( ! dir.exists(ve.runtime) ) {
[…]

### 2022-01-04 sources/framework/VEModel/tests/test_setup.R

@@ -1,0 +1,104 @@
+# test_setup.R
+# Comprehensively test VEModel and related interfaces from within the
+# VEModel Github source directory.
+
+# You should run this from the VEModel package root.
+# Make sure your .libPaths includes the built ve-lib (e.g. by
+# creating a suitable .REnviron file).
+
+# If you're looking at this file in a built runtime, you should look
+# at the test.R function the individual test_*
+# functions are code you can try interactively to test the framework.
+# You should NOT run the "rewind" or "setup" functions...
+# The test functions are a more detailed and up-to-date version of what the
+# walkthrough does: those are the target of test-based development for the
+# framework.
+
+# starting in the root of the package, just do this
+#  source("tests/test_run.r")
+# It will create a temporary folder to use as a runtime
+# and list the available test functions.
+
+if ( ! requireNamespace("pkgload",quietly=TRUE) ) {
+  stop("Missing required package: 'pkgload'")
+}
+
+# function: pseudo_package
+setup <- function(ve.runtime=NULL) {
+  # Creates or uses a fresh minimal runtime environment as a sub-directory of "tests"
+  # Set VE_RUNTIME to some other location if desired (does not need to have a runtime
+  # there yet, and in fact it's better if it doesn't).
+  if ( ! is.character(ve.runtime) ) {
+    ve.runtime <- Sys.getenv("VE_RUNTIME",unset=NA)
+    if ( ! is.na(ve.runtime ) ) {
+      if ( ! dir.exists(ve.runtime) ) {
+        ve.runtime <- NA
+      }
+    }
+    if ( is.na(ve.runtime) ) {
+      ve.runtime <- grep("^(tests/)runtime.*",list.dirs("tests"),value=TRUE)[1]
+      if ( ! dir.exists(ve.runtime) ) {
+        ve.runtime <- normalizePath(tempfile(pattern="runtime",tmpdir="tests"),winslash="/",mustWork=FALSE)
+        dir.create(ve.runtime)
+      }
+    }
+  }
+  ve.runtime <- normalizePath(ve.runtime,winslash="/",mustWork=TRUE)
+  Sys.setenv(VE_RUNTIME=ve.runtime)
+  pkgload::load_all()
+  ve.env <- VEModel::runtimeEnvironment()
[…]

### 2022-01-04 sources/framework/visioneval/tests/test.R

@@ -1,5 +1,5 @@
 # Test.R
-# Comprehensively test VEModel and related interfaces
+# Comprehensively test visioneval interfaces (basic model runs)
 # Also provides working examples of the API
 
 # You should run this from the visioneval package root.
@@ -17,127 +17,10 @@
 # It will create a temporary folder to use as a runtime
 # and list the available test functions.
 
-# function: pseudo_package
-# 
-if ( ! requireNamespace("pkgload",quietly=TRUE) ) {
-  stop("Missing required package: 'pkgload'")
-}
+# TODO: if not using pkgload from source, require visioneval
+
 if ( ! requireNamespace("stringr",quietly=TRUE) ) {
   stop("Missing required package: 'stringr'")
-}
-
-# Working directory should be the package base directory
-setup <- function(ve.runtime=NULL) {
-  # Creates or uses a fresh minimal runtime environment as a sub-directory of "tests"
-  # Set VE_RUNTIME to some other location if desired (does not need to have a runtime
-  # there yet, and in fact it's better if it doesn't).
-  if ( ! is.character(ve.runtime) ) {
-    ve.runtime <- Sys.getenv("VE_RUNTIME",unset=NA)
-    if ( ! is.na(ve.runtime ) ) {
-      if ( ! dir.exists(ve.runtime) ) {
-        ve.runtime <- NA
-      }
-    }
-    if ( is.na(ve.runtime) ) {
-      ve.runtime <- grep("^(tests/)runtime.*",list.dirs("tests"),value=TRUE)[1]
-      if ( ! dir.exists(ve.runtime) ) {
-        ve.runtime <- normalizePath(tempfile(pattern="runtime",tmpdir="tests"),winslash="/",mustWork=FALSE)
-        dir.create(ve.runtime)
-      }
-    }
-  }
-  ve.runtime <- normalizePath(ve.runtime,winslash="/",mustWork=TRUE)
-  Sys.setenv(VE_RUNTIME=ve.runtime)
-  pkgload::load_all()
-  if ( ! "ve.env" %in% search() ) {
-    message("Setup attaching ve.env")
-    ve.env <- attach(name="ve.env",NULL)
-  } else {
-    ve.env <- as.environment("ve.env")
-  }
[…]

### 2023-06-26 sources/modules/VETravelDemandMM/inst/tutorial/Review.Rmd

---
title: "VETravelDemandMM Module Contribution Review"
author: "Liming Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{VETravelDemandMM Module Contribution Review}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette documents responses to [VisionEval Contribution Review Criteria](https://github.com/gregorbj/VisionEval/wiki/Contribution-Review-Criteria) for the VETravelDemandMM module.

## Contribution Review Criteria

> 1. Does it contain all the elements that are required by the VisionEval [system specifications](https://github.com/gregorbj/VisionEval/blob/master/api/model_system_design.md)?

As a module for VisionEval and a standard R package, this package/module conforms to the recommendations in [Section 8 Module](https://github.com/gregorbj/VisionEval/blob/master/api/model_system_design.md#8-modules), in term of both required elements and the directories and files layout. The only directory included in the package but not appearing in the recommendation is `data-raw`. Here we follow Hadley Wickham's recommendation in his [R packages](http://r-pkgs.had.co.nz/data.html#data-data): "Often, the data you include in `data/` is a cleaned up version of raw data you’ve gathered from elsewhere. I highly recommend taking the time to include the code used to do this in the source version of your package [... and] put this code in `data-raw/`". The scripts in `data-raw` are code used to estimate all models and save estimation results into the `data` directory.

> 2. Why is it better, and/or different than [existing modules](https://github.com/gregorbj/VisionEval/tree/master/sources/modules)? Does it do good science and provide documentation justifying this claim? Is it consistent with good practice in strategic modeling? How might it overlap with existing modules?

The primary objectives of the module is to better represent multi-modal travel for in models for strategic planning and to update models with the latest and best data available. In addition to these two primary objectives, the module uses rigorous selection and benchmark of different model structures in choosing model structure and takes advantage of R infrastructure and new packages. Justification and objectives can be found in these project reports:

- [SPR 788 Project Report for Task 2 Model Design and Estimation Report](https://cities-lab.github.io/SPR788/Task2_Report.html)
- [SPR 788 Project Report for Task 3 VETravelDemand (VisionEval Travel Demand) Implementation](https://cities-lab.github.io/SPR788/Task3_Report.html)
- [SPR 788 Project Report for Task 4 Model Testing](https://cities-lab.github.io/SPR788/Task4_Report.html)
- [Development of a Multi-modal Travel Demand Module for the Regional Strategic Planning Model (manuscript under review)](https://www.dropbox.com/s/y594fz44achoqkq/jtlu_rspm.pdf?dl=0)

> 3. Is the module documentation complete? Does it include documentation of model estimation, algorithms, and instructions for using?

There are documents of the module in various form:

1. manual for each of the R functions in the module;
1. vignettes (including this document) that provide a long form document of the package;
1. code and comments for estimation scripts
1. Project reports and manuscripts under review, linked in the Introduction vignette

> 4. If the module allows the estimation of regional parameters, does it provide default data, does it have clear documentation of what the estimation data needs to be and how it is to be formatted, and does it include proper data specifications to ensure that the user’s input data are correct?

This module provides default model parameters estimated with US nationwide data, and it is also possible to re-estimate model paramters with region-specific data. The main estimation data are drawn from two external data package ([NHTS2009](https://github.com/cities-lab/NHTS2009) and [SLD](https://github.com/cities-lab/SLD), documented therein) and `data-raw/LoadDataforEstimation.R` joins data from different data sources and creates a single household data frame for estimation. Because the estimation script utilizes confidential residence block ground information in file hhctbg.csv (which contains these fields in character/string format: "HOUSEID","HHSTFIPS10","HHCNTYFP10","HHCT10","HHBG10","HHZIP"), users cannot directly run `data-raw/LoadDataforEstimation.R` as it is. Any user needs to obtain a copy of hhctbg.csv should contact NHTS (https://nhts.ornl.gov/contactUs.shtml). `data-raw/LoadDataforEstimation.R` provides code and comments needed to replace the estimation data with region specific data.

Since the model estimation does not use VisionEval specification, there is little check on data quality except for a few informational print out in the script.

> 5. Is it based on geographic definitions that are consistent with the model system definitions?

The module uses two geographies: Bzone (block-group) and Marea (region/UZA), which are consistent with the model system definitions.

> 6. Does the module compute quickly enough and provide documentation justifying this claim?

The tests running all 6 models in the module using Rogue Valley data take between 1 minute to 1 minutes and 30 seconds across different builds on Travis CI (See https://travis-ci.org/cities-lab/VETravelDemandMM/builds), on average 10 - 15 seconds per model, which is almost the same as the GreenSTEP/RSPM VETravelDemand module (https://travis-ci.org/gregorbj/VisionEval/builds).

> 7. Does it includes all source files and data?  If a contributed module does not include all source data, it should include a minimal example data file for testing and so it is clear what data structure is needed to run the module.  It should also include clear instructions on how to fetch the data and/or a clear explanation of why non-included data is confidential and contact information for data owners.

Except for the confidential residential block group information for households in NHTS (see also response to Question 4 above), all data and code are included in the package or in other open source (data) packages (NHTS2009 and SLD).

> 8. Does the module only call R code and packages that work on all operating systems? If the code includes any non-R code (e.g. FORTRAN, C++) will that code compile on all operating systems?

The module only includes R code and should work on all operating systems.

> 9. Is it licensed with the VisionEval [license](https//github.com/gregorbj/VisionEval/blob/master/LICENSE) that allows the code to be freely distributed and modified and includes attribution so that the ‘provenance’ of the code can be tracked?

The package is licensed with the same Apache 2.0 license as VisionEval. A LICENSE file is added to the package directory.

> 10. Does it only interact with the computing environment by returning a properly structured list to the framework (i.e. it does not modify the global environment, does not read or write files, and only calls framework functions that are allowed)?

Primary functions of the module are implemented in `R/Predict*.R`. The development of the package follows the VisionEval system design guide and the template module and only calls framework functions and a few internal helper functions. The module does not modify the global environment or read/write files when running as a VisionEval module.

> 11. Does it include regression tests to enable checking that consistent results will be returned when updates are made to the framework and/or R programming environment?

`checkModuleOutputs` provided by `testModule` checks outputs against specification of each model in automated tests of the package.

> 12. Does it include [sufficient test coverage](Automated-Testing) and test data? Does it pass the [‘testModule’](https://github.com/gregorbj/VisionEval/blob/master/api/functions_summary.md#testmodule) test which validates that it will run correctly in the model system?

Automated tests including `devtools.check()`, installation testing, and test runs of models in the module with `visioneval::testModule()` are in place with [Travis-CI](https://travis-ci.org/cities-lab/VETravelDemandMM). All tests pass in the current and recent builds.
